---
title: "Data Summarization"
output:
  ioslides_presentation:
    css: ../styles.css
    widescreen: yes
---


```{r, echo = FALSE, message=FALSE, error = FALSE}
library(knitr)
opts_chunk$set(comment = "", message = FALSE)
suppressWarnings({library(dplyr)})
library(readr)
library(tidyverse)
library(jhur)
```

<style type="text/css">
pre { /* Code block - slightly smaller in this lecture */
    font-size: 20px;
}
</style>

## Data Summarization

* Basic statistical summarization
    * `mean(x)`: takes the mean of x
    * `sd(x)`: takes the standard deviation of x
    * `median(x)`: takes the median of x
    * `quantile(x)`: displays sample quantiles of x. Default is min, IQR, max
    * `range(x)`: displays the range. Same as `c(min(x), max(x))`
    * `sum(x)`: sum of x
    * `max(x)`: maximum value in x
    * `min(x)`: minimum value in x
    * **all have a **`na.rm` for missing data 
* Transformations
    * `log` - log (base `e`) transformation
    * `log10` - log base 10 transform
    * `sqrt` - square root

## Statistical summarization

The vector getting summarized goes inside the parentheses:

```{r}
x <- c(1, 5, 7, 4, 2, 8)
mean(x)
range(x)
sum(x)
```

## Statistical summarization

Note that many of these functions have additional inputs regarding missing data, typically requiring the `na.rm` argument ("remove NAs"). 

```{r error = TRUE}
x <- c(1, 5, 7, 4, 2, 8, NA)
mean(x)
mean(x, na.rm = TRUE)
quantile(x)
quantile(x, na.rm = TRUE)
```

## Some examples

We can use the `jhu_cars` to explore different ways of summarizing data.  The `head` command displays the first rows of an object:

```{r}
library(jhur)
head(jhu_cars)
```

## Statistical summarization

Note - the `$` references/selects columns from a `data.frame`/`tibble`:

```{r}
mean(jhu_cars$hp)
quantile(jhu_cars$hp)
```

## Statistical summarization

The "tidy" way:

```{r}
jhu_cars %>% pull(hp) %>% mean() # alt: pull(jhu_cars, hp) %>% mean()
jhu_cars %>% pull(hp) %>% quantile()
```

## Statistical summarization

```{r}
jhu_cars %>% pull(wt) %>% median()
jhu_cars %>% pull(wt) %>% quantile(probs = 0.6)
```

## Data Summarization on data frames

* Basic statistical summarization
    * `rowMeans(x)`: takes the means of each row of x
    * `colMeans(x)`: takes the means of each column of x
    * `rowSums(x)`: takes the sum of each row of x
    * `colSums(x)`: takes the sum of each column of x
    * `summary(x)`: for data frames, displays the quantile information

## TB Incidence 

Let's read in a `tibble` of values from TB incidence.

If you have the `jhur` package installed successfully:

```{r}
tb <- jhur::read_tb()
```

<br>
If not, download the `xlsx` file from this link and read it in using `read_csv()`: http://jhudatascience.org/intro_to_R_class/data/tb_incidence.xlsx

## TB Incidence 

Check out the data:

```{r}
head(tb)
colnames(tb)
```

## Indicator of TB

Before we go further, let's rename the first column to be the country measured using the `rename` function in `dplyr`.

In this case, we have to use the backticks (\`) because there are spaces and funky characters in the name:

```{r}
library(dplyr)
tb <- tb %>% rename(country = `TB incidence, all forms (per 100 000 population per year)`)
```

<br>

`colnames` will show us the column names and show that country is renamed:
```{r}
colnames(tb)
```

## Summarize the data: `dplyr` `summarize` function

`dplyr::summarize` will allow you to summarize data. Format is `new = SUMMARY`. 

```{r, eval = FALSE}
# General format - Not the code!
{data object to update} <- {data to use} %>% 
                          summarize({summary column name} = {operator(source column)}) 
```

```{r}
tb %>% summarize(mean_2006 = mean(`2006`, na.rm = TRUE))
```

## Summarize the data: `dplyr` `summarize` function

`summarize` can do multiple operations at once. Just separate by a comma.

```{r}
tb %>% 
  summarize(mean_2006 = mean(`2006`, na.rm = TRUE),
            median_2007 = median(`2007`, na.rm = TRUE),
            median(`2004`, na.rm = TRUE))
```

<br>
Notice how when we forget to provide a new name, output is still provided, but the column name is messy.

## Iterative summaries: `dplyr` `summarize` and `across` functions

Use the [`across`](https://dplyr.tidyverse.org/reference/across.html) function with `summarize` to summarize across multiple columns of your data.

```{r}
tb %>% 
  summarize(across( c(`1990`, `1991`, `1992`, `1993`), ~ sum(.x, na.rm = TRUE)))
tb %>% 
  summarize(across( starts_with("2"), ~ range(.x, na.rm = TRUE)))
```

## Row means

`colMeans` and `rowMeans` require **all numeric data**. 

Let's see what the mean is across each row (country):

```{r}
tb_2 <- column_to_rownames(tb, "country") # opposite of rownames_to_column() !
head(tb_2, 2)
rowMeans(tb_2, na.rm = TRUE)
```

## Row means

`colMeans` gives you very similar output to functions we've seen previously in this lecture (`summarize` and `across`).

```{r}
colMeans(tb_2, na.rm = TRUE)
tb %>% 
  summarize(across( colnames(tb_2), ~ mean(.x, na.rm = TRUE)))
```

## `summary` Function

Using `summary` can give you rough snapshots of each column, but you would likely use `mean`, `min`, `max`, and `quantile` when necessary (and number of NAs):

```{r}
summary(tb)
```

## Lab Part 1

[Website](http://jhudatascience.org/intro_to_R_class/index.html)

## Youth Tobacco Survey

Here we will be using the Youth Tobacco Survey data:
http://jhudatascience.org/intro_to_R_class/data/Youth_Tobacco_Survey_YTS_Data.csv

```{r}
yts <- jhur::read_yts()
head(yts)
```

## Length and unique

`unique(x)` will return the unique elements of `x`

```{r, message = FALSE}
locations <- yts %>% pull(LocationDesc)
unique(locations) %>% head()
```

`length` will tell you the length of a vector. Combined with `unique`, tells you the number of unique elements:

```{r}
length(unique(locations))
```

## `table` and `dplyr`: `count` 

`table(x)` will return a frequency table of unique elements of `x`

```{r, message = FALSE}
table(locations)
```

## `table` and `dplyr`: `count` 

Use `count` directly on a data.frame and column without needing to use `pull`.

```{r, message = FALSE}
yts %>% count(LocationDesc)
```

## `table` and `dplyr`: `count` 

Multiple columns listed further subdivides the count.

```{r, message = FALSE}
yts %>% count(LocationDesc, TopicDesc)
```

# Grouping

## Perform Operations By Groups: dplyr

`group_by` allows you group the data set by grouping variables:

```{r}
#
yts
```

## Perform Operations By Groups: dplyr

`group_by` allows you group the data set by grouping variables:

```{r}
yts <- yts %>% group_by(Response)
yts
```

## Summarize the grouped data

It's grouped! Grouping doesn't change the data in any way, but how **functions operate on it**. Now we can summarize `Data_Value` (percent of respondents) by group:

```{r}
yts %>% summarize(avg_percent = mean(Data_Value, na.rm = TRUE))
```

## Using the `pipe` to connect these

Pipe `yts` into `group_by`, then pipe that into `summarize`:

```{r}
yts %>%
  group_by(Response) %>%
  summarize(avg_percent = mean(Data_Value, na.rm = TRUE),
            max_percent = max(Data_Value, na.rm = TRUE))
```

## Ungroup the data

The `ungroup` function will allow you to clear the groups from the data. You can also overwrite the first `group_by` with a new one.

```{r}
yts = ungroup(yts)
yts
```

## `group_by` with `mutate` - just add data

We can also use `mutate` to calculate the mean value for each year and add it as a column:

```{r}
yts %>%
  group_by(YEAR) %>%
  mutate(year_avg = mean(Data_Value, na.rm = TRUE)) %>%
  select(LocationDesc, Data_Value, year_avg)
```

## Counting

There are other functions, such as `n()` count the number of observations.

```{r}
yts %>%
  group_by(YEAR) %>%
  summarize(n = n(),
            mean = mean(Data_Value, na.rm = TRUE))
```

## Lab Part 2

[Website](http://jhudatascience.org/intro_to_R_class/index.html)

# Preview: plotting

## Basic Plots

Plotting is an important component of exploratory data analysis. These are some rough one-line plots that you can use in realtime while exploring your data. We will go over formatting and making plots look nicer in additional lectures.

* Basic summarization plots: 
    * `plot(x,y)`: scatterplot of x and y
    * `boxplot(y~x)`: boxplot of y against levels of x
    * `hist(x)`: histogram of x
    * `plot(density(x))`: kernel density plot of x

## Scatterplot

```{r}
plot( pull(jhu_cars,hp), pull(jhu_cars,mpg) ) # alt: plot(jhu_cars$hp, jhu_cars$mpg)
```

## Boxplot

```{r}
boxplot( pull(jhu_cars,hp) ~ pull(jhu_cars,cyl) )
```

## Histogram

```{r}
hist(pull(jhu_cars,mpg))
```

## Histogram

Use the `breaks =` argument to tweak the resolution:

```{r}
hist(pull(jhu_cars,mpg), breaks = 10)
```

## Density

```{r}
plot(density(pull(jhu_cars,mpg)))
```

## Lab Part 3

[Website](http://jhudatascience.org/intro_to_R_class/index.html)
