---
title: "Data_IO_Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

In this lab you can use the interactive console to explore or **Knit** the document. 
Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Ctrl-Enter (Windows/Linux) in an R code chunk.

# Part 1  

1. Read in the Youth Tobacco study from URL 
http://jhudatascience.org/intro_to_r/data/Youth_Tobacco_Survey_YTS_Data.csv
and assign it to an object named `youth`. 

```{r}
library(readr)
youth <- read_csv(file = "http://jhudatascience.org/intro_to_r/data/Youth_Tobacco_Survey_YTS_Data.csv")
```

2. Use `head()` to display the first few rows of the data frame. 

```{r}
head(youth)
```

3. Use `spec()` to display the list of column names and their type specification. 

```{r}
spec(youth)
```

4. Load the `readxl` package with the `library()` command. 

If it is not installed, install it via: `RStudio --> Tools --> Install Packages`. You can also try `install.packages("readxl")`.

```{r}
library(readxl)
```

5. Download the dataset of monuments from: http://jhudatascience.org/intro_to_r/data/Monuments.xlsx file to `Monuments.xlsx`.

```{r}
download.file(
  url = "http://jhudatascience.org/intro_to_r/data/Monuments.xlsx",
  destfile = "Monuments.xlsx",
  overwrite = TRUE,
  mode = "wb"
)
```

6. Use the `read_excel()` function in the `readxl` package to read the `Monuments.xlsx` file and call the output `mon`. 

```{r}
mon <- read_excel(path = "Monuments.xlsx")
```

Bonus: Run the following code - is there a problem? How do you know?
```{r}
read_delim(file = "http://jhudatascience.org/intro_to_r/data/Youth_Tobacco_Survey_YTS_Data.csv", delim = "\t")
# It should be a red flag to see that there is only one column that looks like: "YEAR,LocationAbbr,LocationDesc,TopicType,TopicDesc,MeasureDesc..."
# This file is comma delimited not tab delimited!
```

Bonus: Install and load the `haven` package. Look at the help page for `read_dta()` function, and scroll to the very bottom of the page. Try running some of the examples provided.

```{r eval = FALSE}
install.packages("haven")
library(haven)
?read_dta

path <- system.file("examples", "iris.dta", package = "haven")
read_dta(path)
```

# Part 2

7. Learn your working directory by running `getwd()`

```{r}
getwd()
```

8. Write out the `mon` object as a CSV file calling it "monuments.csv", using `write_csv()`. Where is the file now?

```{r}
write_csv(mon, file = "monuments.csv")
```


# Bonus

9. Write one of the variables from the Basic R lab to your working directory in `rds` format. Call the file `my_variable.rds`.

```{r}
y <- c(10, 20, 30, 40, 50, 60)
write_rds(y, file = "my_variable.rds")
```
