---
title: "Data Visualization Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1

0. Load the libraries

```{r, message=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(jhur)
```

0. Read Bike Lanes Dataset using `read_bike()` function from `jhur` package. Assign it to `bike` variable. 

Then, use the provided code to compute a data frame `bike_agg` with aggregate summary of bike lanes: average length of lanes (`lane_avg_length`) for each year (`dateInstalled`). 

```{r}
bike <- read_bike()

bike_agg <-
  bike %>%
  # filter data to keep only these observations for which year is non-0
  filter(dateInstalled != 0) %>%
  group_by(dateInstalled) %>%
  summarise(lane_avg_length = mean(length))

bike_agg
```

1. Use `ggplot2` package make plot of average length of lanes (`lane_avg_length`; y-axis) for each year (`dateInstalled`; x-axis). You can use lines layer (`+ geom_line()`) or points layer (`+ geom_point()`), or both!

Assign the plot to variable `my_plot`. Type `my_plot` in the console to have it displayed 

```{r}
## ggplot(???, aes(x = ???, y = ???)) +
# ??? +
# ???
```

2. "Update" your plot by adding a title and changing the x and y axis titles. 

```{r}



```

3. Use the `scale_x_continuous()` function and its arguments to modify x-axis look. Use `scale_y_continuous()` function and its arguments to modify the y-axis. 

```{r}
##  my_plot +
# scale_x_continuous(?????) +
# scale_y_continuous(????)  # force y-axis to start from 0, do not force upper bound
```


4. Observe several different versions of the plot by displaying `my_plot` while adding a different "theme" to it.

```{r}
# my_plot + theme_bw()
```


# Part 2

1. Use the provided code to compute a data frame `bike_agg_2` with aggregate summary of bike lanes: number of lanes (`lane_count`) -- separately for each year (`dateInstalled`) and for each lane type. 

```{r}
bike_agg_2 <-
  bike %>%
  filter(dateInstalled != 0) %>%
  group_by(dateInstalled, type) %>%
  summarise(lane_count = n())

bike_agg_2
```


2.  Use `ggplot2` package to make a plot showing trajectories of number of lanes (`lane_count`; y-axis) over year (`dateInstalled`; x-axis), where each bike line type has a different line (hint: use `group = type` in mapping) and a different color (hint: use `color = type` in mapping). 

```{r}




```

3. Redo the above plot by adding a faceting  (`+ facet_wrap(~ type, ncol = 3)`) to have data for each bike line type in a separate plot panel. 

(You may see `geom_path: Each group consists of only one observation. Do you need to adjust the group aesthetic?` warning as some bike lane types will have only 1 point plotted while trying to plot a line).

```{r}



```

Bonus: Modify `facet_plot` to remove the legend and change the names of the axis titles to be "Number of bike lanes" for the x axis and "Date bike lane was installed" for the y axis.  

```{r}

```
