---
title: "Variables: Objects in R"
author: "Introduction to R for Public Health Researchers"
output:
  ioslides_presentation:
    css: ../styles.css
    widescreen: yes
subtitle: Basic R Functionality
---

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(comment = "")
```

## Common new users frustations

1.  **Different versions of software**
1.  **Data type problems (is that a string or a number?)**
2.  Working directory problems: trying to read files that R "can't find"
    - RStudio can help, and so do RStudio Projects
    - discuss in Data Input/Output lecture
3.  Typos (R is **case sensitive**, `x` and `X` are different)
    - RStudio helps with "tab completion"
    - discussed throughout

<!-- ## Getting Started -->

<!-- * You should have the latest version of R installed! -->
<!-- * Open RStudio -->
<!-- * Files → New → R Script →
<!-- * Save the blank R script in a directory of your choosing -->
<!-- * Add a comment header -->

<!-- ## Commenting in Scripts -->

<!-- Add a comment header to day1.R :`#` is the comment symbol -->

<!-- ```{r} -->
<!-- ################# -->
<!-- # Title: Demo R Script -->
<!-- # Author: John Muschelli -->
<!-- # Purpose: Demonstrate comments in R -->
<!-- ################### -->

<!-- # nothing to its right is evaluated -->

<!-- # this # is still a comment -->
<!-- ### you can use many #'s as you want -->

<!-- # sometimes you have a really long comment, -->
<!-- #    like explaining what you are doing  -->
<!-- #    for a step in analysis.  -->
<!-- # Take it to another line -->
<!-- ``` -->

## Explaining output on slides

In slides, a command (we'll also call them code or a code chunk) will look like this
```{r code}
print("I'm code")
```

And then directly after it, will be the output of the code.  
So `print("I'm code")` is the code chunk and [1] "I'm code" is the output.

## R as a calculator

```{r calcDemo}
2 + 2
2 * 4
2 ^ 3
```

Note, when you type your command, R inherently thinks you want to print the result. 

## R as a calculator

* The R console is a full calculator
* Try to play around with it:
    * +, -, /, * are add, subtract, divide and multiply
    * ^ or ** is power
    * parentheses -- ( and ) -- work with order of operations 
    * %% finds the remainder

## R as a calculator

```{r calcDemo2}
2 + (2 * 3)^2
(1 + 3) / 2 + 45
6 / 2 * (1 + 2)
```

## R as a calculator

Try evaluating the following:

* `2 + 2 * 3 / 4 -3`
* `2 * 3 / 4 * 2`
* `2^4 - 1`

## Commenting in Scripts

`#` is the comment symbol

```{r}
# this is a comment

# nothing to its right is evaluated

# this # is still a comment
### you can use many #'s as you want

1 + 2 # Can be the right of code

```


## R variables 

* You can create variables from within the R environment and from files on your computer
* R uses "=" or "<-" to assign values to a variable name
* Variable names are case-sensitive, i.e. X and x are different

```{r assign}
x = 2 # Same as: x <- 2
x
x * 4
x + 2
```

## R variables

* The most comfortable and familiar class/data type for many of you will be `data.frame`
* You can think of these as essentially Excel spreadsheets with rows (usually subjects or observations) and columns (usually variables)

## R variables

* `data.frame`s are somewhat advanced objects in R; we will start with simpler objects;
* Here we introduce "1 dimensional" classes; often referred to as 'vectors'
* Vectors can have multiple sets of observations, but each observation has to be the same class.

```{r assignClass}
class(x)
y = "hello world!"
print(y)
class(y)
```

## R variables

Try assigning your full name to an R variable called `name`

## R variables

Try assigning your full name to an R variable called `name`

```{r myName}
name = "Ava Hoffman"
name
```

## The 'combine' function

The function `c()` collects/combines/joins single R objects into a vector of R objects. It is mostly used for creating vectors of numbers, character strings, and other data types. 

```{r assign3a}
x <- c(1, 4, 6, 8)
x
class(x)
```

## The 'combine' function

Try assigning your first and last name as 2 separate character strings into a single vector called `name2`

## The 'combine' function

Try assigning your first and last name as 2 separate character strings into a length-2 vector called `name2`

```{r myName2}
name2 = c("Ava","Hoffman")
name2
```

## R variables

`length()`: Get or set the length of vectors (including lists) and factors, and of any other R object for which a method has been defined.

```{r assign3b}
length(x)
y
length(y)
```

##  R variables

What do you expect for the length of the `name` variable? What about the `name2` variable?

What are the lengths of each?

##  R variables

What do you expect for the length of the `name` variable? What about the `name2` variable?

What are the lengths of each?

```{r myName3}
length(name)
length(name2)
```

## R variables

You can perform functions to entire vectors of numbers very easily.

```{r assign4}
x + 2
x * 3
x + c(1, 2, 3, 4)
```

## Lab Part 1

[Website](http://jhudatascience.org/intro_to_r/index.html)

## R variables

But things like algebra can only be performed on numbers.

```{r, error=TRUE}
name2 + 4
```

## R variables

Save these modified vectors as a new vector.

```{r assign5}
y = x + c(1, 2, 3, 4)
y 
```

Note that the R object `y` is no longer "Hello  World!" - It has effectively been overwritten by assigning new data to the variable

## R variables

You can get more attributes than just class.  The function `str` gives you the structure of the object.

```{r assign2}
str(x)
str(y)
```
This tells you that `x` is a numeric vector and tells you the length.

## Review

* Creating a new script
* Using R as a calculator
* Assigning values to variables
* Performing algebra on numeric variables


## Lab Part 2

[Website](http://jhudatascience.org/intro_to_r/index.html)
