---
title: "Subsetting Data in R"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this lab you can use the interactive console to explore but please record your commands here.  Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Ctrl-Enter (Windows/Linux) (But only inside the code chunks designated with the ```{r}``` areas).

# Part 1

Now we will work with the `county_pop` data from the same case study.


```{r, message = FALSE}
# don't forget to load the packages that you will need!
library(tidyverse)
```


Import the data from "https://jhudatascience.org/intro_to_r/data/county_pop.csv" using `read_csv` and assign the data to an object called `county_pop`. 
Check that it worked by seeing if you have the `county_pop` data.

```{r 0response}

```

### 1.1

What class is `county_pop`?

```{r 1.1response}

```

### 1.2

How many observations (rows) and variables (columns) are in the dataset - try the `dim()` function?

```{r 1.2response}

```

### 1.3

Next, rename the column `BUYER_STATE`  to be `State_name` (hint - use `rename()` and watch out for the order of the new and old names!). 

```{r 1.3response}

```

### 1.4

Convert the column names of `county_pop` to be all lower case. Use `rename_with()`, and the `tolower` command.

```{r 1.4response}

```


# Practice on Your Own!

### P.1

How can you show the first 3 rows and the last 3 rows of `county_pop` (in two lines of code)?

```{r P.1response}

```


# Part 2

### 2.1

Create a subset of the `county_pop` that only contains the columns: `county_name`, `year`, and `population` and assign this object to `pop_sub` - what are the dimensions of this dataset?

```{r 2.1response}

```

### 2.2

Start with `county_pop` again instead of the dataset you just made. Subset the data to only include the `population` column and the columns that end with "_name". Hint: use  `select()` and `ends_with()`. Assign this subset of the data to be `pop2`. Again take a look at the data and check the dimensions.

```{r 2.2response}

```

### 2.3

Pull the variable `county_name` from `pop2`. How does this differ form selecting it? Use head() to take a look at both options.

```{r 2.3response}

```

### 2.4

Subset the rows of `county_pop` that have **more** than 100000 for population - how many rows are there? Use `filter()`.

```{r 2.4response}

```

### 2.5

Subset the rows of `county_pop` that have a population of **more** than 100000 and have a year value **less** than 2010 - how many are there?

```{r 2.5response}

```

### 2.6

Subset the rows of `county_pop` that have a population of **more** than 100000 and a year value of **less than or equal to ** 2010 - how many are there?

```{r 2.6response}

```

### 2.7

Why do the answers for 2.5 and 2.6 differ?

```{r 2.7response}

```


# Practice on Your Own!

### P.2

Subset the rows of `county_pop` for rows that have `state_name` of `NM`,  **or** **less** than 10000 `population`.
How many rows have both?

```{r P.2response}

```

### P.3

Select the variables that contain the letter "a" from `cars_mt_copy`.

```{r P.3response}

```


# Part 3

### 3.1

Create a subset called `county_2012` from `county_pop` that only contains the rows for the year 2012 and only the columns: `county_name`,	`state_name`, `population`. The column `year` should not be included in `county_2012`.
	What are the dimensions of this dataset? Don't use pipes (`%>%`) and instead do this in two steps creating the `county_2012` object with `filter` and updating it with `select`.

```{r 3.1response}

```

### 3.2

Try the same thing again with pipes `%>%`.

```{r 3.2response}

```

What happens if you do the steps in a different order? Why does this not work?

```{r, error = TRUE, label = '3.2aresponse'}

```

### 3.3

Use `arrange()` to order the rows of `county_2012` by `population` in increasing order. What is county with the smallest population?

```{r 3.3response}

```

### 3.4

Create a new variable in `county_2012` called `pop1000`, which  is equal to `population` divided by 1000, using `mutate()`(don't forget to reassign `county_2012`). Use pipes `%>%`.

```
# General format
NEWDATA <- OLD_DATA %>% mutate(NEW_COLUMN = OLD_COLUMN)
```

```{r 3.4response}

```


# Practice on Your Own!

### P.4

Move the `state_name` column to be before `county_name` in the `county_2012` dataset. Use `relocate()`.

```{r P.4response}

```

### P.5

How can you find the value of `population` for Chambers, AL without just looking at the data manually and instead use functions we learned today?

```{r P.5response}

```
