---
title: "Subsetting Data in R - Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this lab you can use the interactive console to explore but please record your commands here.  Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Ctrl-Enter (Windows/Linux) (But only inside the code chunks designated with the ```{r}``` areas).

```{r, message = FALSE}
# don't forget to load the packages that you will need!!
library(dplyr)
library(tidyverse)
library(jhur)
```

# Part 1

First check to see if you have the `mtcars` dataset by running this chunk.Remember that code is denoted with backticks "`" around it in the main text, but do not copy backticks when you paste code elements into code chunks (like the gray section below). 

```{r}
mtcars
```


1. What class is `mtcars`?

```{r}

```

2. How many observations (rows) and variables (columns) are in the `mtcars` dataset - try the `dim()` function?

```{r}

```

3. Copy mtcars into an object called `cars_mt_copy` using the code supplied in the chunk below. Next, rename the column `mpg` in `cars_mt_copy` to `MPG` (hint - use `rename()`).

```{r}
cars_mt_copy <- mtcars

```

4. Convert the column names of `cars_mt_copy` to all upper case. Use `rename_with()`, and the `toupper` command.

```{r}

```

Bonus: How can you print the first 3 rows and the last 3 rows of `cars_mt_copy` (in two lines of code)?

```{r}

```

# Part 2

1. Using `cars_mt_copy` create a column called `CAR` using the `rownames_to_column()` function. Create a tibble from the data called `cars_mt_copy`.  Take a look at your data to check it. How many dimensions does it have?

```{r}

```

2. Subset both the column `CAR` that you just created and the columns from `cars_mt_copy` that end in `"P"` and call it `pvars`. Hint: use `ends_with()` and `select()`. Again take a look at the data.

```{r}

```

3. Create a subset of the `cars_mt_copy` that only contains the columns: `CAR`, `WT`, `QSEC`, and `HP` and assign this object to `carsSub` - what are the dimensions of this dataset?

```{r}


```

Bonus: Convert the column names of `carsSub` to all lower case. Take a look at the data.

```{r}


```

Bonus: Pull the variable `car` from `carsSub`. How does this differ form selecting it?

```{r}

```

Bonus: Select the variables that contain the letter a from `cars_mt_copy`.

```{r}

```

# Part 3

1. Subset the rows of `cars_mt_copy` that get **more** than 20 miles per gallon (`MPG`) of fuel efficiency - how many are there? Use `filter()`.

```{r}

```



2. Subset the rows of `cars_mt_copy` that get **less** than 16 miles per gallon (`MPG`) of fuel efficiency and **more** than 150 horsepower (`HP`) - how many are there?

```{r}

```


3. Subset the rows of `cars_mt_copy` that get **less** than 16 miles per gallon (`MPG`) of fuel efficiency **and** **greater than or equal** to 150 horsepower (`HP`) - how many are there?

```{r}

```

4. Why do the answers for 2 and 3 differ?


Bonus:

Subset the rows of `cars_mt_copy` for cars that have 4 `CYL`,  **or** **less** than 1.9 WT.
How many cars have both?




# Part 4

1. Create a subset called `carsSub` from the `cars_mt_copy` data that only contains the columns:
`CAR`,	`WT`, `QSEC`, and `HP` for only the cars with 8 cylinders (`CYL`).
	What are the dimensions of this dataset? Don't use pipes (`%>%`).

```{r}

```

2. Try the same thing again with pipes `%>%`.

```{r}

```

What happens if you do the steps in a different order? Why?


3. Re-order the rows of `carsSub` by weight (`WT`) in increasing order. Use `arrange()`. What is the heaviest car?

```{r}

```


4. Create a new variable in `carsSub` called `WT2`, which  is equal to `WT^2`, using `mutate()`(don't forget to reassign `carsSub`). Use pipes `%>%`:

```{r}

```


Bonus: Move the `WT2` column to be first. Use `relocate()`.

```{r}

```


BONUS: How can you find the value of `WT2` for the Cadillac Fleetwood without just looking at the data manually and instead use functions we learned today?

```{r}


```

