---
title: "Functions that need pull first"
output: 
  html_document:
    toc: TRUE
---

When can you use a function directly on a column of a tibble and when do you need a vector in R?

There are some inconsistencies across different functions. Typically if it is a base R function, you often need to pull the data out into a vector first.  Sometimes you just need to try them out. Always check that your work did what you expect!

## In general

In general these are the conventions for working with values (recall that values or samples within rows):

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(kableExtra)
library(tidyverse)
text_tbl <- data.frame(
  Type = c("base R and tidyverse function", "special tidyverse functions", "base R functions", "stat functions that come with R"),
  Example = c("setdiff()", "stringr functions like str_replace()", "mean()", "cor.test()"))
kbl(text_tbl) %>%
  # kable_classic(full_width = F, html_font = "helvetica") %>%
  kable_paper(full_width = F, html_font = "helvetica", font_size = "18") %>%
  column_spec(1:2, bold = T, border_right = T, border_left = TRUE) %>%
  row_spec(0, align = "c") %>%
  row_spec(1, background = "yellow") %>%
  row_spec(2, background = "lightgreen") %>%
  row_spec(3:4, background = "lightblue")
```

## `setdiff()`

This function was originally a base R function and it got adopted by a tidyverse package and is therefore used as it was previously and requires vectors.

```{r}
animal1 <-tibble(animals = c("cat", "dog", "mouse"), 
                 size  = c("medium", "medium", "small"))
animal2 <-tibble(animals = c("cat", "hourse", "fish"),
                 size  = c("medium", "large", "small"))

# shows what is in the first data set that is not in the second dataset as a tibble
dplyr::setdiff(select(animal1, animals), 
               select(animal2, animals))
# shows what is in the first data set that is not in the second dataset as a vector
dplyr::setdiff(pull(animal1, animals), 
               pull(animal2, animals))
# shows what is in the first data set that is not in the second dataset as a vector
base::setdiff(animal1$animals, 
               animal2$animals)
# shows what is in the first data set that is not in the second dataset as a vector
base::setdiff(pull(animal1,animals), 
              pull(animal2,animals))

base::setdiff(select(animal1, animals), 
              select(animal2, animals)) # works a bit differently

lubridate::setdiff(select(animal1, animals), 
              select(animal2, animals)) 
```


