---
output:
  html_document
---

## Online Videos of Lectures
***

### From Summer Institute 2021

```{r, echo = FALSE, message = FALSE, results='asis'}
mat = matrix(c(
  "RStudio", "https://youtu.be/zpAQrglIJb0",
  "Basic R", "https://youtu.be/md30mwFJ2_Q",
  "Data IO", "https://youtu.be/6xNyETqIqZU",
  "Subsetting Data", "https://youtu.be/YeBSc2YXr4U",
  "Data Summarization Part 1 + 2", "https://youtu.be/yL3BGDWtVC4",
  "Data Classes", "https://youtu.be/zBCvbikMTAc",
  "Data Cleaning", "https://youtu.be/TaREvr5evwk",
  "Data Manipulation", "https://youtu.be/-039V99I-PE",
  "Reproducibility", "https://youtu.be/ui9dJLqtdCs",
  "Statistics", "https://youtu.be/EJwlBoBHoz4",
  "Data Visualization", "https://youtu.be/ngQX9khx7UQ",
  "Functions", "https://youtu.be/epIujSKgNi4"
  ), ncol = 2, byrow = TRUE)
mat = data.frame(mat, stringsAsFactors = FALSE)
colnames(mat) = c("Day", "Link to Video")
knitr::kable(mat, format = "html") %>% 
  kable_styling() 
```

### From Winter Institute 2020

```{r, echo = FALSE, message = FALSE, results='asis'}
mat = matrix(c(
  "RStudio and Data Classes", "https://youtu.be/vyIsDnsq5jY",
  "Subsetting Data", "https://youtu.be/mT8lSagYbjM",
  "Data Summarization Part 1", "https://youtu.be/SZYpzt9zy0g",
  "Data Classes", "https://youtu.be/82zSA1N0mnA",
  "Data Cleaning", "https://youtu.be/G3V2YPaQN34",
  "Data Manipulation", "https://youtu.be/43MPdw5bf4o",
  "Statistics 1", "https://youtu.be/Jr4ljyzrr4U",
  "Statistics 2", "https://youtu.be/ub2BSbK9lMM"
  ), ncol = 2, byrow = TRUE)
mat = data.frame(mat, stringsAsFactors = FALSE)
colnames(mat) = c("Day", "Link to Video")
knitr::kable(mat, format = "html") %>% 
  kable_styling() 
```

### From Summer Institute 1 2017

```{r videos, echo = FALSE, message = FALSE, results='asis'}
mat = matrix(c(
  "Day 1", "https://youtu.be/Xi-wsACc7p0",
"Day 2", "https://youtu.be/u1FQt9Hr8iw",
"Day 3", "https://youtu.be/woc7AGJTzZw",
"Day 4", "https://youtu.be/RZ7eVIMPIes",
"Day 5", "https://youtu.be/e8cFV8wluC0"), ncol = 2, byrow = TRUE)
mat = data.frame(mat, stringsAsFactors = FALSE)
colnames(mat) = c("Day", "Link to Video")
knitr::kable(mat, format = "html") %>% 
  kable_styling() 
```

<br>

## Additional Resources
***

**Need help?** 

- [Various "Cheat Sheets"](https://www.rstudio.com/resources/cheatsheets/)   

- [R reference card](http://cran.r-project.org/doc/contrib/Short-refcard.pdf) 

- [R jargon](https://link.springer.com/content/pdf/bbm%3A978-1-4419-1318-0%2F1.pdf)  

- [R terminology](https://cran.r-project.org/doc/manuals/r-release/R-lang.pdf)

- [What is the Tidyverse?](https://www.tidyverse.org/packages/)

-  Animations of join functions:
    [`full-join()`](https://github.com/gadenbuie/tidyexplain/blob/master/images/full-join.gif)
    [`inner_join()`](https://github.com/gadenbuie/tidyexplain/blob/master/images/inner-join.gif)
    [`left-join()`](https://github.com/gadenbuie/tidyexplain/blob/master/images/left-join.gif)
    [`right-join()`](https://github.com/gadenbuie/tidyexplain/blob/master/images/right-join.gif)

- PC users who want to see how to move files around (especially from downloads), check out this video:   https://youtu.be/we6vwB7DsNU

- Mac users who want to see how to move files around (especially from downloads), check out this video:   https://www.youtube.com/watch?v=Ao9e0cDzMrE

**Want more?**  

- [Tidyverse Skills for Data Science Book](https://jhudatascience.org/tidyversecourse/)  
(more about the tidyverse, some modeling, and machine learning)

- [Tidyverse Skills for Data Science Course](https://www.coursera.org/specializations/tidyverse-data-science-r)  
(same content with quizzes, can get certificate with $)

- [R for Data Science](http://r4ds.had.co.nz/)  
(great general information)
 
- [Open Case Studies](https://www.opencasestudies.org/)  
(resource for specific public health cases with statistical implementation and interpretation)

- [Dataquest](https://www.dataquest.io/)  
(general interactive resource)

- [Quick R Guide]( http://statmethods.net/)  
(nice free general resource)

- [Building up a `ggplot2` figure](https://hopstat.wordpress.com/2016/02/18/how-i-build-up-a-ggplot2-figure/)  
(guide to making plots)

**Interested in Reproducibility?**

Check out Candace's courses:

- [Introduction](https://jhudatascience.org/Reproducibility_in_Cancer_Informatics/)

- [Advanced](https://jhudatascience.org/Adv_Reproducibility_in_Cancer_Informatics/)

**R for Stata, SPSS, and SAS files**

- The [Haven](https://haven.tidyverse.org/) package  
 (This package is super useful for reading and writing files so that they are compatible across Stata, SPSS, SAS, and R) 
- [R vs Stata](https://link.springer.com/content/pdf/bbm%3A978-1-4419-1318-0%2F1.pdf)  
(See page 505)
