---
title: "Day 1"
output:
  html_document:
    fig_caption: false
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
source("../../scripts/utils.R")
schedule <- read_markdown("../_schedule_table.Rmd")
```

## Objectives

Upon completion of this session, you will be able to do the following:

1. Begin interfacing with the RStudio IDE.
2. Locate basic help tools within the RStudio IDE.
3. Learn how to open and create RMarkdown files (.Rmd).
4. Understand R as a calculator.


## Lectures

Lecture recording links will be emailed directly to students.


## Material

```{r, results='asis', echo = FALSE, message=FALSE, warning=FALSE}
# The str_replace_all creates an absolute url to linked documents. Important because there are word and pdf rendered versions of the syllabus.
# The filter() step removes all dates but this date to make a smaller table.
pander::pandoc.table(
  schedule %>%
    urls_for_module_details() %>%
    filter(
      row_number() >= which(str_detect(schedule$`**Day**`, "\\[Day 1\\]")) &
        row_number() < which(str_detect(schedule$`**Day**`, "\\[Day 2\\]"))
    ),
  missing = "",
  split.tables = Inf,
  style = "rmarkdown"
)
``` 


## Homework

`r emo::ji("rotating_light")` **SURVEYS**: [Best times for Office Hours and first impressions feedback](`r config::get("courseplus_surveys")`)

`r emo::ji("pencil")` **HOMEWORK 1**: [Dataquest](https://app.dataquest.io/login?target-url=%2Fm%2F499%2Fintroduction-to-programming-in-r) 


## Drop Boxes

[Assignment Drop Boxes](`r config::get("courseplus_dropboxes")`)

**Note**: only people taking the course for credit must turn in the assignments. However, we will evaluate all submitted assignments in case others would like feedback on their work.


## Library

[RStudio cheatsheets](https://www.rstudio.com/resources/cheatsheets/)

[RStudio IDE cheatsheet (PDF)](../../resources/rstudio_IDE_cheatsheet.pdf)

[R jargon](https://link.springer.com/content/pdf/bbm%3A978-1-4419-1318-0%2F1.pdf)  

[R terminology](https://cran.r-project.org/doc/manuals/r-release/R-lang.pdf)

