---
output:
  html_document:
    fig_caption: false
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
source("../../scripts/utils.R")
schedule <- read_markdown("../_schedule_table.Rmd")
```

# Day 2

## Objectives

Upon completion of this session, you will be able to do the following:

1. Learn how to open and create RMarkdown files (.Rmd).
2. Understand R as a calculator.
3. Import common data types like csv, xls/xlsx, and txt files into RStudio.


## Lectures

Lecture recording links will be emailed directly to students.


## Material

```{r, results='asis', echo = FALSE, message=FALSE, warning=FALSE}
# The str_replace_all creates an absolute url to linked documents. Important because there are word and pdf rendered versions of the syllabus.
# The filter() step removes all dates but this date to make a smaller table.
pander::pandoc.table(
  schedule %>%
    urls_for_module_details() %>% 
    filter(
      row_number() >= which(str_detect(schedule$`**Day**`, "\\[Day 2\\]")) &
        row_number() < which(str_detect(schedule$`**Day**`, "\\[Day 3\\]"))
    ),
  missing = "",
  split.tables = Inf,
  style = "rmarkdown"
)
``` 


## Homework

`r emo::ji("rotating_light")` **SURVEY**: [Best times for Office Hours](https://courseplus.jhu.edu/core/index.cfm/go/survey2:student.home/coid/18000/#!/surveyid/36566/surveyview)

`r emo::ji("rotating_light")` **SURVEY**: [Day 2 check-in](https://courseplus.jhu.edu/core/index.cfm/go/survey2:student.home/coid/18000/#!/surveyid/36570/surveyview) 

`r emo::ji("pencil")` **HOMEWORK 1**: [Dataquest](https://app.dataquest.io/login?target-url=%2Fm%2F499%2Fintroduction-to-programming-in-r) 


## Drop Boxes

[Assignment Drop Boxes](https://courseplus.jhu.edu/core/index.cfm/go/db:main/coid/18000/)

**Note**: only people taking the course for credit must turn in the assignments. However, we will evaluate all submitted assignments in case others would like feedback on their work.


## Library

[RStudio cheatsheets](https://www.rstudio.com/resources/cheatsheets/)

[RStudio IDE cheatsheet (PDF)](../../resources/rstudio_IDE_cheatsheet.pdf)
