---
title: "Day 4"
output:
  html_document:
    fig_caption: false
---

```{r setup, echo = FALSE, message = FALSE, warning=FALSE}
library(tidyverse)
source("../../scripts/utils.R")
schedule <- read_markdown("../_schedule_table.Rmd")
```

## Objectives

Upon completion of this session, you will be able to do the following:

1. Evaluate data integrity with object summarization functions.
2. Compute simple statistical summaries on a data set.
3. Clearly express a sequence of multiple operations using pipes.
4. Selectively perform operations on groups of interest from the data and visualize basic statistical outputs from those groups. 
5. Learn the data classes used in R programming and explore how to extract, encode, and manipulate the data types. 

## Lectures

Lecture recording links will be emailed directly to students.


## Material

```{r, results='asis', echo = FALSE, message=FALSE, warning=FALSE}
# The str_replace_all creates an absolute url to linked documents. Important because there are word and pdf rendered versions of the syllabus.
# The filter() step removes all dates but this date to make a smaller table.
pander::pandoc.table(
  schedule %>%
    urls_for_module_details() %>%
    filter(
      row_number() >= which(str_detect(schedule$`**Day**`, "\\[Day 4\\]")) &
        row_number() < which(str_detect(schedule$`**Day**`, "\\[Day 5\\]"))
    ),
  missing = "",
  split.tables = Inf,
  style = "rmarkdown"
)
``` 


## Homework

`r emo::ji("rotating_light")` **SURVEY**: [Day 4 check-in](`r config::get("courseplus_surveys")`)

`r emo::ji("pencil")` **HOMEWORK 1**: [Dataquest](https://app.dataquest.io/login?target-url=%2Fm%2F499%2Fintroduction-to-programming-in-r) 

`r emo::ji("pencil")` **HOMEWORK 2**: [Questions (Rmd)](../../modules/HW/homework2.Rmd), [Questions (HTML)](../../modules/HW/homework2.html)


## Drop Boxes

[Assignment Drop Boxes](`r config::get("courseplus_dropboxes")`)

**Note**: only people taking the course for credit must turn in the assignments. However, we will evaluate all submitted assignments in case others would like feedback on their work.


## Library

[RStudio cheatsheets](https://www.rstudio.com/resources/cheatsheets/)

[RStudio IDE cheatsheet (PDF)](../../resources/rstudio_IDE_cheatsheet.pdf)
