---
title: "Manipulating Data in R Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data in this lab comes from the CDC (https://covid.cdc.gov/covid-data-tracker/#vaccinations_vacc-total-admin-rate-total - snapshot from January 12, 2022) and the Bureau of Economic Analysis (https://www.bea.gov/data/income-saving/personal-income-by-state).

```{r message=FALSE}
library(readr)
library(dplyr)
library(tidyr)
```

# Part 1

1. Read in the SARS-CoV-2 Vaccination data from http://jhudatascience.org/intro_to_r/data/USA_covid19_vaccinations.csv. You can use the url or download the data. Assign the data the name "vacc". We will be reviewing new concepts here and incorporating some from week 1. Remember to use `read_csv()` from the `readr` package.

```{r}

```

2. Look at the column names using `colnames` - do you notice any patterns?

```{r}

```

3. Let's rename the column "State/Territory/Federal Entity" in "vacc" to "Entity" using `rename`. Make sure to reassign to `vacc` here and in subsequent steps.

```
Usage: rename(newname = oldname)
```

```{r}

```

4. Select only the columns "Entity", and those that start with "Percent". Use `select` and `starts_with("Percent")`. 

```
Usage: select(colname1, colname2, ...)
```

```{r}

```

5. Create a new dataset "vacc_long" that does `pivot_longer()` on all columns except "Entity". Remember that `!Entity` means all columns except "Entity".

```
Usage: pivot_longer(cols = colname(s))
```

```{r}

```

6. Filter the "Entity" column so it only includes values in the following list: "Maryland","Virginia","Florida","Massachusetts", "United States". **Hint**: use `filter` and `%in%`.

```
Usage: filter(colname %in% c(1, 2, 3, ...))
```

```{r}

```

7. Use `pivot_wider` to reshape "vacc_long". Use "Entity" for the `names_from` argument. Use "value" for the `values_from` argument. Call this new data `vacc_wide`. Look at the data. How do these states compare to one another.

```
Usage: pivot_wider(names_from = ???, values_from = ???)
```

```{r}

```

**Bonus / optional**

Combine blocks 1 and 3-7 by chaining everything together using the pipe ` %>% `. Call your data `vacc_compare`.

```{r}

```

Tweak the code in the block above to:

- look for columns that start with "Total" (instead of "Percent") and 
- select different states/Entities to compare

```{r}

```


# Part 2

8. Read in the GDP and Personal Income Data from http://jhudatascience.org/intro_to_r/data/gdp_personal_income.csv. You can use the url or download the data. Call it "gdp". 

```{r}

```

9. Use `pivot_wider` to reshape "gdp". Use "Description" for the `names_from` argument. Use "`2020`" for the `values_from` argument. Reassign this data to "gdp".

```
Usage: pivot_wider(names_from = ???, values_from = ???)
```

```{r}

```

10. Join the data so that you only retain data in both "vacc" and "gdp". 

- First, try merging without using the `by` argument - what happens? 
- Next, try merging using `by = c("Entity" = "GeoName")`. 
- Call the output "merged". How many observations are there?

```
Usage: inner_join(x, y, by = c("colname1" = "colname2"))
```

```{r}

```

11. Join the data using a `full_join.` Call the output "full". How many observations are there?

```
Usage: full_join(x, y, by = c("colname1" = "colname2"))
```

```{r}

```

**Bonus / optional**

Do a left join of "vacc" and "gdp". How many observations are there?

```{r}

```

Repeat above with a `right_join` with the same order of the arguments. How many observations are there?

```{r}

```

Perform a `setdiff` on "vacc" and "gdp" to determine what Entities are missing from the GDP data and which are missing from the vaccine data. Remember you need to `pull()` the columns you want to compare.

```{r}

```
