---
title: "Manipulating Data in R Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data in this lab comes from the CDC (https://covid.cdc.gov/covid-data-tracker/#vaccinations_vacc-total-admin-rate-total - current as of January 12, 2022!) and the Bureau of Economic Analysis (https://www.bea.gov/data/income-saving/personal-income-by-state).

```{r message=FALSE}
library(readr)
library(dplyr)
library(tidyr)
```

# Part 1

1. Read in the SARS-CoV-2 Vaccination data from http://jhudatascience.org/intro_to_r/data/USA_covid19_vaccinations.csv. Assign the data the name "vacc". We will be reviewing new concepts here and incorporating some from week 1.

```{r}
vacc <- read_csv("http://jhudatascience.org/intro_to_r/data/USA_covid19_vaccinations.csv")
# If downloaded
# vacc <- read_csv("USA_covid19_vaccinations.csv")
```

2. Look at the column names using `colnames` - do you notice any patterns?

```{r}
colnames(vacc)
# Looks like many start with "Percent" and some start with "Total" - this indicates there are different units of measure for these different variables!
```

3. Let's rename the column "State/Territory/Federal Entity" in "vacc" to "Entity" using `rename`.

```{r}
vacc <- vacc %>% rename(Entity = `State/Territory/Federal Entity`)
```

4. Create a new dataset "vacc_long" that does `pivot_longer` on all columns that start with the word "Percent". Use `starts_with("Percent")` for the `cols` argument. Use "Measure" for the `names_to` argument. Use "Percent" for the `values_to` argument.

```{r}
# ??? <- ??? %>% pivot_longer(cols = ???,
#                             names_to = ???,
#                             values_to = ???)
vacc_long <- vacc %>% pivot_longer(
  cols = starts_with("Percent"),
  names_to = "Measure",
  values_to = "Percent"
)
```

5. Select only the columns "Entity", "Measure", and "Percent" from "vacc_long". Use `select`. Filter "Entity" so it only includes values in the following list: "Maryland","Virginia","Florida","Massachusetts", "United States". **Hint**: use `filter` and `%in%`.

```{r}
vacc_long <- vacc_long %>%
  select(Entity, Measure, Percent) %>%
  filter(Entity %in% c("Maryland", "Virginia", "Florida", "Massachusetts", "United States"))
```

6. Use `pivot_wider` to reshape "vacc_long". Use "Entity" for the `names_from` argument. Use "Percent" for the `values_from` argument.

```{r}
# ??? <- ??? %>%
#   pivot_wider(names_from = ???,
#               values_from = ???)
vacc_wide <- vacc_long %>%
  pivot_wider(
    names_from = Entity,
    values_from = Percent
  )
```

# Part 2

7. Read in the GDP and Personal Income Data from http://jhudatascience.org/intro_to_r/data/gdp_personal_income.csv. Call it "gdp". 

```{r}
gdp <- read_csv("http://jhudatascience.org/intro_to_r/data/gdp_personal_income.csv")
# If downloaded
# gdp <- read_csv("gdp_personal_income.csv")
```

8. Use `pivot_wider` to reshape "gdp". Use "Description" for the `names_from` argument. Use "`2020`" for the `values_from` argument. Reassign this data to "gdp".

```{r}
gdp <- gdp %>%
  pivot_wider(
    names_from = Description,
    values_from = `2020`
  )
```

9. Join the data so that you only retain data in both "vacc" and "gdp". First, try merging without using the `by` argument - what happens? Next, try merging using `by = c("Entity" = "GeoName")`. Call the output "merged". How many observations are there?

```{r}
# merged <- inner_join(vacc, gdp) does not work!
merged <- inner_join(vacc, gdp, by = c("Entity" = "GeoName"))
nrow(merged)
```

10. Join the data using a `full_join.` Call the output "full". How many observations are there?

```{r}
full <- full_join(vacc, gdp, by = c("Entity" = "GeoName"))
nrow(full)
```

11. Do a left join of "vacc" and "gdp". How many observations are there?

```{r}
left <- left_join(vacc, gdp, by = c("Entity" = "GeoName"))
nrow(left)
```

12. Repeat above with a `right_join` with the same order of the arguments. How many observations are there?

```{r}
right <- right_join(vacc, gdp, by = c("Entity" = "GeoName"))
nrow(right)
```

13. Perform a `setdiff` on "vacc" and "gdp" to determine what Entities are missing from the GDP data and which are missing from the vaccine data.

```{r}
setdiff(vacc %>% pull(Entity), gdp %>% pull(GeoName))
setdiff(gdp %>% pull(GeoName), vacc %>% pull(Entity))
```

**Bonus**

14. Repeat #4 - #6 but instead look for columns that start with "Total" and select different states/Entities to compare. Also try combining all functions in one long block using the pipe `%>%`.

```{r}
vacc_wide <-
  vacc %>%
  pivot_longer(starts_with("Total"), names_to = "Measure", values_to = "Total") %>%
  select(Entity, Measure, Total) %>%
  filter(Entity %in% c("Alaska", "Kansas", "California", "United States")) %>%
  pivot_wider(names_from = Entity, values_from = Total)
```
