---
title: "Manipulating Data in R Lab - Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data in this lab comes from the CDC (https://covid.cdc.gov/covid-data-tracker/#vaccinations_vacc-total-admin-rate-total - snapshot from January 12, 2022) and the Bureau of Economic Analysis (https://www.bea.gov/data/income-saving/personal-income-by-state).

```{r message=FALSE}
library(tidyverse)
library(janitor)
```

# Part 1

### 1.1

Read in the SARS-CoV-2 Vaccination data from http://jhudatascience.org/intro_to_r/data/USA_covid19_vaccinations.csv. You can use the url or download the data. Assign the data the name "vacc". We will be reviewing new concepts here and incorporating some from week 1. 

-  Remember to use `read_csv()` from the `readr` package.
-  Do NOT use `read.csv()`.

```{r 1.1response}
vacc <- read_csv("http://jhudatascience.org/intro_to_r/data/USA_covid19_vaccinations.csv")
# If downloaded
# vacc <- read_csv("USA_covid19_vaccinations.csv")
```

### 1.2

First, clean up the column names with the `clean_names()` function in the `janitor` package. Then, look at the column names using `colnames` - do you notice any patterns?

```{r 1.2response}
vacc <- clean_names(vacc)
colnames(vacc)
# Looks like many start with "percent" and some start with "total" - this indicates there are different units of measure for these different variables!
```

### 1.3

Let's rename the column "state_territory_federal_entity" in "vacc" to "entity" using `rename`. Make sure to reassign to `vacc` here and in subsequent steps.

```
# General format
new_data <- old_data %>% rename(newname = oldname)
```

```{r 1.3response}
vacc <- vacc %>% rename(entity = state_territory_federal_entity)
```

### 1.4

Select only the columns "entity", and those that start with "Percent". Use `select` and `starts_with("percent")`. 

```
# General format
new_data <- old_data %>% select(colname1, colname2, ...)
```

```{r 1.4response}
vacc <- vacc %>% select(entity, starts_with("percent"))
```

### 1.5

Using the data from 1.4, create a new dataset "vacc_long" that does `pivot_longer()` on all columns except "entity". Remember that `!entity` means all columns except "entity".

```
# General format
new_data <- old_data %>% pivot_longer(cols = colname(s))
```

```{r 1.5response}
vacc_long <- vacc %>% pivot_longer(cols = !entity)
```

### 1.6

Using `vacc_long`, filter the "entity" column so it only includes values in the following list: "Maryland", "Virginia", "Florida", "Massachusetts", "United States". **Hint**: use `filter` and `%in%`.

```
# General format
new_data <- old_data %>% filter(colname %in% c(1, 2, 3, ...))
```

```{r 1.6response}
vacc_long <- vacc_long %>%
  filter(entity %in% c("Maryland", "Virginia", "Mississippi", "Massachusetts", "United States"))
```

### 1.7

Use `pivot_wider` to reshape "vacc_long". Use "entity" for the `names_from` argument. Use "value" for the `values_from` argument. Call this new data `vacc_wide`. Look at the data. How do these states compare to one another.

```
# General format
new_data <- old_data %>% pivot_wider(names_from = column1, values_from = column2)
```

```{r 1.7response}
vacc_wide <- vacc_long %>%
  pivot_wider(
    names_from = entity,
    values_from = value
  )
vacc_wide
```


# Practice on Your Own!

### P.1

Take the code from Questions 1.1-1.7. Chain all of this code together using the pipe `%>%`. Call your data `vacc_compare`.

```{r P.1response}
vacc_compare <-
  read_csv("http://jhudatascience.org/intro_to_r/data/USA_covid19_vaccinations.csv") %>%
  clean_names() %>%
  rename(entity = state_territory_federal_entity) %>%
  select(entity, starts_with("percent")) %>%
  pivot_longer(cols = !entity) %>%
  filter(entity %in% c("Maryland", "Virginia", "Mississippi", "Massachusetts", "United States")) %>%
  pivot_wider(names_from = entity, values_from = value)
vacc_compare
```

### P.2

Modify the code from Question P.1:

-  Look for columns that start with "Total" (instead of "Percent") and 
-  Select different states/Entities to compare
-  Call your data `vacc_compare2`

```{r P.2response}
vacc_compare2 <-
  read_csv("http://jhudatascience.org/intro_to_r/data/USA_covid19_vaccinations.csv") %>%
  clean_names() %>%
  rename(entity = state_territory_federal_entity) %>%
  select(entity, starts_with("total")) %>%
  pivot_longer(cols = !entity) %>%
  filter(entity %in% c("Alaska", "Kansas", "California", "United States")) %>%
  pivot_wider(names_from = entity, values_from = value)
vacc_compare2
```


# Part 2

### 2.1

Read in the GDP and Personal Income Data from http://jhudatascience.org/intro_to_r/data/gdp_personal_income.csv. You can use the url or download the data. Call it "gdp".

```{r 2.1response}
gdp <- read_csv("http://jhudatascience.org/intro_to_r/data/gdp_personal_income.csv")
# If downloaded
# gdp <- read_csv("gdp_personal_income.csv")
```

### 2.2

First, clean up the column names with the `clean_names()` function in the `janitor` package. Then, use `pivot_wider` to reshape "gdp". Use `description` for the `names_from` argument. Use `x2020` for the `values_from` argument. Reassign this data to "gdp".

*Note that clean_names() has added an "x" before the original column name `2020`.*

```
# General format
new_data <- old_data %>% pivot_wider(names_from = column1, values_from = column2)
```

```{r 2.2response}
gdp <- gdp %>%
  clean_names() %>%
  pivot_wider(
    names_from = description,
    values_from = x2020
  )
```

### 2.3

Join the data. Keep only data that is found in both "vacc" and "gdp". 

- First, try joining without using the `by` argument - what happens?
- Next, try joining using `by = c("entity" = "geo_name")`. 
- Call the output "merged". How many observations (rows) are there?

```
# General format
new_data <- inner_join(x, y, by = c("colname1" = "colname2"))
```

```{r 2.3response}
# merged <- inner_join(vacc, gdp) does not work!
merged <- inner_join(vacc, gdp, by = c("entity" = "geo_name"))
nrow(merged)
```

### 2.4

Change your code from 2.3 to do a `full_join.` Call the output "full". How many observations (rows) are there?

```
# General format
new_data <- full_join(x, y, by = c("colname1" = "colname2"))
```

```{r 2.4response}
full <- full_join(vacc, gdp, by = c("entity" = "geo_name"))
nrow(full)
```


# Practice on Your Own!

### P.3

Do a left join of "vacc" and "gdp". Call the output "left". How many observations are there?

```{r P.3response}
left <- left_join(vacc, gdp, by = c("entity" = "geo_name"))
nrow(left)
```

### P.4

Copy your code from Question P.3 and change it to a `right_join` with the same order of the arguments. Call the output "right". How many observations are there?

```{r P.4response}
right <- right_join(vacc, gdp, by = c("entity" = "geo_name"))
nrow(right)
```

### P.5

Perform two `anti_join` operations on "vacc" and "gdp" to determine what Entities are missing from the GDP data and which are missing from the vaccine data.

```
# General format
anti_join(L, R, by = c("name_L" = "name_R")) %>% select(name_L)
```

```{r P.5response}
in_vacc_only <- anti_join(vacc, gdp, by = c("entity" = "geo_name")) %>% select(entity)
in_gdp_only <- anti_join(gdp, vacc, by = c("geo_name" = "entity")) %>% select(geo_name)
in_vacc_only
in_gdp_only
```
