---
title: "Manipulating Data in R Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data in this lab comes from the CDC (https://covid.cdc.gov/covid-data-tracker/#vaccinations_vacc-total-admin-rate-total - current as of January 12, 2022!) and the Bureau of Economic Analysis (https://www.bea.gov/data/income-saving/personal-income-by-state).

```{r message=FALSE}
library(readr)
library(dplyr)
library(tidyr)
```

# Part 1

1. Read in the SARS-CoV-2 Vaccination data from http://jhudatascience.org/intro_to_r/data/USA_covid19_vaccinations.csv. Assign the data the name "vacc". We will be reviewing new concepts here and incorporating some from week 1.

```{r}

```

2. Look at the column names using `colnames` - do you notice any patterns?

```{r}

```

3. Let's rename the column "State/Territory/Federal Entity" in "vacc" to "Entity" using `rename`.

```{r}

```

4. Create a new dataset "vacc_long" that does `pivot_longer` on all columns that start with the word "Percent". Use `starts_with("Percent")` for the `cols` argument. Use "Measure" for the `names_to` argument. Use "Percent" for the `values_to` argument.

```{r}
# ??? <- ??? %>% pivot_longer(cols = ???,
#                             names_to = ???,
#                             values_to = ???)
```

5. Select only the columns "Entity", "Measure", and "Percent" from "vacc_long". Use `select`. Filter "Entity" so it only includes values in the following list: "Maryland","Virginia","Florida","Massachusetts", "United States". **Hint**: use `filter` and `%in%`.

```{r}

```

6. Use `pivot_wider` to reshape "vacc_long". Use "Entity" for the `names_from` argument. Use "Percent" for the `values_from` argument.

```{r}
# ??? <- ??? %>%
#   pivot_wider(names_from = ???,
#               values_from = ???)
```

# Part 2

7. Read in the GDP and Personal Income Data from http://jhudatascience.org/intro_to_r/data/gdp_personal_income.csv. Call it "gdp". 

```{r}

```

8. Use `pivot_wider` to reshape "gdp". Use "Description" for the `names_from` argument. Use "`2020`" for the `values_from` argument. Reassign this data to "gdp".

```{r}

```

9. Join the data so that you only retain data in both "vacc" and "gdp". First, try merging without using the `by` argument - what happens? Next, try merging using `by = c("Entity" = "GeoName")`. Call the output "merged". How many observations are there?

```{r}

```

10. Join the data using a `full_join.` Call the output "full". How many observations are there?

```{r}

```

11. Do a left join of "vacc" and "gdp". How many observations are there?

```{r}

```

12. Repeat above with a `right_join` with the same order of the arguments. How many observations are there?

```{r}

```

13. Perform a `setdiff` on "vacc" and "gdp" to determine what Entities are missing from the GDP data and which are missing from the vaccine data.

```{r}

```

**Bonus**

14. Repeat #4 - #6 but instead look for columns that start with "Total" and select different states/Entities to compare. Also try combining all functions in one long block using the pipe `%>%`.

```{r}




```
