---
title: "Data Summarization Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Data used

Bike Lanes Dataset: BikeBaltimore is the Department of Transportation's bike program. 
The data is from http://data.baltimorecity.gov/Transportation/Bike-Lanes/xzfj-gyms

You can Download as a CSV in your current working directory.  Note its also available at: 	http://jhudatascience.org/intro_to_R_class/data/Bike_Lanes.csv 

```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(jhur)

bike = read_csv(
  "http://jhudatascience.org/intro_to_R_class/data/Bike_Lanes.csv")
```


or use 
```{r}
library(jhur)
bike = read_bike()
```

# Part 1

1. How many bike "lanes" are currently in Baltimore?  You can assume each observation/row is a different bike "lane".  (hint: how do you get the number of rows of a data set? You can use `dim()` or `nrow()`).
```{r q1}


```

2. How many (a) feet and (b) miles of bike "lanes" are currently in Baltimore, based on the `length` column? (`5280` feet in a mile).

```{r q2}
# sum(???)


```

3. How many types of bike lanes are there?   (Hints: `unique`, `table`, or `count()` on the column named `type`).
   
```{r q3}
# ??? %>% count(???)


```

4. Which bike lane`type` has (a) the most number of and (b) longest average bike lane length? (Hint: `group_by` and `summarize`)

```{r q4}
# ??? %>% 
#   group_by(???) %>% 
#   summarize(??? = n(),
#             ??? = mean(???))
```

5. Take your code from the above question, adding another pipe (`%>%`) and `arrange()` to sort the output by `mean`.

```{r q5}



```

6. Zeros aren't a very meaningful value for year. Change any zeros in `bike$dateInstalled` to `NA`. For the curious,`ifelse()` in R works takes the same arguments as the "IF" function in Excel!

```{r q6}
# ??? = ??? %>% 
#   mutate(??? = ifelse(dateInstalled == ???, NA, dateInstalled))


```

7. What was the average bike lane length grouped by `dateInstalled`? Remove `NA`s with `na.rm = TRUE`.

```{r q7}
# ??? %>%
#   group_by(???) %>% 
#   summarize(??? = mean(???, ???))

            
```


# Part 2

8.  Describe the distribution of bike "lane" lengths:
    (a) Numerically [hint: `quantile()`, using `bike$length`] and 
		(b) graphically [hint: `qplot()` with `geom = "histogram"` or `geom = "density")`].
```{r q8}

# (b)
# qplot(x = ???, data = ???, geom = ???)

```

9. Then describe as above, after stratifying by  i) type (`type`) then ii) number of lanes (`numLanes`) using `geom = "boxplot"`: 
```{r q9}



```

