---
title: "Data Summarization Lab Key"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Data used

Bike Lanes Dataset: BikeBaltimore is the Department of Transportation's bike program. 
The data is from http://data.baltimorecity.gov/Transportation/Bike-Lanes/xzfj-gyms

You can Download as a CSV in your current working directory.  Note its also available at: 	http://johnmuschelli.com/intro_to_r/data/Bike_Lanes.csv 

```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(jhur)

bike = read_csv(
  "http://johnmuschelli.com/intro_to_r/data/Bike_Lanes.csv")
```


or use 
```{r}
library(jhur)
bike = read_bike()
```

1. How many bike "lanes" are currently in Baltimore?  You can assume each observation/row is a different bike "lane"
```{r q1}
nrow(bike)
dim(bike)
```

2. How many (a) feet and (b) miles of bike "lanes" are currently in Baltimore?

```{r q2}
sum(bike$length)
sum(bike$length)/5280
sum(bike$length/5280)
```
3. How many types of bike lanes are there? Which type has (a) the most number of and (b) longest average bike lane length?
   
```{r q3}
table(bike$type, useNA = "ifany")
unique(bike$type)
length(table(bike$type))
length(unique(bike$type))

is.na(unique(bike$type))

bike %>% 
  group_by(type) %>% 
  summarise(n = n(),
            mean = mean(length)) %>% 
  arrange(mean)
```

4. How many different projects do the "bike" lanes fall into? 
		Which `project` category has the longest average bike lane? 

```{r q4}
length(unique(bike$project))

bike %>% 
  group_by(project, type) %>% 
  summarise(n = n(),
            mean = mean(length)) %>% 
  arrange(desc(mean))

arrange(summarize(group_by(bike, project, type), 
          n = n(), mean = mean(length)),
        desc(mean))

avg = bike %>% 
  group_by(type) %>% 
  summarize(mn = mean(length, na.rm = TRUE)) %>% 
  filter(mn == max(mn))
```

			  
5. What was the average bike lane length per year that they were installed? 
 Set `bike$dateInstalled` to `NA` if it is equal to zero.

```{r q5}
bike = bike %>% mutate(
  dateInstalled = ifelse(
    dateInstalled == 0, 
    NA,
    dateInstalled)
)
mean(bike$length[ !is.na(bike$dateInstalled)])

is.na(bike$dateInstalled)
!is.na(bike$dateInstalled)
```

```{r}
b2 = bike %>% 
  mutate(dateInstalled = ifelse(dateInstalled == "0", NA, 
                                dateInstalled))
 b2 = bike %>% 
   mutate(dateInstalled = if_else(dateInstalled == "0", 
                                  NA_integer_, 
                                 dateInstalled))
bike$dateInstalled[bike$dateInstalled == "0"] = NA

bike %>% 
  group_by(dateInstalled) %>% 
  summarise(n = n(),
            mean_of_the_bike = mean(length))
            
```

6. (a) Numerically [hint: `quantile()`] and 
		(b) graphically [hint: `hist()` or `plot(density())`]
		 describe the distribution of bike "lane" lengths.
```{r q6}
quantile(bike$length)

hist(bike$length)
hist(bike$length,breaks=100)
qplot(x = length, data = bike, geom = "histogram")
qplot(x = log10(length), data = bike, geom = "histogram")

hist(log2(bike$length),breaks=100)
hist(log10(bike$length),breaks=100)
```

7. Then describe as above, after stratifying by  i) type then ii) number of lanes
```{r q7}
qplot(y = length, x = type, data = bike, geom = "boxplot")
qplot(y = length, x = numLanes, data = bike, geom = "boxplot")
```

```{r}
boxplot(bike$length~bike$type)
boxplot(bike$length~bike$type, las=3)
boxplot(bike$length~bike$numLanes)

bike$length[1] = NA

boxplot(log10(bike$length)~bike$type)


bike %>% 
  group_by(type) %>% 
  summarise(q0.7 = quantile(length, na.rm = TRUE, probs =0.7))
```

