---
title: "Basic R Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



In this lab you can use the interactive console to explore or **Knit** the document. 
Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Ctrl-Enter (Windows/Linux) in an R code chunk.


# Part 1

1. Create a new object called `my.num` that contains 6 numbers.
```{r}
my.num <- c(5, 4, 7, 8, 12, 14) # These can be any real numbers
```


2. Multiply `my.num` by 4.
```{r}
my.num * 4
```


3. Create a second object called `my.char` that contains 5 character strings.
```{r}
my.char <- c("banana", "garlic", "pickles", "hot chocolate", "7-Up") # Again,these can be anything
```


4. Combine the two objects `my.num` and `my.char` into an object called `both`.
```{r}
both <- c(my.num, my.char)
```


5. What is the length of `both`?
```{r}
length(both)
```


6. What class is `both`?
```{r}
class(both)
```


**Bonus / Extra practice**: Create a vector that contains 4 sets of the numbers 1, 2, 3, and 4.
```{r}
vec_1 <- c(1, 2, 3, 4)
vec_2 <- c(vec_1, vec_1, vec_1, vec_1)
```


# Part 2

7. Divide `both` by 3, what happens?
```{r, error=TRUE}
both / 3
```


8. Create a vector with elements 1, 2, 3, 4, 5, 6 and call it `x`.
```{r}
x <- c(1, 2, 3, 4, 5, 6)
```


9. Create another vector with elements 10, 20, 30, 40, 50 and call it `y`.
```{r}
y <- c(10, 20, 30, 40, 50)
```


10. What happens when you add `x` and `y` together?
```{r}
x + y
# R "auto replicates" the shorter vector when the vectors aren't the same length - this is not what we want!
# In this case, it reads y as c(10, 20, 30, 40, 50, 10)!
```


12. Append the value 60 onto the vector `y` (hint: you can use the c() function).
```{r}
y <- c(y, 60)
```


13. Add `x` and `y` together.
```{r}
x + y
```


**Bonus / Extra practice**: Multiply `x` and `y` together. How is this similar to the way R performs addition in #13 ?
```{r}
x * y
# R adds together the first element of each vector, then the second, and so on. This order also applies when multiplying!
```


# Part 3

14. Create a vector object called `int_vect` that starts at 1 and goes up to 10. Use `seq()`.
```
# General format
seq(from = NUMBER, to = NUMBER)
```
```{r}
int_vect <- seq(from = 1, to = 10)
int_vect
```


15. Repeat the `int_vect` object this sequence 3 times using `rep()` and store the new object as `int_vect_3`.
```
# General format - times and each are optional
rep(x = OBJECT_TO_REPEAT, times = NUM_TIMES_TO_REPEAT, each = NUM_TIMES_TO_REPEAT_EACH_ELEMENT)
```
```{r}
int_vect_3 <- rep(int_vect, times = 3)
int_vect_3
```


16. What is the length of `int_vect_3`?
```{r}
length(int_vect_3)
```


17. Create a vector that takes the sequence "Strongly Agree", "Agree", "Neutral", "Disagree", "Strongly Disagree" and repeats each element 10 times.
```{r}
my_seq <- c("Strongly Agree", "Agree", "Neutral", "Disagree", "Strongly Disagree")
rep(my_seq, each = 10)

rep( c("Strongly Agree", "Agree", "Neutral", "Disagree", "Strongly Disagree"), each = 10 )
```


**Bonus / Extra practice**: "Strongly Agree", "Agree", "Neutral", "Disagree", "Strongly Disagree" are often responses to surveys. Create a randomly sampled vector of 30 survey responses. (hint use `sample()` and set the replace argument to `TRUE`). Store the output as `my_responses`. Examine the data by typing the name in the Console using a function.
```{r}
my_responses <- sample(
  x = c("Strongly Agree", "Agree", "Neutral", "Disagree", "Strongly Disagree"),
  size = 30,
  replace = TRUE
)
my_responses
```


**Bonus / Extra practice**: Let's say you change your survey so participants can rank their response 1-10 (inclusive). Create a randomly sampled vector of 30 survey responses. (hint use `seq()` and `sample()` and set the replace argument to `TRUE`). Store the output as `my_responses_2`. Examine the data by typing the name in the Console using a function.
```{r}
my_responses_2 <- sample(
  x = seq(from = 1, to = 10),
  size = 30,
  replace = TRUE
)
```
