---
title: "Basic R Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In this lab you can use the interactive console to explore or **Knit** the document. 
Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Ctrl-Enter (Windows/Linux) in an R code chunk.


# Part 1

1. Create a new object called `my.num` that contains 6 numbers.
```{r}
# my.num <- ???
```


2. Multiply `my.num` by 4.
```{r}

```


3. Create a second object called `my.char` that contains 5 character strings.
```{r}
# my.char <- ???
```


4. Combine the two objects `my.num` and `my.char` into an object called `both`.
```{r}

```


5. What is the length of `both`?
```{r}

```


6. What class is `both`?
```{r}

```


Bonus: Create a vector that contains 4 sets of the numbers 1, 2, 3, and 4.
```{r}

```

# Part 2

7. Divide `both` by 3, what happens?
```{r, error=TRUE}

```


8. Create a vector with elements 1 2 3 4 5 6 and call it `x`.
```{r}

```


9. Create another vector with elements 10 20 30 40 50 and call it `y`.
```{r}

```


10. What happens when you add `x` and `y` together?
```{r}

```


12. Append the value 60 onto the vector `y` (hint: you can use the c() function).
```{r}

```


13. Add `x` and `y` together.
```{r}

```


Bonus: Multiply `x` and `y` together. How is this similar to the way R performs addition in #13 ?
```{r}

```
