---
title: "Basic R Lab - Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Instructions: Run any code for the questions below in the console.

# Part 1

### 1.1 

Create a new object called `my.num` that contains any number.

```
# General format
my.num <- ?
```

```{r 1.1response}
my.num <- 2
```

### 1.2

Multiply `my.num` by 4.

```{r 1.2response}
my.num * 4
```

### 1.3

Create a second object called `my.char` that contains 5 character strings.

```
# General format
my.char <- c("character1", "character2", ...)
```

```{r 1.3response}
my.char <- c("banana", "garlic", "pickles", "hot chocolate", "7-Up") # Again,these can be anything
```

### 1.4

Combine the two objects `my.num` and `my.char` into an object called `both`.

```{r 1.4response}
both <- c(my.num, my.char)
```

### 1.5

What is the length of `both`? Use the `length()` function.

```{r 1.5response}
length(both)
```

### 1.6 

What class is `both`?

```{r 1.6response}
class(both)
```

### 1.7

Divide `both` by 3, what happens?

```{r, error=TRUE, label='1.7response'}
both / 3
```

# Practice on Your Own!

### P.1

Create a vector that contains 4 sets of the numbers 1, 2, 3, and 4.

```{r P.1response}
vec_1 <- c(1, 2, 3, 4)
vec_2 <- c(vec_1, vec_1, vec_1, vec_1)
```


# Part 2

### 2.1 

Create a vector with elements 1, 10, 100, 1000 and call it `z`.

```
# General format
z <- c(...)
```
 
```{r 2.1response}
z <- c(1, 10, 100, 1000)
```

### 2.2

Multiply each number in `z` by 5.
 
```{r 2.2response}
z * 5
```

### 2.3

Create a vector object called `int_vect` that starts at 1 and goes up to 10. Use `seq()`.

```
# General format
seq(from = NUMBER, to = NUMBER)
```

```{r 2.3response}
int_vect <- seq(from = 1, to = 10)
int_vect
```

### 2.4

What is the length of `int_vect`?

```{r 2.4response}
length(int_vect)
```

### 2.5

Install the tidyverse package using `install.packages("tidyverse")`. Load this package using `library(tidyverse)`.

# Practice on Your Own!

### P.2

Create a vector with elements 1, 2, 3, 4, 5 and call it `x`.

```
# General format
x <- c(...)
```
 
```{r P.2response}
x <- c(1, 2, 3, 4, 5)
```

### P.3

Create another vector with elements 10, 20, 30, 40, 50 and call it `y`.

```
# General format
y <- c(...)
```

```{r P.3response}
y <- c(10, 20, 30, 40, 50)
```

### P.4

Determine the length of `x` and `y`. Next, add the vectors x and y together.

```{r P.4response}
length(x)
length(y)
x + y
# [1] 11 22 33 44 55
```

### P.5

Append the value 60 onto the vector `y` (hint: you can use the `c()` function).

```
# General format
y <- c(y, ...)
```

```{r P.5response}
y <- c(y, 60)
```

### P.6

Determine the length of `x` and `y`.

```{r P.6response}
length(x)
length(y)
```

### P.7

Add `x` and `y` together. What happens?

```{r P.7response}
x + y
# [1] 11 22 33 44 55 61
# Warning message:
# In x + y : longer object length is not a multiple of shorter object length
# R "auto replicates" the shorter vector when the vectors aren't the same length - this is not what we want!
# In this case, it reads x as c(1, 2, 3, 4, 5, 1)!
```

### P.8

Multiply the following `a` and `b` together. How is this similar to the way R performs addition in question 2.6 ?

```{r}
a <- c(1, 2, 3)
b <- c(10, 100, 1000)
```

```{r P.8response}
a * b
# R adds together the first element of each vector, then the second, and so on. This order also applies when multiplying!
```

### P.9

Create a vector that takes the sequence "Strongly Agree", "Agree", "Neutral", "Disagree", "Strongly Disagree" and repeats each element 10 times.

```{r P.9response}
my_seq <- c("Strongly Agree", "Agree", "Neutral", "Disagree", "Strongly Disagree")
rep(my_seq, each = 10)

rep(c("Strongly Agree", "Agree", "Neutral", "Disagree", "Strongly Disagree"), each = 10)
```

### P.10

"Strongly Agree", "Agree", "Neutral", "Disagree", "Strongly Disagree" are often responses to surveys. Create a randomly sampled vector of 30 survey responses. (hint use `sample()` and set the replace argument to `TRUE`). Store the output as `my_responses`. Examine the data by typing the name `my_responses` in the Console.

```{r P.10response}
my_responses <- sample(
  x = c("Strongly Agree", "Agree", "Neutral", "Disagree", "Strongly Disagree"),
  size = 30,
  replace = TRUE
)
my_responses
```

### P.11

Let's say you change your survey so participants can rank their response 1-10 (inclusive). Create a randomly sampled vector of 30 survey responses. (hint use `seq()` and `sample()` and set the replace argument to `TRUE`). Store the output as `my_responses_2`. Examine the data by typing the name `my_responses_2` in the Console.

```{r P.11response}
my_responses_2 <- sample(
  x = seq(from = 1, to = 10),
  size = 30,
  replace = TRUE
)
```
