---
title: "Data Input Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Part 1

## Helpful tips before we start

### TROUBLESHOOTING: Common new user mistakes we have seen

- Check the file path -- is the file there?  
- Typos (R is **case sensitive**, `x` and `X` are different)  
- Open ended quotes, parentheses, and brackets     
- Deleting part of the code chunk  
- For any function, you can write `?FUNCTION_NAME`, or `help("FUNCTION_NAME")` to look at the help file

### 1.1

Set up your R Project.

* File, New Project or click the new project button  
* New Directory  
* New Project  
* Type a name and choose a location  
* Check that the folder is there!  

Check out our resource here: https://jhudatascience.org/intro_to_r/resources/R_Projects.html

### 1.2 

Load the package by adding "library(tidyverse)" below and running the code.

```{r 1.2response}

```

### 1.3

Use the manual import method (File > Import Dataset > From Text (`readr`)) to Read in SARS-CoV-2 vaccination data from this URL:

https://jhudatascience.org/intro_to_r/data/vaccinations.csv

You can learn more about how the data was collected here: https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-Jurisdi/unsk-b7fc

### 1.4

What is the dataset object called? You can find this information in the Console or the Environment. Enter your answer as a comment using `#`.

```{r 1.4response}

```

### 1.5

Preview the data by examining the Environment. How many observations and variables are there? Enter your answer as a comment using `#`.

```{r 1.5response}

```

# Practice on Your Own!

### P.1

Download the data from https://jhudatascience.org/intro_to_r/data/vaccinations.csv and move the file to your project folder. Import the data by browsing for the file on your computer.

> *Download the data*
> *Put data in the project folder*
> File, Import Dataset, From Text (`readr`)
> browse for the file
> click "Update" and "Import"


# Part 2

### 2.1

Read in SARS-CoV-2 vaccination data from URL https://jhudatascience.org/intro_to_r/data/vaccinations.csv. Assign it to an object named `vacc`. Use the code structure below.

```
# General format
# OBJECT <- read_csv(FILE)
```

```{r 2.1response}

```

### 2.2

Take a look at the data. Do these data objects (`vaccinations` and `vacc`) appear to be the same? Why or why not?

```{r 2.2response}

```

### 2.3

Learn your working directory by running `getwd()`. This is where R will look for files unless you tell it otherwise.

```{r 2.3response}

```


# Practice on Your Own!

### P.2

Run the following code - is there a problem? How do you know?

```{r eval=FALSE}
vacc2 <- read_delim("https://jhudatascience.org/intro_to_r/data/vaccinations.csv", delim = "\t")
vacc2
```

```{r P.2response}

```

### P.3

Try reading in some data on your computer using any method we discussed!
