---
title: "Data_Input_Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

In this lab you can use the interactive console to explore or **Knit** the document. 
Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Ctrl-Enter (Windows/Linux) in an R code chunk.

# Part 1  

1\. Use the manual import method (File > Import Dataset > From Text (`readr`)) to Read in SARS-CoV-2 vaccination data from URL 
http://jhudatascience.org/intro_to_r/data/vaccinations.csv. You can learn more about how the data was collected here: https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-Jurisdi/unsk-b7fc

2\. What is the dataset object called? You can find this information in the Console or the Environment. Enter your answer as a comment using `#`.

```{r}
# vaccinations
```


3\. Preview the data by clicking the table button in the Environment. How many observations and variables are there? Enter your answer as a comment using `#`.

```{r}
# 37272 obs. of 109 variables
```


# Part 2

4\. Read in SARS-CoV-2 vaccination data from URL 
http://jhudatascience.org/intro_to_r/data/vaccinations.csv
and assign it to an object named `vacc`. Use the code structure below.

```
# General format
library(readr)
# OBJECT <- read_csv(FILE)
```

```{r}
library(readr)
vacc <- read_csv(file = "http://jhudatascience.org/intro_to_r/data/vaccinations.csv")
```


5\. Use `head()` to display the first few rows of the data frame. 

```{r}
head(vacc)
```


6\. Use `str()` to display the list of column names and their type specification. 

```{r}
str(vacc)
```


7\. Load the `readxl` package with the `library()` command. 

If it is not installed, install it via: `RStudio --> Tools --> Install Packages`. You can also try `install.packages("readxl")`.

```{r}
library(readxl)
```


8\. Download the dataset of asthma prevalence in the USA from: http://jhudatascience.org/intro_to_r/data/asthma.xlsx file to `asthma.xlsx` by running the following code chunk. This only downloads the file, it does NOT bring the file into R.

```{r}
download.file(
  url = "http://jhudatascience.org/intro_to_r/data/asthma.xlsx",
  destfile = "asthma.xlsx",
  overwrite = TRUE,
  mode = "wb"
)
```


9\. Use the `read_excel()` function in the `readxl` package to read the `asthma.xlsx` file and call the output `asthma`. 

```{r}
asthma <- read_excel(path = "asthma.xlsx")
```


10\. Use the `head()` and `str()` functions to inspect the data.

```{r}
head(asthma)
str(asthma)
```


**Bonus / Extra practice**: Run the following code - is there a problem? How do you know?

```{r}
yts <- read_delim("http://jhudatascience.org/intro_to_r/data/Youth_Tobacco_Survey_YTS_Data.csv", delim = "\t")
yts
# It should be a red flag to see that there is only one column that looks like: "YEAR,LocationAbbr,LocationDesc,TopicType,TopicDesc,MeasureDesc..."
# This file is comma delimited not tab delimited!
```


**Bonus / Extra practice**: By default, R reads the first sheet of an excel file. Copy your code from question 9 and add the following argument: `sheet = 2`. Inspect the data using `head()`.

```{r}
asthma <- read_excel(path = "asthma.xlsx", sheet = 2)
head(asthma)
```


**Bonus / Extra practice**: Install and load the `haven` package. Look at the help page for `read_dta()` function, and scroll to the very bottom of the page. Try running some of the examples provided.

```{r eval = FALSE}
install.packages("haven")
library(haven)
?read_dta

path <- system.file("examples", "iris.dta", package = "haven")
read_dta(path)
```


# Part 3

11\. Learn your working directory by running `getwd()`

```{r}
getwd()
```


12\. Create a folder in your R project called data. Move the "asthma.xlsx" file there. 


13\. Modify the following code so that it finds "asthma.xlsx" in the "data" directory.

```{r eval=FALSE}
asthma <- read_excel(path = "data/asthma.xlsx")
```


**Bonus / Extra practice**: Practice importing a dataset of your choice, give it an object name, and use `head()` to preview the first few lines.



lines


asdasdsd
