---
title: "Data Input Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Helpful tips before we start

### TROUBLESHOOTING: Common new user mistakes we have seen

- Check the file path -- is the file there?  
- Typos (R is **case sensitive**, `x` and `X` are different)  
- Open ended quotes, parentheses, and brackets     
- Deleting part of the code chunk  
- For any function, you can write `?FUNCTION_NAME`, or `help("FUNCTION_NAME")` to look at the help file

### R Projects

R Projects can help you keep files organized and avoid issues with working directories. Check out our resource here: https://jhudatascience.org/intro_to_r/resources/R_Projects.html


## Lab

In this lab you can use the interactive console to explore or **Knit** the document. 
Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Ctrl-Enter (Windows/Linux) in an R code chunk.

```{r}
# Load the necessary package
library(readr)
```

### 1.1

Use the manual import method (File > Import Dataset > From Text (`readr`)) to Read in SARS-CoV-2 vaccination data from this URL:

https://jhudatascience.org/intro_to_r/data/vaccinations.csv. 

You can learn more about how the data was collected here: https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-Jurisdi/unsk-b7fc

### 1.2

What is the dataset object called? You can find this information in the Console or the Environment. Enter your answer as a comment using `#`.

```{r 1.2response}

```

### 1.3

Preview the data by clicking the table button in the Environment. How many observations and variables are there? Enter your answer as a comment using `#`.

```{r 1.3response}

```

### 1.4

Read in SARS-CoV-2 vaccination data from URL 
https://jhudatascience.org/intro_to_r/data/vaccinations.csv
and assign it to an object named `vacc`. Use the code structure below.

```
# General format
library(readr)
# OBJECT <- read_csv(FILE)
```

```{r 1.4response}

```

### 1.5

Take a look at the data. Do these data objects (`vaccinations` and `vacc`) appear to be the same? Why or why not?

```{r 1.5response}

```

### 1.6

Learn your working directory by running `getwd()`. This is where R will look for files unless you tell it otherwise.

```{r 1.6response}

```


# Practice on Your Own!

### P.1

Load the `readxl` package with the `library()` command. 

If it is not installed, install it via: `RStudio --> Tools --> Install Packages`. You can also try `install.packages("readxl")`.

```{r P.1response}

```

### P.2

Download the dataset of asthma prevalence in the USA from: https://jhudatascience.org/intro_to_r/data/asthma.xlsx file to `asthma.xlsx` by running the following code chunk. This only downloads the file, it does NOT bring the file into R.

```{r}
download.file(
  url = "https://jhudatascience.org/intro_to_r/data/asthma.xlsx",
  destfile = "asthma.xlsx",
  overwrite = TRUE,
  mode = "wb"
)
```

Note: the "wb" option makes sure the file can be read correctly on Windows and Apple machines.

### P.3

Use the `read_excel()` function in the `readxl` package to read the `asthma.xlsx` file and call the output `asthma`. 

```{r P.3response}

```

### P.4

Run the following code - is there a problem? How do you know?

```{r}
yts <- read_delim("https://jhudatascience.org/intro_to_r/data/Youth_Tobacco_Survey_YTS_Data.csv", delim = "\t")
yts
```

```{r P.4response}

```

### P.5

By default, R reads the first sheet of an excel file. Copy your code from question P.3 and add the following argument: `sheet = 2`. Inspect the data using `head()`.

```{r P.5response}

```

### P.6

Install and load the `haven` package. Look at the help page for `read_dta()` function, and scroll to the very bottom of the page. Try running some of the examples provided.

```{r eval = FALSE, label = 'P.6response'}

```

