---
title: "Data Class Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

Dataset used is the Bike Lanes Dataset: BikeBaltimore is the Department of Transportation's bike program. 

The data is from http://data.baltimorecity.gov/Transportation/Bike-Lanes/xzfj-gyms

# Part 1

0. Load all the libraries we will use in this lab. 

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(jhur)
```

1. Read Bike Lanes Dataset using `read_bike()` function from `jhur` package. Assign it to `bike` variable. Use `str()` function to take a look at the data and learn column types.

Note: alternatively you can read data to R from URL `"http://jhudatascience.org/intro_to_r/data/Bike_Lanes.csv"` using `read_csv()` function. 

```{r}
bike = read_bike()
str(bike)
```

2. Use `mutate()` function to create a new variable in the data set named `type_fact` that is of class `factor` made from the `type` variable. 

```{r}
bike <- mutate(bike, type_fact = factor(type))
```

3. Use `str()` function again. Note the difference between `type` and `type_fact` variables 

```{r}
str(bike)
```

4. Use `levels()` function on `type_fact` variable to display levels of `type_fact` variable. Hint: recall you can use `data$variable_name` to access variable of interest. 

```{r}
levels(bike$type_fact)
```

5. Use `table()` function on `type_fact` variable to display the count of observations with each bike type. 

Use `table()` function again and specify additional argumet `useNA = "always"`. 

```{r}
table(bike$type_fact)
table(bike$type_fact, useNA = "always")
```


# Part 2

1. Read Read Charm City Circulator data using `read_circulator()` function from `jhur` package. Assign it to `circ` variable. Use `str()` function to take a look at the data and learn column types.

```{r}
circ = read_circulator()
str(circ)
```

2. Use `mutate()` function to create a new column named `date_formatted` that is of `Date` class. The new variable is created from `date` column. Hint: use `mdy()` function.

```{r}
circ = mutate(circ, date_formatted = mdy(date))
```

3. Use `str()` function again. Note the difference between `date` and `date_formatted` columns.  

```{r}
str(circ)
```

5. Use `range()` function on `date_formatted` variable to display the range of dates in the data set.

```{r}
range(circ$date_formatted)
```
 
6. Use `mutate()` function to create a new column named `date_year` that represents year of the calendar date. The new variable is created from `date_formatted` column. Hint: use `year()` function.

```{r}
circ = mutate(circ, date_year = year(date_formatted))
```

7. Use `table()` function on `date_year` variable to display the count of observations with each calendar year. 

```{r}
table(circ$date_year)
```
 
 
# Part 3

1. Use function `seq()` to generate 16 numeric values ranging from 1 to 16 Assign them to a variable named `vals`. 

```{r}
vals <- seq(from = 1, to = 16)
vals
```

2. Use values from the `vals` vector to create 5x5 matrix. Assign the matrix to variable named `mat`. 

```{r}
mat <- matrix(vals, nrow = 4)
mat
```

3. What happens when you type `mat + 100`? 

```{r}
mat + 100
```

4. What happens when you type `mat + mat`? 

```{r}
mat + mat
```



