---
title: "Data Class Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

Dataset used is the Bike Lanes Dataset: BikeBaltimore is the Department of Transportation's bike program. 

The data is from http://data.baltimorecity.gov/Transportation/Bike-Lanes/xzfj-gyms

# Part 1

0. Load all the libraries we will use in this lab. 

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(jhur)
```

1. Create some data to work with. 
First, create a vector that has class integer called `int_vect` that starts at 1 and goes up to 10 and repeats this sequence 3 times using `rep`  (hint use `seq()`).
Next, create a random vector of integers from 1 to 30 called `rand_vect` (hint use `sample()`).
Create a vector that repeats `c(TRUE, TRUE, FALSE)` 10 times called `TF_vect`. 
Create a vector that repeats `c("TRUE", "TRUE", "FALSE")` 10 times called `TF_vect2`.

Because we are using a random vector let's use the `set.seed()` function to make sure we all have the same result - this code will already be in the code chunk for you. Simply create the vectors below the `set.seed` line.

```{r}
set.seed(1234)
int_vect <- rep(seq(1, 10), 3)
rand_vect <- sample( 1:30, size = 30, replace = TRUE)
TF_vect <- rep(c(TRUE, TRUE, FALSE), times = 10)
TF_vect2 <- rep(c("TRUE", "TRUE", "FALSE"), times = 10)
```

Create a tibble combining these vectors together called vect_data using the following code.

```{r}
vect_data <- tibble(int_vect, rand_vect, TF_vect, TF_vect2)
```


2. Take a look at 5 random rows using the `slice_sample()` function. Try this a few times to see how the results change.
```{r}
slice_sample(vect_data, n = 5)
slice_sample(vect_data, n = 5)
slice_sample(vect_data, n = 5)
```

3. Check to see if the `TF_vect` is logical. Check to see if `TF_vect2` is logical. Why are the results what they are?
```{r}
is.logical(TF_vect)
is.logical(TF_vect2)
```

Logical vectors do not have quotes around `TRUE` and `FALSE` values.

4. Use `mutate()` function to create a new variable in the data set named `type_fact` that is of class `factor` made from the `int_vect` variable. Take a look at the data. 

```{r}
vect_data <- mutate(vect_data, type_fact = factor(int_vect))
vect_data
```



5. Use `levels()` function on `type_fact` variable to display levels of `type_fact` variable. Hint: recall you can use `data$variable_name` to access variable of interest. 

```{r}
levels(vect_data$type_fact)
```



# Part 2

1. Read Read Charm City Circulator data using `read_circulator()` function from `jhur` package. Assign it to `circ` variable. Use `str()` function to take a look at the data and learn column types.

```{r}
circ = read_circulator()
str(circ)
```

2. Use `mutate()` function to create a new column named `date_formatted` that is of `Date` class. The new variable is created from `date` column. Hint: use `mdy()` function.

```{r}
circ = mutate(circ, date_formatted = mdy(date))
```

3. Use `str()` function again. Note the difference between `date` and `date_formatted` columns.  

```{r}
str(circ)
```

5. Use `range()` function on `date_formatted` variable to display the range of dates in the data set.

```{r}
range(circ$date_formatted)
```
 
6. Use `mutate()` function to create a new column named `date_year` that represents year of the calendar date. The new variable is created from `date_formatted` column. Hint: use `year()` function.

```{r}
circ = mutate(circ, date_year = year(date_formatted))
```

7. Use `table()` function on `date_year` variable to display the count of observations with each calendar year. 

```{r}
table(circ$date_year)
```
 
 
# Part 3

1. Use function `seq()` to generate 16 numeric values ranging from 1 to 16 Assign them to a variable named `vals`. 

```{r}
vals <- seq(from = 1, to = 16)
vals
```

2. Use values from the `vals` vector to create 5x5 matrix. Assign the matrix to variable named `mat`. 

```{r}
mat <- matrix(vals, nrow = 4)
mat
```

3. What happens when you type `mat + 100`? 

```{r}
mat + 100
```

4. What happens when you type `mat + mat`? 

```{r}
mat + mat
```



