---
title: "Data Class Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---


# Data used

Bike Lanes Dataset: BikeBaltimore is the Department of Transportation's bike program. 
The data is from http://data.baltimorecity.gov/Transportation/Bike-Lanes/xzfj-gyms

You can Download as a CSV in your current working directory.  Note its also available at: 	http://johnmuschelli.com/intro_to_r/data/Bike_Lanes.csv 

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(jhur)

bike = read_csv(
  "http://johnmuschelli.com/intro_to_r/data/Bike_Lanes.csv")
```


or use 
```{r}
library(jhur)
bike = read_bike()
```



# Part 1

1.  Get all the differet types of bike types from the `type` column.
 Use `sort(unique())`.  Assign this to an object `btypes.` Type `dput(btypes)`
 
```{r q1}
head(factor(bike$type))

btypes = sort(unique(bike$type))
x = c("SIDEPATH","BIKE BOULEVARD", "BIKE LANE", "CONTRAFLOW",
        "SHARED BUS BIKE",  "SHARROW",  "SIGNED ROUTE")
dput(btypes)
dput(btypes)[c(6,1:5,7)]
dput(btypes[c(6,1:5,7)])

lev = c( "SIDEPATH", "BIKE BOULEVARD", "BIKE LANE", "CONTRAFLOW", "SHARED BUS BIKE", 
        "SHARROW", "SIGNED ROUTE")
```

2. Reorder the output of the `dput` command so `SIDEPATH` is first.  
 Recode type as a factor using this vector of levels.  
 Print `head(bike$type)`.  Note what you see.  Run 
 `table(bike$type)` afterwards and note the order
```{r q2}
bike$type = factor(bike$type)
bike$type = relevel(bike$type, "SIDEPATH")
bike$type = factor(bike$type,
          levels = dput(btypes[c(6,1:5,7)]))
bike = bike %>% mutate(type = factor(type, 
                levels = dput(btypes[c(6,1:5,7)])))

table(bike$type)
```

3. Make a column `type2`, which is a factor of type, with the 
 levels: 
 `c( "SIDEPATH", "BIKE BOULEVARD", "BIKE LANE")`.  Run `table(bike$type2)`,
 with the option `useNA = "always"`.  Note, we do not have to 
 make type a character again before doing this
```{r q3}
bike = bike %>% 
  mutate(type2 = factor(type, 
             levels = c( "SIDEPATH", "BIKE BOULEVARD", 
                         "BIKE LANE") ) )
table(bike$type)
table(bike$type2)
table(bike$type2, useNA = "always")
```

4. 
 a) Reassign `dateInstalled` into a character using `as.character`.
 Run `head(bike$dateInstalled)`.
```{r q4a}
bike = bike %>% 
  mutate(dateInstalled = 
           as.character(dateInstalled)
  )
head(bike$dateInstalled)
```


 b) Reassign `dateInstalled` a factor, using the default levels.
 Run `head(bike$dateInstalled)`.
```{r q4b}
bike = bike %>% 
  mutate(dateInstalled = 
           factor(dateInstalled)
  )
head(bike$dateInstalled)

table(factor(bike$dateInstalled, levels = 2005:2017))
table(factor(bike$dateInstalled, levels = 2005:2017), 
        useNA="ifany")
```

 c) Do not reassign `dateInstalled`, but simply run
 `head(as.numeric(bike$dateInstalled))`  we are looking to see what happens when we try to go from factor to numeric
```{r q4c}
head(as.numeric(bike$dateInstalled)) 
```


 d) Do not reassign `dateInstalled`, but simply run
 `head(as.numeric(as.character(bike$dateInstalled)))`
 this is how you get "numeric" values back if they were "incorrectly" 
 factors
```{r q4d}
head(as.numeric(as.character(
    bike$dateInstalled)))
```


# Part 2

5.  Convert `type` back to a character.
 Make a column `type2` (replacing the old one)
 where if the type is one of these categories:
 `c("CONTRAFLOW", "SHARED BUS BIKE", "SHARROW", "SIGNED ROUTE")`
 call it `"OTHER"`.  Use `%in%` and `ifelse`.  Make `type2` a factor
 with the levels `c( "SIDEPATH", "BIKE BOULEVARD", "BIKE LANE", "OTHER")`
```{r q5}
 
bike = bike %>% mutate(
    type = as.character(type),
    type2 = ifelse(type %in% c("CONTRAFLOW", "SHARED BUS BIKE", 
                               "SHARROW", "SIGNED ROUTE"), "OTHER", type),
    type2 = factor(type2, levels = c( "SIDEPATH", "BIKE BOULEVARD", 
                               "BIKE LANE", "OTHER") ))

table(bike$type2)

bike2 = bike %>% 
  mutate(
    type = factor(type,
                  levels = c( "SIDEPATH", "BIKE BOULEVARD", 
                              "BIKE LANE", "CONTRAFLOW", 
                              "SHARED BUS BIKE", 
                              "SHARROW", "SIGNED ROUTE")
                  ),
    type2 = recode_factor(type, 
                          "CONTRAFLOW" = "OTHER",
                          "SHARED BUS BIKE" = "OTHER",
                          "SHARROW" = "OTHER",
                          "SIGNED ROUTE" = "OTHER")
  )
table(bike2$type2)
```

# Part 3

6.  Parse the following dates with the correct `lubridate` function:

a) "2014/02-14"

```{r q6a}
ymd("2014/02-14")
```

 b) "04/22/14 03:20"
 
```{r q6b}
mdy_hm("04/22/14 03:20")
```

c) "4/5/2016 03:2:22"  assume `mdy`

```{r q6c}
mdy_hms("4/5/2016 03:2:22")
```










