---
title: "Cleaning complicated column names"
output: 
  html_document:
    css: ../docs/web_styles.css
    toc: true
---


## Cleaning a common pattern from names

Let's say that we already have technically clean names - in that they don't have spaces or punctuation or start with a number. However, let's say that there is a redundant word ("percent") that we want to remove or add to multiple columns.


First let's load the packages we will need. We will show some functions from   `janitor` and the `tidyverse`:

```{r, echo = FALSE}
install.packages("janitor", repos='http://cran.us.r-project.org')
```

```{r}
#install.packages("janitor")
library(tidyverse)
library(janitor)
```

First let's make some data:

```{r}
data_to_clean <- tibble(State = c("Texas", "Utah", "Maryland", "Ohio"),
                        tax_percent = c(10, 20, 60, 40),
                        literacy_percent = c(70, 80, 80, 75),
                        above_poverty_percent = c(60, 70, 50, 60))
data_to_clean

```


We can use the `rename_with` function of `dplyr` and `str_remove` of `stringr` to remove the pattern "_percent" from each of the column names.

Here we use the `~` and the `.` to indicate that we are using `str_remove` and all the column names. If it finds the pattern it will remove it.


```{r}
data_to_clean %>% rename_with(~str_remove(., '_percent'))

```

Nice! That simplified our names very easily!

## Cleaning names with numbers and punctuation


We can use patterns with regex - see this [regex cheatsheet](https://evoldyn.gitlab.io/evomics-2018/ref-sheets/R_strings.pdf) for help to remove unwanted characters! We adapted some code from this [source](https://stackoverflow.com/questions/71151470/remove-characters-from-column-names).

First we will make some very messy data:

```{r}

d <- tibble("Year" = 1:5,
       "Info" = 1:5,
       "1. Products" = 1:5,
       "2. Rate" = 1:5,
       "3. Price" = 1:5,
       "29. Other" = 1:5)
d
```

Now we can remove the numbers and punctuation in a similar way as we did before using `rename_with` and `str_remove`, but this time we specify a few things:

- that we want to remove digits with `[:digits:]` (based on the [regex cheatsheet](https://evoldyn.gitlab.io/evomics-2018/ref-sheets/R_strings.pdf))

- that we want to remove possibly one or more digits with the `+` (based on the [regex cheatsheet](https://evoldyn.gitlab.io/evomics-2018/ref-sheets/R_strings.pdf))

- that we want to remove a period (which needs two `\\` based on the (based on the [regex cheatsheet](https://evoldyn.gitlab.io/evomics-2018/ref-sheets/R_strings.pdf)) too!) and a space

Here we go:

```{r}
d %>% 
  rename_with(~str_remove(., "[:digit:]+\\. "))

```

Nice, that is better!

## Using values of a specific row for column names

First let's make some messy data that is missing values in the first row and has possible better column names in the second row. We adapted code from this [source](https://cran.r-project.org/web/packages/janitor/vignettes/janitor.html#remove_constant-columns).

This can often happen when we read in data.

```{r}

dirt <- data.frame(X_1 = c(NA, "ID", 1:3),
           X_2 = c(NA, "Value", 4:6))

dirt
```


The function `row_to_names` from the `janitor` package (not part of the `tidyverse` - so make sure you install and load it!) can be really helpful for this. 

We can use the `row_number` argument of `row_to_names` to specify that the column names can be found in the second row.

```{r}
row_to_names(dirt, row_number = 2) # our column names can be found in row 2!

```


