---
title: "Functions that need pull first"
output: 
  html_document:
    css: ../docs/web_styles.css
---

When can you use a function directly on a column of a tibble and when do you need a vector in R?

There are some inconsistencies across different functions. Typically if it is a base R function, you often need to pull the data out into a vector first.  Sometimes you just need to try them out. Always check that your work did what you expect!

## In general {.tabset .tabset-fade}

In general these functions need to use vectors:

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(kableExtra)
library(tidyverse)
text_tbl <- data.frame(
  Type = c("base R and tidyverse function", "special tidyverse functions", "base R functions", "stat functions that come with R"),
  Example = c("setdiff()", "stringr functions like str_replace()", "mean()", "cor.test()"))
kbl(text_tbl) %>%
  # kable_classic(full_width = F, html_font = "helvetica") %>%
  kable_paper(full_width = F, html_font = "helvetica", font_size = "18") %>%
  column_spec(1:2, bold = T, border_right = T, border_left = TRUE) %>%
  row_spec(0, align = "c") %>%
  row_spec(1, background = "yellow") %>%
  row_spec(2, background = "lightgreen") %>%
  row_spec(3:4, background = "lightblue")
```

<br>

Click below to take a look at each in more detail.

<br>

### `setdiff()` function

This function was originally a base R function and it got adopted by tidyverse packages (`dplyr` and `lubridate`) and is therefore can behave a bit differently. This is confusing and part of the struggle of an open development environment.

When using `setdiff()` from base R, vectors are required - meaning we need to use `pull` first. It shows the **elements in the first that are missing from the second**. 

Using `::` we can specify which package we want to use the function from.


```{r}
data_As <- tibble(State = c("Alabama", "Alaska"),
                 state_bird = c("wild turkey", "willow ptarmigan"))
data_cold <- tibble(State = c("Maine", "Alaska", "Alaska"),
                    vacc_rate = c(0.795, 0.623, 0.626),
                    month = c("April", "April", "May"))

A_states_vector <- data_As %>% pull(State)
cold_states_vector <- data_cold %>% pull(State)

A_states_vector
cold_states_vector

base::setdiff(A_states_vector, cold_states_vector)
```
"Alabama" is in the first vector, but not in the second vector.

If we do this with the `dplyr` version this still works the same:

```{r}
dplyr::setdiff(A_states_vector, cold_states_vector)
```

We can select the `State` variable to compare (keeping it as a tibble) and it still works as the above with the `dplyr` version, even though `A_states` and `cold_states` are tibbles and not vectors. (Remember `select()` creates a smaller tibble and `pull()` creates a vector of the values.) In this case we see the rows in the first tibble that are not in the second.

```{r}
A_states_tibble <- data_As %>% select(State)
A_states_tibble
cold_states_tibble <- data_cold %>% select(State)
cold_states_tibble
dplyr::setdiff(A_states_tibble, cold_states_tibble)

```

However when using `setdiff()` from base R with the tibble versions of the data, it does not work properly and just gives us the first tibble.

```{r}
base::setdiff(A_states_tibble, cold_states_tibble) 
base::setdiff(cold_states_tibble, A_states_tibble) 
```

We could also use `setdiff()` from `dplyr` to tell us what rows were removed when filtering a dataframe or tibble:

```{r}
mt_cars_high_mpg <- mtcars %>% filter(mpg > 20)
dplyr::setdiff(mtcars, mt_cars_high_mpg)
```

Here we see all the rows with `mpg` below 20. 

### `stringr` functions

These functions are often applied within `filter()` or `mutate()` for a data frame. When they are not used inside these functions they need to be used on a vector. 

```{r}
iris %>% filter(str_detect(string = Species, pattern = "set")) %>% head() # this will work

iris %>% pull(Species) %>% str_detect(pattern = "set") %>% head() # so will this
```

This however, would not work well:

```{r, eval = FALSE}
iris %>% select(Species) %>% str_detect(pattern = "set")
```


### `base` R math functions

Functions like `sum()` or `mean()` need vectors. We can use them well within `summarize` but would need to use a vector otherwise. 

```{r}
iris %>% summarize(mean_Petal_Length = mean(Petal.Length))

iris %>% pull(Petal.Length) %>% mean()
```

This does not work:

```{r, eval = FALSE}
iris %>% select(Petal.Length) %>% mean()

```



### `stats` functions

Functions like `cor.test` of the `stats` package (which come automatically with R) need vectors too. Note that some of the other `stats` functions are tolerant to using tibbles. 

```{r}

x <-iris %>% pull(Petal.Length)
y <-iris %>% pull(Petal.Width)

cor.test(x,y)
```

This does not work:

```{r, eval = FALSE}
x <-iris %>% select(Petal.Length)
y <-iris %>% select(Petal.Width)

cor.test(x,y)
```
