---
title: "Data Cleaning Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Data used

Bike Lanes Dataset: BikeBaltimore is the Department of Transportation's bike program. 
The data is from http://data.baltimorecity.gov/Transportation/Bike-Lanes/xzfj-gyms

You can Download as a CSV in your current working directory.  Note its also available at: 	http://jhudatascience.org/intro_to_r/data/Bike_Lanes.csv 

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(jhur)
library(tidyverse)
library(broom)
library(naniar)

bike <- jhur::read_bike()
```

# Part 1


1.  Use the `is.na()`  and `any()` functions to check if the bike `dateInstalled` variable has any `NA` values.

```{r}
any(is.na(bike$dateInstalled))
# or
bike %>%
  pull(dateInstalled) %>%
  is.na() %>%
  any()
```



2.  Filter rows of bike, so that only rows remain that do NOT have missing values  for the `route` variable,  using `drop_na`. Assign this to the object `have_route.`

```{r}
have_rout <- bike %>% drop_na(route)
```

3.  Use `naniar` to make a visual of the amount of data missing for each variable of `bike` (use `naniar::gg_miss_var()`). Check out more about this package here: https://www.njtierney.com/post/2018/06/12/naniar-on-cran/

```{r}
naniar::gg_miss_var(bike)
```


4. What percentage of the `subType` variable is complete of `bike` ? Hint: use another `naniar` function.
 
```{r}
naniar::pct_complete(bike$subType) # this or..
pull(bike, subType) %>% pct_complete() # this
```

# Part 2

**New Data set**
Now imagine we work in a clinic and we are trying to understand more about blood types of patients. 

Let's say we the data like so:

```{r}
BloodType <- tibble(
  gender =
    c(
      "M", "Male", "Female", "F", "M",
      "Male", "Other", "M", "F", "Other",
      "F", "Male", NA, "Male", "Female"
    ),
  type = c(
    "A.-", "AB.+", "O.-", "O.+", "AB.-",
    "B.+", "B.-", "o.-", "O.+", "A.-",
    "A.+", "O.-", "B.-", "o.+", "AB.-"
  ),
  infection = c(
    "Yes", "No", "Yes", "No", "No",
    "No", "Yes", "No", "Yes", "No",
    "No", "Yes", "Yes", "Yes", "NotSure"
  )
)

BloodType
```

There are some issues with this data that we need to figure out!

1. Determine how many `NA` values there are for `gender`.

```{r}
count(BloodType, gender)
sum(is.na(pull(BloodType, gender)))
BloodType %>%
  pull(gender) %>%
  is.na() %>%
  sum()
```

2. Recode the `gender` variable of the `BloodType` data so that it is consistent. Use `case_when()`.

```{r}

BloodType <- BloodType %>%
  mutate(gender = case_when(
    gender %in% c("M", "m", "Male") ~ "Male",
    gender %in% c("F", "female", "Female") ~ "Female",
    gender %in% c("Other") ~ "Other"
  ))

count(BloodType, gender)
```

Check to see how many values `gender` has for each category (hint: use `count`). It's good practice to regularly check your data throughout the data wrangling process.

3.
```{r}
BloodType %>% count(gender)
```

4. Recode the `type` variable of the `BloodType` data to be consistent. Use `recode`.
```{r}
BloodType <- BloodType %>%
  mutate(type = recode(type,
    "o.-" = "O.-",
    "o.+" = "O.+"
  ))
BloodType
```

5. Check to see that `type` only has these possible values:
 "A.-","A.+", "AB.-", "AB.+", "B-","B+", "O.-", "O.+"

```{r}
BloodType %>%
  pull(type) %>%
  table(useNA = "ifany")
# or
BloodType %>% count(type)
```

6. Make a new tibble of `BloodType` called `Bloodtype_split` that splits the `type` variable into two called `blood_type` and  `Rhfactor`. Note: periods are special characters that generally are interpreted as wild cards thus we need `\\` to tell R that we want it to be interpreted as a period.

The following commented code within the chunk (starts with `#`) can be used as a guide to help you. Type your code below this commented code. 

```{r}
## ______________ <- ________ %>%
# ________(____, ____ = c(__________, __________), sep = "\\.")

BloodType_split <- BloodType %>% separate(type, into = c("blood_type", "Rhfactor"), sep = "\\.")
```

Bonus: How many observations are there for each `Rhfactor`:
```{r}
count(BloodType_split, Rhfactor)
```

Bonus: Filtering for patients with type O, how many had the infection?
```{r}
BloodType_split %>%
  filter(blood_type == "O") %>%
  count(infection)
```


