---
output:
  html_document:
    fig_caption: false
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)

schedule <- googlesheets4::read_sheet(
    "https://docs.google.com/spreadsheets/d/18Hcl_nOodBqjbwYX9RtusEIt-y76Uf3ZaATvwl5mPlI/edit?usp=sharing"
  )
```

# Day 1

## Objectives

Upon completion of this session, you will be able to do the following:

1. Begin interfacing with the RStudio IDE.
2. Locate basic help tools within the RStudio IDE.
3. Learn how to open and create RMarkdown files (.Rmd).
4. Understand R as a calculator.
5. Import common data types like csv, xls/xlsx, and txt files into RStudio.


## Lectures

Lecture recording links will be emailed directly to students.


## Material

```{r, results='asis', echo = FALSE, message=FALSE}
# Read in schedule from excel - easier to read than md :)
# https://docs.google.com/spreadsheets/d/18Hcl_nOodBqjbwYX9RtusEIt-y76Uf3ZaATvwl5mPlI/edit?usp=sharing
pander::pandoc.table(
  schedule  %>%
    mutate(across(
      .fns = ~ str_replace_all(., "\\./", "https://jhudatascience.org/intro_to_r/")
    )) %>%
    filter(
      row_number() >= which(str_detect(schedule$`**Day**`, "\\[Day 1\\]")) &
        row_number() < which(str_detect(schedule$`**Day**`, "\\[Day 2\\]")) - 1
    ),
  missing = "",
  split.tables = Inf,
  style = "rmarkdown"
)
``` 


## Homework

`r emo::ji("rotating_light")` **SURVEY**: [Best times for Office Hours](fixthislink)

`r emo::ji("rotating_light")` **SURVEY**: [First impressions feedback](fixthislink) 
