---
title: "Basic R Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



In this lab you can use the interactive console to explore or **Knit** the document. 
Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Cntr-Enter (Windows/Linux) in an R code chunk.


# Part 1

1. Create a new variable called `my.num` that contains 6 numbers.
```{r}
my.num <- c(5,4,7,8,12,14) # These can be any real numbers
```


2. Multiply `my.num` by 4.
```{r}
my.num * 4 
```


3. Create a second variable called `my.char` that contains 5 character strings.
```{r}
my.char <- c("banana", "garlic", "pickles", "hot chocolate", "7-Up") # Again,these can be anything
```


4. Combine the two variables `my.num` and `my.char` into a variable called `both`.
```{r}
both <- c(my.num, my.char)
```


5. What is the length of `both`?
```{r}
length(both)
```


6. What class is `both`?
```{r}
class(both)
```


# Part 2

7. Divide `both` by 3, what happens?
```{r, error=TRUE}
both / 3
```


8. Create a vector with elements 1 2 3 4 5 6 and call it `x`.
```{r}
x <- c(1, 2, 3, 4, 5, 6)
```


9. Create another vector with elements 10 20 30 40 50 60 and call it `y`.
```{r}
y <- c(10, 20, 30, 40, 50, 60)
```


10. The code `z <- y[1:5]` saves the first five items in `y` as the new variable `z`. Print `z`.
```{r}
z <- y[1:5]
print(z)
```


11. What happens when you add `x` and `z` together? How is this different from #10 ?
```{r}
x + z
# R "auto replicates" the shorter vector when the vectors aren't the same length - this is not what we want!
# In this case, it reads y as c(10, 20, 30, 40, 50, 10)!
```


12. Append the value 100 onto the vector `z` (hint: you can use the c() function).
```{r}
z <- c(z, 100)
```


13. Add `x` and `z` together.
```{r}
x + z
```


14. Multiply `x` and `z` together. How is this similar to the way R performs addition in #13 ?
```{r}
x * z
# R adds together the first element of each vector, then the second, and so on. This order also applies when multiplying!
```
