---
title: "Basic R Lab Key"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



In this lab you can use the interactive console to explore or **Knit** the document. 
Remember anything you type here can be "sent" to the console with Cmd-Enter (OS-X) or Cntr-Enter (Windows/Linux) in an R code chunk.


# Part 1

1. create a new variable called `my.num` that contains 6 numbers
```{r}
my.num = c(5,4,7,8,12,14)
```

2. mulitply `my.num` by 4
```{r}
my.num * 4 
```


3. create a second variable called `my.char` that contains 5 character strings
```{r}
my.char = c("Andrew", "John", "John", "Andrew","John")
```


4. combine the two variables `my.num` and `my.char` into a variable called `both`
```{r}
both = c(my.num, my.char)
```

5. what is the length of `both`?
```{r}
length(both)
```


6. what class is `both`?
```{r}
class(both)
```


# Part 2

7. Divide `both` by 3, what happens?
```{r, error=TRUE}
both / 3
```

8. create a vector with elements 1 2 3 4 5 6 and call it `x`
```{r}
x = c(1,2,3,4,5,6)
```


9. create another vector with elements 10 20 30 40 50 and call it `y`
```{r}
y =  c(10,20,30,40,50)
```


10. what happens if you try to add x and y together? why?
```{r}
x + y
```



11. append the value 60 onto the vector y (hint: you can use the c() function)
```{r}
y = c(y, 60)
```

12. add x and y together
```{r}
x + y
```


13. multiply x and y together. pay attention to how R performs operations on vectors of the same length.
```{r}
x * y
```


