---
title: '**KEY** Introduction to R: Homework 3'
output:
  html_document: default
---

### **Instructions**

1. Please submit your filled out `.Rmd` file to the assignment drop box on CoursePlus: https://courseplus.jhu.edu/core/index.cfm/go/db:assignment.submit/assignmentID/44235/coID/18000/

2. All assignments are due by Friday, July 1, 2022 at 11:59pm EST. Please reach out to us if you need help before this time!

```{r initiatePackages, message=FALSE}
## you can add more, or change...these are suggestions
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

### **Problem Set**

1\. Bring the dataset into R. The dataset is located at: https://jhudatascience.org/intro_to_r/data/mortality.csv. You can use the link, download it, or use whatever method you like for getting the file. Once you get the file, read the dataset in using `read_csv()` and assign it the name "mort".

```{r}

```

2\. Run the `colnames()` function to take a look at the dataset column names. You should see that there was originally no name for the first column and that R replaced it with "...1". Rename the first column of "mort" to "country" using the `rename()` function in `dplyr`. 

```{r}

```

3\. Select only the numeric type columns (`select()`). Then, create the variable "year" from column names by using the `colnames()` function to extract them. 

```{r}

```

4\. What is the `typeof()` for "year"? If it's not an integer, turn it into integer form with `as.integer()`.

```{r}

```

5\. Use the `pct_complete()` function in the `naniar` package to determine the percent missing data in "mort". You might need to load and install `naniar`!

```{r}

```

6\. Are there any countries that have a complete record in "mort" across all years? Just look at the output here, don't reassign it. **Hint:** look for complete records by dropping all NAs from the dataset using `drop_na()`.

```{r}

```

7\. Reshape the "complete" data to long form. 

- There should be a column for country ("country"), a column for year ("year"), and a column for the mortality value ("mortality"). 
- Use `pivot_longer()`. 
- You should pivot all columns except "country". 
- **Hint:** listing `!COLUMN` or `-COLUMN` means everything except COLUMN. 
- Assign the reshaped data to "long".

```{r}

```

8\. Bring an additional dataset into R. The dataset is tab-delimited and located at: https://jhudatascience.org/intro_to_r/data/country_pop.txt. You can use the link, download it, or use whatever method you like for getting the file. Once you get the file, read the dataset in using `read_tsv()` and assign it the name "pop".

```{r}

```

9\. Rename the second column in "pop" to "country" and the column "% of world population", to "percent". Use the `rename()` function. Don't forget to reassign the renamed data to "pop".

```{r}

```

10\. Sort the data in "pop" by "Population" from largest to smallest using `arrange()` and `desc()`. After sorting, `select()` "country" to create an one-column tibble of countries ordered by population. Assign this data the name "country_ordered".

```{r}

```

11\. Subset "long"  based on years 2000-2010, including 2000 and 2010 and call this "long_sub" using `&` or the `between()` function. Confirm your filtering worked by looking at the range of "year". If you're getting a strange error, make sure you created the "year" column in problem #7.

```{r}

```

12\. Further subset `long_sub`. You will filter for specific countries using `filter()` and the `%in%` operator. Only include countries in this list: `c("Venezuela", "Bahrain", "Estonia", "Iran", "Thailand", "Canada")`. Make sure to reassign to "long_sub".

```{r}

```

13\. Use `pivot_wider()` to turn the "year" column of "long_sub" into multiple columns, each representing a different year. Fill values (`values_from=`) with "mortality". Assign this pivoted dataset the name "mort_sub".

```{r}

```

14\. Using "country_ordered" and "mort_sub", `right_join()` the two datasets by "country". Use the pipe `%>%` to join this dataset to "pop", keeping only the data on the lefthand side of the join. Call this "joined".

```{r}

```

15\. The values in the table are percentages of the total population (not proportion). Create a new column called "mort_count" that estimates the total number of child deaths per year based on the total population. You can use any year, or an average of all of them, to make your calculation. Whatever you choose, justify your choice. Finally, `select()` only "country", "Population", and "mort_count" and view the data.

```{r}

```

### **Bonus Practice**

Please check back soon!
