---
title: "Data Summarization Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data used

Bike Lanes Dataset: BikeBaltimore is the Department of Transportation's bike program. 
The data is from http://data.baltimorecity.gov/Transportation/Bike-Lanes/xzfj-gyms

You can Download as a CSV in your current working directory.  Note its also available at: 	http://jhudatascience.org/intro_to_r/data/Bike_Lanes.csv 

```{r, echo = TRUE, message=FALSE, error = FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(jhur)

bike <- read_csv(file = "http://jhudatascience.org/intro_to_r/data/Bike_Lanes.csv")
```

or use 
```{r}
library(jhur)
bike <- read_bike()
```

# Part 1

1. How many bike "lanes" are currently in Baltimore?  You can assume each observation/row is a different bike "lane".  (hint: how do you get the number of rows of a data set? You can use `dim()` or `nrow()` or another function).

```{r}
nrow(bike)
dim(bike)
bike %>%
  nrow()
```

2. How many feet of bike "lanes" are currently in Baltimore, based on the `length` column? (use `sum()`)

```{r}
sum(bike$length)
sum(bike %>% pull(length))
```

3. Summarize the data to get the `max` and `min` of `length` using the `summarize` function.

```
# General format 
{data to use} %>% 
    summarize({summary column name} = {operator(source column)})
```

```{r}
bike %>% summarize(
  max = max(length),
  min = min(length)
)
```

**Bonus / Optional:**

A. Summarize the `bike` data to get the mean of `length` and `dateInstalled`. Do this three ways: 1) with `summarize`, 2) with `summarize` and `across`, and 3) with `colMeans()`.

```{r}
bike %>% summarize(
  mean_length = mean(length, na.rm = TRUE),
  mean_date = mean(dateInstalled, na.rm = TRUE)
)
bike %>%
  summarize(across(c(length, dateInstalled), ~ mean(.x, na.rm = TRUE)))

bike %>%
  select(length, dateInstalled) %>%
  colMeans()
```

You should have gotten a mean date sometime in the 1800s - that doesn't make much sense! Hypothesize why the average date is a date from before bike lanes were being built in Baltimore.

> There are probably some zeros or other incorrect low values in the data.

B. Filter any zeros out of `bike` `dateInstalled`. Use `filter()`. Assign this "cleaned" dataset object the name `bike_2`.

```{r}
bike_2 <- bike %>% filter(dateInstalled != 0)
```

* Sometimes it's not a good idea to drop data! You can also change any zeros in `bike` `dateInstalled` to `NA` using `mutate` and `ifelse()`. `ifelse()` in R works takes the same arguments as the "IF" function in Excel, and can come in handy here: `ifelse(condition, yes = VAL_IF_TRUE, no = VAL_IF_FALSE)`.

```{r}
bike_nas <- bike %>%
  mutate(dateInstalled = ifelse(dateInstalled == 0, yes = NA, no = dateInstalled))

# How to find NAs?
# is.na(bike$dateInstalled)
# !is.na(bike$dateInstalled)
```


# Part 2

4. How many types of bike lanes are there?   (Hints: `unique`, `table`, or `bike %>% count()` on the column named `type`).
   
```{r}
# So many ways!!

bike %>% count(type)

bike %>%
  pull(type) %>%
  table()

unique(bike %>% pull(type))
table(bike$type, useNA = "ifany")
unique(bike$type)
length(table(bike$type))
length(unique(bike$type))

is.na(unique(bike$type))

bike %>%
  group_by(type) %>%
  summarize(count = n())
```

5. Which bike lane`type` has (a) the most number of and (b) longest average bike lane length? (Hint: `group_by` and `summarize`). In your summarized output, make sure you call the new summarized average bike lane length variable (column name) "mean". In other words, the head of your output should look like:

```
# A tibble: 
  type            number_of_rows  mean
  <chr>                    <int> <dbl>
1 BIKE BOULEVARD              49  197.
...
```

```{r}
bike %>%
  group_by(type) %>%
  summarize(
    number_of_rows = n(),
    mean = mean(length)
  )
```

6. Take your code from the above question and do the following:

- Add another pipe (`%>%`)
- Add the `arrange()` to sort the output by the summarized column "mean".

```{r}
bike %>%
  group_by(type) %>%
  summarize(
    number_of_rows = n(),
    mean = mean(length)
  ) %>%
  arrange(mean)
```

7. Make a new column using `mutate`. This new column should be different for each year (`dateInstalled`), and indicate the total sum number of lanes (`numLanes`) built in that year. Call this new column `year_total` and make sure to reassign the dataset to `bike`. (hint: use `group_by` first)

```{r}
bike <- bike %>%
  group_by(dateInstalled) %>%
  mutate(year_total = sum(numLanes, na.rm = TRUE))
```

8. How does your data from above change if you also group by `type`? Ungroup your data when you are done.

```{r}
bike %>%
  group_by(dateInstalled, type) %>%
  mutate(year_total = sum(numLanes, na.rm = TRUE))

bike <- ungroup(bike)
```

**Bonus / Optional:**

C. Using `bike_2` from Part 1 above, what was the average bike lane length grouped by `dateInstalled`? Remove `NA`s with `na.rm = TRUE`.

```{r}
bike_2 %>%
  group_by(dateInstalled) %>%
  summarise(mean_of_the_bike = mean(length, na.rm = F))
```


# Part 3

9. Create a histogram for the `length` column in the `bike` dataset. Try playing with the `breaks =` argument.

```{r}
hist(x = pull(bike, length), breaks = 100)
```

10. Create a scatterplot with `dateInstalled` on the x axis and `length` on the y axis.

```{r}
plot(x = pull(bike, dateInstalled), y = pull(bike, length))
```

**Bonus / Optional:**

D. Does the plot from question 10 improve if you use the `bike_2` dataset?

```{r}
plot(x = pull(bike_2, dateInstalled), y = pull(bike_2, length))
```

> Yes!

E. What kind of plot would be better for showing the length by each year group? Make this plot.

> A boxplot would be more appropriate since year behaves more as a category than a continuous number.

```{r}
boxplot(pull(bike_2, length) ~ pull(bike_2, dateInstalled))
```
