---
title: "Homework 3"
author: "Untitled"
date: "January 3, 2019"
output: html_document
---

This is the second homework assignment for Intro to R. You must submit both the RMD and "knitted" HTML files are due before class on Friday (Day 5)


Use this dataset on infant mortality for the following questions: http://johnmuschelli.com/intro_to_r/data/indicatordeadkids35.csv


```{r initiatePackages, message=FALSE}
## you can add more, or change...these are suggestions
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

Questions

1.	Read the data using `read_csv()` and name it `mort`. Rename the first column to `country` using the `rename()` command in `dplyr`.  Create an object `year` variable by extracting column names (using `colnames()`) and make it to an integer `as.integer()`), excluding the first column either with string manipulations or bracket subsetting or subsetting with `is.na`. 

```{r, dataImport}

```

2.	Reshape the data so that there is a variable named `year` corresponding to `year` (key) and a column of the mortalities named `mortality` (value). Hint: use the `tidyr` package and its `gather()` function. Name the output `long` and make `year` a numeric variable. Hint: remember that -COLUMN_NAME removes that column, gather all the columns but country.  

```{r, question2}

```


3.	Read in this the tab-delim file: http://johnmuschelli.com/intro_to_r/data/country_pop.txt and call it `pop`, which contains population information on each country (hint: use `read_tsv()`).  Rename the second column to `"Country"` and the column `"% of world population"`, to `percent`

```{r, question3}


```

4.	Determine the population of each country in `pop` using `arrange()`. Get the order of the countries based on this (first is the highest population), and extract that column and call it `pop_levels`.  Make a variable in the `long` data set named `sorted` that is the `country` variable coded as a factor based on `pop_levels`.

```{r, question4}


```



5.	Subset `long` based on years 1975-2010, including 1975 and 2010 and call this `long_sub` using `&` or the `between()` function.  Subset the following countries using `dplyr::filter()` and the `%in%` operator on the sorted country factor (`sorted`) for `long_sub` with
`c("Venezuela", "Bahrain", "Estonia", "Iran", "Thailand", "Chile",  "Western Sahara", "Azerbaijan", "Argentina", "Haiti")` and reassign to `long_sub`. Lastly, remove missing rows for `mortality` using `filter()` and `is.na()`.

```{r, question5}


```


6.  Plotting: create "spaghetti"/line plots for the countries in `long_sub`, using different colors for different countries, using `sorted`. The x-axis should be `year`, and the y-axis should be `mortality`.  Make the plot using `qplot` and also `ggplot`.


7.  Bonus, load the `plotly` package (`library(plotly)`) and assign the plot from 6) to `g` and run `ggplotly(g)`.

```{r}


```


