---
title: "Homework 3 - Key"
author: "Introduction to R for Public Health Researchers"
output: html_document
---


Use this dataset on infant mortality for the following questions: http://johnmuschelli.com/intro_to_r/data/indicatordeadkids35.csv


```{r initiatePackages, message=FALSE}
## you can add more, or change...these are suggestions
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

Questions

1.	Read the data using `read_csv()` and name it `mort`. Rename the first column to `country` using the `rename()` command in `dplyr`.  Create an object `year` variable by extracting column names (using `colnames()`) and make it to an integer `as.integer()`), excluding the first column either with string manipulations or bracket subsetting or subsetting with `is.na`. 

```{r, dataImport}
mort = read_csv("http://johnmuschelli.com/intro_to_r/data/indicatordeadkids35.csv")
mort = mort %>% 
  rename(country = X1)
```

Using Bracket notation:
```{r}
year = colnames(mort)
year = year[-1]
year = as.integer(year)
```

or using `is.na`:
```{r}
year = colnames(mort)
year = as.integer(year)
year = year[ !is.na(year)]
```

or using string manipulations
```{r}
year = colnames(mort)
# start with a  digit
year = year[str_detect(year, "^\\d")]
year = as.integer(year)
```

2.	Reshape the data so that there is a variable named `year` corresponding to `year` (key) and a column of the mortalities named `mortality` (value). Hint: use the `tidyr` package and its `gather()` function. Name the output `long` and make `year` a numeric variable. Hint: remember that -COLUMN_NAME removes that column, gather all the columns but country.  

Let's create a new vector for the years:

```{r, question2}
# can use quotes
long = mort %>% 
  gather(key = "year", value = "mortality", -country)
# or without
long = mort %>% 
  gather(year, mortality, -country)

long = long %>% 
  mutate(year = as.numeric(year))
```


3.	Read in this the tab-delim file: http://johnmuschelli.com/intro_to_r/data/country_pop.txt and call it `pop`, which contains population information on each country (hint: use `read_tsv()`).  Rename the second column to `"Country"` and the column `"% of world population"`, to `percent`

```{r, question3}
pop = read_tsv("http://johnmuschelli.com/intro_to_r/data/country_pop.txt")
pop = pop %>% 
  rename(Country = `Country (or dependent territory)`,
         percent = `% of world population`)
```

4.	Determine the population of each country in `pop` using `arrange()`. Get the order of the countries based on this (first is the highest population), and extract that column and call it `pop_levels`.  Make a variable in the `long` data set named `sorted` that is the `country` variable coded as a factor based on `pop_levels`.

```{r, question4}
pop = pop %>% 
  arrange(desc(Population))
# this is sorted !
pop_levels = pop$Country
long = long %>% 
  mutate(sorted = factor(country, levels = pop_levels))
```

As an aside, we should do some cleaning and checking before doing this, as we see not all the countries in the `long` data set exactly match those in the `pop` data set:
```{r}
# you would want to clean these up in practice
# setdiff shows the "set difference"
setdiff(long$country, pop$Country)

# some are now set to missing (as factors do)
sum(is.na(long$country))
sum(is.na(long$sorted))
```


5.	Subset `long` based on years 1975-2010, including 1975 and 2010 and call this `long_sub` using `&` or the `between()` function.  Subset the following countries using `dplyr::filter()` and the `%in%` operator on the sorted country factor ("sorted") for `long_sub` with
`c("Venezuela", "Bahrain", "Estonia", "Iran", "Thailand", "Chile",  "Western Sahara", "Azerbaijan", "Argentina", "Haiti")`. Lastly, remove missing rows for mortality using `filter()` and `is.na()`.

Subsetting long:
```{r, question5}
long_sub = long %>% 
  filter(year >= 1975 & year <= 2010)
range(long_sub$year)
```

There is a function `between` that helps us with this for shorthand

```{r, question5}
long_sub = long %>% 
  filter(between(year, 1975, 2010))
range(long_sub$year)

```
