---
title: "Introduction to R"
author: "Introduction to R for Public Health Researchers"
output:
  ioslides_presentation:
    css: ../styles.css
    widescreen: yes
--- 

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(comment = "")
```

## Welcome to class!

1. Introductions
2. Class overview
3. Getting R up and running


## About Us

**John Muschelli** 

Assistant Scientist, Department of Biostatistics

PhD in Biostatistics, ScM in Biostatistics

Email: jmusche1@jhu.edu

## About Us

**Andrew Jaffe**

Lead Investigator, Lieber Institute for Brain Development

Assistant Professor, Department of Mental Health, JHSPH

PhD in Epidemiology, MHS in Bioinformatics

Email: ajaffe@jhu.edu


## What is R?

- R is a language and environment for statistical computing and graphics

- R is the open source implementation of the S language, which was developed by Bell laboratories

- R is both open source and open development

(source: http://www.r-project.org/)

## Why R?

* Powerful and flexible

* Free (open source)

* Extensive add-on software (packages)

* Designed for statistical computing

* High level language

## Why not R?

* Fairly steep learning curve

    * "Programming" oriented
    
    * Minimal interface
    
* Little centralized support, relies on online community and package developers

* Annoying to update

* Slower, and more memory intensive, than the more traditional programming languages (C, Java, Perl, Python)


## Introductions

What do you hope to get out of the class?

Why else to use R?

## Course Website

http://johnmuschelli.com/intro_to_r

Materials will be uploaded the night before class

## Learning Objectives

- Reading data into R
- Recoding and manipulating data
- Writing R functions and using add-on packages
- Making exploratory plots
- Understanding basic programming syntax
- Performing basic statistical tests

<!-- ## Course Format -->

<!-- * 3 modules per class session, each approximately 1 hour -->
<!--     * "Interactive" Lecture with RStudio + slides -->
<!--     * Lab/Practical experience -->

<!-- ## Grading -->

<!-- 1. Attendance/Participation: 20% -->
<!-- 2. Nightly Homework: 3 x 15% -->
<!-- 3. Final "Project": 35%  -->

<!-- ## Grading -->

<!-- * **Homework 1**: Due Tuesday by class -->

<!-- * **Homework 2**: Due Thursday by class -->

<!-- * **Homework 3**: Due Friday by class -->

<!-- * **Project**: Due 2 weeks after class ends -->

## Installing R

* Install the latest version from: [http://cran.r-project.org/](http://cran.r-project.org/ )

* [Install RStudio](https://www.rstudio.com/products/rstudio/download/)

## Collection of R packages

We have an R package called jhur that will make sure all the packages are installed.

You can just copy and paste the below code into your console - we'll explain what it all means in the next day or two

```{r, packageInstall, eval=FALSE}
install.packages("remotes")
remotes::install_github("muschellij2/jhur")
```

Note it may take ~5-10 minutes to run.

## Useful (+Free) Resources

* R for Data Science: http://r4ds.had.co.nz/ 
* Various "Cheat Sheets": https://www.rstudio.com/resources/cheatsheets/
* DataCamp http://www.datacamp.com
* R reference card: http://cran.r-project.org/doc/contrib/Short-refcard.pdf
* UCLA Institute for Digital Research and Education: http://www.ats.ucla.edu/stat/r/
* Quick R: http://statmethods.net/

## Website

[Website](http://johnmuschelli.com/intro_to_r/index.html)
