---
title: "Factors Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

0. Load all the libraries we will use in this lab. 

```{r message=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
```

1. Create a function that takes one argument, a vector, and returns the sum of the vector squared. Call it "sum_squared". Test your function on the vector `c(2,7,21,30,90)` - you should get the answer 22500.

```
# General format
NEW_FUNCTION <- function(x, y) x + y ..
```

```{r}
x <- c(2, 7, 21, 30, 90)

sum_squared <- function(x) sum(x)^2
sum_squared(x)

sum_squared <- function(x) {
  out <- sum(x)^2
  return(out)
}
sum_squared(x)
```

2. Create a function that takes two arguments, (1) a vector and (2) a numeric value. This function tests whether the number (2) is contained within the vector (1). Call it "has_n". Test your function on the vector `c(2,7,21,30,90)` and number `21` - you should get the answer TRUE.

```{r}
x <- c(2, 7, 21, 30, 90)
n <- 21

has_n <- function(x, n) n %in% x
has_n(x, n)
```

3. Amend the function "has_n" from #2 so that it takes a default value of `21` for the numeric argument.

```{r}
x <- c(2, 7, 21, 30, 90)
n <- 21

has_n <- function(x, n = 21) n %in% x
has_n(x)
```

4. Read in the SARS-CoV-2 Vaccination data from http://jhudatascience.org/intro_to_r/data/USA_covid19_vaccinations.csv. Assign the data the name "vacc".

```{r message=FALSE}
vacc <- read_csv("http://jhudatascience.org/intro_to_r/data/USA_covid19_vaccinations.csv")
# If downloaded
# vacc <- read_csv("USA_covid19_vaccinations.csv")
```

5. We want to get some summary statistics on the Moderna vaccines. Use `across` inside `summarize` to get the sum total number vaccine doses for any variable containing the word "Moderna" or starting with "Total". **Hint**: use `matches()` AND `starts_with()` to select the right columns inside `across`. Keep in mind that this includes the United States as a whole and so is not totally accurate!

```
# General format
data %>%
  summarize(across(
    .cols = {vector or tidyselect},
    .fns = {some function},
    {additional arguments}
  ))
```

```{r}
vacc %>%
  summarize(across(
    .cols = matches("Moderna") & starts_with("Total"),
    .fns = sum
  ))
vacc %>%
  summarize(across(
    .cols = matches("Moderna") & starts_with("Total"),
    .fns = sum,
    na.rm = TRUE
  ))
```

6. Use `across` and `mutate` to convert all columns containing the word "Percent" into proportions (i.e., divide that value by 100). **Hint**: use `matches()` to select the right columns. Use a "function on the fly" to divide by 100. It will also be easier to check your work if you `select()` columns that match "Percent".

```{r}
vacc %>%
  mutate(across(
    .cols = matches("Percent"),
    .fns = function(x) x / 100
  )) %>%
  select(matches("Percent"))
```

7. Use `across` and `mutate` to convert all columns starting with the word "Total" into a binary variable: TRUE if the value is greater than 10,000,000 and FALSE if less than or equal to 10,000,000. **Hint**: use `starts_with()` to select the columns starting with "Total". Use a "function on the fly" to do a logical test if the value is greater than 10,000,000.

```{r}
vacc %>%
  mutate(across(
    .cols = starts_with("Total"),
    .fns = function(x) x > 10000000
  ))
```

**Bonus / Extra practice**

Take your code from #7 and assign it to the variable `vacc_dat`. 

- use `filter()` to drop any rows where "United States" appears in `State/Territory/Federal Entity`. Make sure to reassign this to `vacc_dat`.
- Create a ggplot boxplot (`geom_boxplot()`) where (1) the x-axis is `Total Doses Delivered` and (2) the y-axis is `Percent of fully vaccinated people with booster doses`.
- You change the `labs()` layer so that the x-axis is "Total Doses Delivered: Greater than 10,000,000"

```{r}
vacc_dat <-
  vacc %>%
  mutate(across(
    .cols = starts_with("Total"),
    .fns = function(x) x > 10000000
  )) %>%
  filter(`State/Territory/Federal Entity` != "United States")

ggplot(vacc_dat) +
  geom_boxplot(aes(
    x = `Total Doses Delivered`,
    y = `Percent of fully vaccinated people with booster doses`
  )) +
  labs(x = "Total Doses Delivered: Greater than 10,000,000")
```
