---
title: "Data Cleaning Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Data used

Bike Lanes Dataset: BikeBaltimore is the Department of Transportation's bike program. 
The data is from http://data.baltimorecity.gov/Transportation/Bike-Lanes/xzfj-gyms

You can Download as a CSV in your current working directory.  Note its also available at: 	http://jhudatascience.org/intro_to_r/data/Bike_Lanes.csv 

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(jhur)
library(tidyverse)
library(broom)
library(naniar)

bike = read_csv(
  "http://jhudatascience.org/intro_to_r/data/Bike_Lanes.csv")
```

# Part 1

1.  Count the number of rows of the bike data and count the number of complete cases of the bike data.  Use `sum` and `complete.cases`.
```{r}

```

2.  Create a data set called `namat` which is equal to `is.na(bike)`.  
What is the class of `namat`?  Run rowSums and colSums on `namat.`
These represent the number of missing values in the rows and columns of
`bike`.  Don't print `rowSums`, but do a table of the `rowSums`
Try `nabike = bike %>% mutate_all(is.na)`.  Then try `nabike %>% summarise_all(sum)`.  What do you get?  
```{r}

```

3.  Filter rows of bike that are NOT missing the `route` variable, assign
this to the object `have_route.`  Do a table of the `subType` using `table`, 
including the missing `subType`s.  Get the same frequency distribution
using `group_by(subType)` and `tally().`
```{r}

```

4. Use `naniar` to make a visual of the amount of data missing for each variable of `bike` (use `naniar::gg_miss_var()`).

```{r}

```

# Part 2

**New Data set**

Download the "Real Property Taxes" Data from my website (via OpenBaltimore):
http://jhudatascience.org/intro_to_r/data/Real_Property_Taxes.csv.gz
note you don't need to unzip it to read it into R



5. Read the Property Tax data into R and call it the variable `tax`
```{r}

```

6. How many addresses pay property taxes? 
```{r}

```

7. What is the total city tax paid?  
You need to remove the `$` from the `CityTax` variable
then you need to make it numeric.   Try `str_replace`, but 
$ is "special"  and you need `fixed()` around it.
```{r}

```

8. a. how many observations/properties are in each ward?
```{r}

```

b. what is the maximum amount still due in each ward?  different summarization (`max`)
```{r}

```

10. Subset the data to only retain those houses that are principal residences. 
which command subsets rows? Filter or select?

How many such houses are there?
```{r}

```


**New Data Set**

Read in the Salary FY2015 dataset
```{r}

sal = read_csv("http://jhudatascience.org/intro_to_r/data/Baltimore_City_Employee_Salaries_FY2015.csv")
```

11. Make an object called health.sal using the salaries data set, with only agencies of those with `"fire"` (or any forms), if any, in the name (remember `fixed( ignore_case = TRUE)` will ignore cases).
```{r}

```

12. Make a data set called `trans` which contains only agencies that contain "TRANS".
```{r}

```

13. What is/are the profession(s) of people who have `"abra"` in their name for Baltimore's Salaries?  Case should be ignored
```{r}

```


14. Create a smaller dataset that only includes the Police Department,  Fire Department and Sheriff's Office.  Use the Agency variable with string matching. Call this `emer`.  How many employees are in this new dataset?
```{r}

```

15. Create a variable called `dept` in the emer data set.
`dept = str_extract(Agency, ".*(ment|ice)")`.  E.g. want to extract all characters
up until `ment` or `ice` (we can group in `regex` using parentheses) and then discard the rest.
```{r}

```



16 (Bonus). Convert the 'LotSize' variable to a numeric square feet variable in the tax data set. Some tips: 
a) 1 acre = 43560 square feet
b) The hyphens represent inches (not decimals)

```{r}

```

