---
title: "Data Cleaning"
output:
  ioslides_presentation:
    css: ../styles.css
    widescreen: yes
---
 

```{r, echo = FALSE, message = FALSE}
library(knitr)
opts_chunk$set(comment = "")
library(readr)
suppressPackageStartupMessages(library(dplyr))
library(tidyverse)
```

<!-- # Before Cleaning - Subsetting with Brackets -->

<!-- ## Select specific elements using an index -->

<!-- Often you only want to look at subsets of a data set at any given time.  Elements of an R object are selected using the brackets (`[` and `]`). -->

<!-- For example, `x` is a vector of numbers and we can select the second element of `x` using the brackets and an index (2): -->

<!-- ```{r} -->
<!-- x = c(1, 4, 2, 8, 10) -->
<!-- x[2] -->
<!-- ``` -->

<!-- <!-- `dplyr`: --> -->

<!-- <!-- ```{r} --> -->
<!-- <!-- nth(x, n = 2) --> -->
<!-- <!-- ``` --> -->

<!-- ## Select specific elements using an index -->

<!-- We can select the fifth or second AND fifth elements below: -->

<!-- ```{r} -->
<!-- x = c(1, 2, 4, 8, 10) -->
<!-- x[5] -->
<!-- x[c(2,5)] -->
<!-- ``` -->


<!-- ```{r, error=TRUE} -->
<!-- nth(x, n = c(2, 5)) # nth only returns one number -->
<!-- ``` -->


<!-- ## Subsetting by deletion of entries -->

<!-- You can put a minus (`-`) before integers inside brackets to remove these indices from the data. -->

<!-- ```{r negativeIndex} -->
<!-- x[-2] # all but the second -->
<!-- ``` -->

<!-- Note that you have to be careful with this syntax when dropping more than 1 element: -->

<!-- ```{r negativeIndex2} -->
<!-- x[-c(1,2,3)] # drop first 3 -->
<!-- # x[-1:3] # shorthand. R sees as -1 to 3 -->
<!-- x[-(1:3)] # needs parentheses -->
<!-- ``` -->

<!-- ## Select specific elements using logical operators -->

<!-- What about selecting rows based on the values of two variables? We use logical statements.  Here we select only elements of `x` greater than 2: -->

<!-- ```{r} -->
<!-- x -->
<!-- x > 2 -->
<!-- x[ x > 2 ] -->
<!-- ``` -->

<!-- ## Select specific elements using logical operators -->
<!-- You can have multiple logical conditions using the following: -->

<!-- * `&` : AND -->
<!-- * `|` : OR -->

<!-- ```{r} -->
<!-- x[ x > 2 & x < 5 ] -->
<!-- x[ x > 5 | x == 2 ] -->
<!-- ``` -->


<!-- ## which function -->

<!-- The `which` functions takes in logical vectors and returns the index for the elements where the logical value is `TRUE`.   -->

<!-- ```{r} -->
<!-- which(x > 5 | x == 2) # returns index -->
<!-- x[ which(x > 5 | x == 2) ] -->
<!-- x[ x > 5 | x == 2 ] -->
<!-- ``` -->

## Data Cleaning

In general, data cleaning is a process of investigating your data for inaccuracies, or recoding it in a way that makes it more manageable.

MOST IMPORTANT RULE - LOOK AT YOUR DATA!


# Dealing with Missing Data

## Missing data types
One of the most important aspects of data cleaning is missing values.  

Types of "missing" data:

* `NA` - general missing data
* `NaN` - stands for "**N**ot **a** **N**umber", happens when you do 0/0.
* `Inf` and `-Inf` - Infinity, happens when you take a positive number (or negative number) by 0.

## Finding Missing data

Each missing data type has a function that returns `TRUE` if the data is missing:

* `NA` - `is.na`
* `NaN` - `is.nan`
* `Inf` and `-Inf` - `is.infinite`
* `is.finite` returns `FALSE` for all missing data and `TRUE` for non-missing

## Useful checking functions

- `is.na` - is `TRUE` if the data is `FALSE` otherwise
- `!` - negation (NOT) 
    - if `is.na(x)` is `TRUE`, then `!is.na(x)` is `FALSE`

- `all` takes in a `logical` and will be `TRUE` if ALL are `TRUE`
    - `all(!is.na(x))` - are all values of `x` NOT NA
- `any` will be `TRUE` if ANY are true
    - `any(is.na(x))` - do we have any `NA`'s in `x`?

```{r isna}
A = c(1, 2, 4, NA)
B = c(1, 2, 3, 4)
any(is.na(A)) # are there any NAs - YES/TRUE
any(is.na(B)) # are there any NAs- NO/FALSE
```


```{r all}
A = c(1, 2, 4, NA)
B = c(1, 2, 3, 4)
all(is.na(A)) # are all values NA - NO/FALSE
all(!is.na(B)) # are all values not NA - YES/TRUE
```

## Complete.cases

`complete.cases` - returns `TRUE` if EVERY value of a row is NOT `NA`
    - very stringent condition 
    - `FALSE` missing one value (even if not important)
   

```{r}
complete.cases(mtcars)
```

## naniar
Sometimes you need to look at lots of data though... the `naniar` package is a good option.

```{r, error=FALSE}
#install.packageS(naniar)
library(naniar)
x = c(0, NA, 2, 3, 4, -0.5, 0.2)
naniar::pct_complete(x)
```

## Naniar plots
```{r}
naniar::gg_miss_var(iris)
```

## Missing Data with Logicals

Recall that mathematical operations with `NA` will result in `NA`s

This is also true for logicals.  This is a good thing. The NA data could be `> 2` or not, we don't know, 
so `R` says there is no `TRUE` or `FALSE`, so that is missing.
```{r}
x = c(0, NA, 2, 3, 4, -0.5, 0.2)
x > 2
```

<!-- ## Missing Data with Logicals -->

<!-- What to do?  What if we want if `x > 2` and `x` isn't `NA`? -->
<!-- Don't do `x != NA`, do `x > 2` and `x` is NOT `NA`: -->

<!-- ```{r} -->
<!-- x > 2 & x != NA -->
<!-- x > 2 & !is.na(x) -->
<!-- ``` -->

<!-- ## Missing Data with Logicals -->

<!-- What about seeing if a value is equal to multiple values?  You can do `(x == 1 | x == 2) & !is.na(x)`, but that is not efficient.   -->

<!-- ```{r} -->
<!-- (x == 0 | x == 2) # has NA -->
<!-- (x == 0 | x == 2) & !is.na(x) # No NA -->
<!-- ``` -->

<!-- what to do? -->

## filter() and missing data

Be careful with missing data using subsetting: 

filter() removes missing values by default.
To keep them need to add is.na():

```{r}
x # looks like the 1st and 3rd element should be TRUE
x %in% c(0, 2) # uh oh - not good!
x %in% c(0, 2) | is.na(x) # do this

```

## tidyr::drop_na
This function will drop rows with **any** missing data in **any** column.

```{r}
df <-tibble(x = c(0, NA, 2, 0.2), 
            y = c(NA, 1, 6, NA))
df
drop_na(df)
```

 

<!-- ## Missing Data with Operations -->

<!-- Similarly with logicals, operations/arithmetic with `NA` will result in `NA`s: -->

<!-- ```{r} -->
<!-- x = c(0, NA, 2, 3, 4, -0.5, 0.2) -->
<!-- x + 2 -->
<!-- x * 2 -->
<!-- ``` -->


# Missing data with Tables and Tabulations

## Useful checking functions

  - `table(x, useNA = "ifany")` - will have row NA if there are NA values
  - `table(x, useNA = "always")` - will have row NA even if there are no NA values

Here we will use `table` to make tabulations of the data.   Look at `?table` to see options for missing data.

```{r table}
z = c("A", "B", "A", "B")
table(z, useNA = "ifany")
table(z, useNA = "always")
```


<!-- ## Creating One-way Tables  -->

<!-- `useNA = "ifany"` will not have NA in table heading if no `NA`: -->
<!-- ```{r onetab_ifany} -->
<!-- table(c(0, 1, 2, 3, 2, 3, 3, 2, 2, 3),  -->
<!--         useNA = "ifany") -->
<!-- tibble(x = c(0, 1, 2, 3, 2, 3, 3, 2, 2, 3)) %>%  count(x) -->
<!-- ``` -->

<!-- ## Creating One-way Tables  -->

<!-- You can set `useNA = "always"` to have it always have a column for `NA` -->
<!-- ```{r onetab} -->
<!-- table(c(0, 1, 2, 3, 2, 3, 3, 2, 2, 3),  -->
<!--         useNA = "always") -->
<!-- ``` -->

<!-- ## Tables with Factors  -->

<!-- If you use a `factor`, all levels will be given even if no exist!  -->
<!--   - (May be wanted or not): -->

<!-- ```{r onetab_fact, error = TRUE} -->
<!-- fac = factor(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3), -->
<!--              levels = 1:4) -->
<!-- tab = table(fac) -->
<!-- tab -->
<!-- tab[ tab > 0 ] -->
<!-- ``` -->


<!-- ## Factors cont -->
<!-- ```{r} -->
<!-- tibble(x = fac) %>% count(x) -->
<!-- ``` -->


<!-- ## Finding Row or Column Totals -->

<!-- `margin.table` finds the marginal sums of the table.  `margin` is 1 for rows, 2 for columns in general in `R`.  Here is the column sums of the table: -->

<!-- ```{r margin} -->
<!-- tab -->
<!-- margin.table(tab, 2) -->
<!-- ``` -->

<!-- ## Proportion Tables  -->

<!-- `prop.table` finds the marginal proportions of the table. Think of it dividing the table by it's respective marginal totals.  If `margin` not set, divides by overall total. -->

<!-- ```{r table2} -->
<!-- prop.table(tab) -->
<!-- prop.table(tab,1) * 100 -->
<!-- ``` -->

## Lab Part 1

[Website](http://jhudatascience.org/intro_to_r/index.html)



<!-- ```{r readSal_csv, echo= FALSE, eval = TRUE} -->
<!-- Sal = read_csv("https://jhudatascience.org/intro_to_r/data/Baltimore_City_Employee_Salaries_FY2015.csv", col_types = cols( -->
<!--   name = col_character(), -->
<!--   JobTitle = col_character(), -->
<!--   AgencyID = col_character(), -->
<!--   Agency = col_character(), -->
<!--   HireDate = col_character(), -->
<!--   AnnualSalary = col_character(), -->
<!--   GrossPay = col_character() -->
<!-- )) -->
<!-- Sal = rename(Sal, Name = name) -->
<!-- ``` -->

<!-- ## Checking for logical conditions  -->

<!-- * `any()` - checks if there are any `TRUE`s -->
<!-- * `all()` - checks if ALL are true -->

<!-- ```{r isna} -->
<!-- head(Sal,2) -->
<!-- any(is.na(Sal$Name)) # are there any NAs? -->
<!-- ``` -->

# Recoding Variables

## Example of Recoding

Let's say gender was coded as Male, M, m, Female, F, f.  Using Excel to find all of these would be a matter of filtering and changing all by hand or using if statements. This can be hectic!

In `dplyr` you can use the `recode` function:

```{r, eval = FALSE}
data = data %>% 
  mutate(gender = recode(gender, M = "Male", m = "Male", Man = "Male"))
```

Or use `ifelse()` or `case_when()`.

<!-- # ```{r, eval = FALSE} -->
<!-- # data %>%  -->
<!-- #   mutate(gender = ifelse(gender %in% c("Male", "M", "m"), -->
<!-- #                          "Male", gender)) -->
<!-- # ``` -->



<!-- ## Example of Cleaning: more complicated -->

<!-- Sometimes though, it's not so simple.  That's where functions that find patterns come in very useful. -->

<!-- ```{r gender, echo=FALSE} -->
<!-- set.seed(4) # random sample below - make sure same every time -->
<!-- gender <- sample(c("Male", "mAle", "MaLe", "M", "MALE", "Ma", "FeMAle", "F", "Woman", "Man", "Fm", "FEMALE"), 1000, replace = TRUE) -->
<!-- ``` -->

<!-- ```{r gentab} -->
<!-- table(gender) -->
<!-- ``` -->


<!-- ## Example of Cleaning: more complicated -->

<!-- ```{r gender2, echo=FALSE} -->
<!-- gender = gender %>%  -->
<!--   tolower %>%  -->
<!--   recode(m = "Male", f = "Female", ma = "Male", -->
<!--          woman = "Female", man = "Male") -->
<!-- ``` -->

<!-- ```{r gentab2} -->
<!-- table(gender) -->
<!-- ``` -->

## Separating columns based on a separator 

* From `tidyr`, you can split a data set into multiple columns:

```{r separate_df}
df = tibble(x = c("I really", "like writing", "R code programs"))
```

```{r}
df %>% separate(x, into = c("first", "second", "third"))
```

## Separating columns based on a separator 
You can specify the separator with `sep`.
* `extra = "merge"` will not drop data.  

```{r}
df %>% separate(x, into = c("first", "second", "third"), 
                extra = "merge", sep = " ")
```

## Uniting columns based on a separator 

* From `tidyr`, you can unite:

```{r unite_df}
df = tibble(id = rep(1:5, 3), visit = rep(1:3, each = 5))
head(df, 4)
```


```{r}
df_united <- df %>% unite(col = "unique_id", id, visit, sep = "_")
head(df_united, 4)
```



# Strings functions
## Splitting/Find/Replace and Regular Expressions

* R can do much more than find exact matches for a whole string
* Like Perl and other languages, it can use regular expressions.
* What are regular expressions?
    * Ways to search for specific strings 
    * Can be very complicated or simple
    * Highly Useful - think "Find" on steroids


## A bit on Regular Expressions

* http://www.regular-expressions.info/reference.html
* They can use to match a large number of strings in one statement
* `.` matches any single character
* `*` means repeat as many (even if 0) more times the last character
* `?` makes the last thing optional
* `^` matches start of vector `^a` - starts with "a"
* `$` matches end of vector `b$` - ends with "b"

## The `stringr` package

The `stringr` package:

* Makes string manipulation more intuitive
* Has a standard format for most functions
    * the first argument is a string like first argument is a `data.frame` in `dplyr`
* We will not cover `grep` or `gsub` - base R functions
    - are used on forums for answers
* Almost all functions start with `str_*`


## Let's look at modifier for `stringr`

`?modifiers`

* `fixed` - match everything exactly
<!-- * `regex` - default - uses **reg**ular **exp**ressions -->
* `ignore_case` is an option to not have to use `tolower`

## Substring and String Splitting

* `str_sub(x, start, end)` - substrings from position start to position end
* `str_split(string, pattern)` - splits strings up - returns list!

```{r str_split_orig}
library(stringr)
x <- c("I really like writing R code")
df = tibble(x = c("I really", "like writing", "R code programs"))
y <- unlist(str_split(x, " "))
y
length(y)
```

## Using a fixed expression

One example case is when you want to split on a period "`.`".  In regular expressions `.` means **ANY** character, so

```{r}
str_split("I.like.strings", ".")
str_split("I.like.strings", fixed("."))
```

<!-- ## Let's extract from `y` -->

<!-- ```{r stsplit2} -->
<!-- y[[2]] -->
<!-- # sapply(y, dplyr::first) # on the fly -->
<!-- # sapply(y, nth, 2) # on the fly -->
<!-- # sapply(y, last) # on the fly -->
<!-- ``` -->




<!-- ## Separating columns based on a separator  -->

<!-- * From `tidyr`, you can split a data set into multiple columns: -->

<!-- ```{r separate_df2_for_easy} -->
<!-- df = tibble(x = c("I really", "like writing", "R code programs")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df %>% separate(x, into = c("first", "second")) -->
<!-- ``` -->


<!-- ## Separating columns based on a separator  -->

<!-- * `extra = "merge"` will not drop data.  Also, you can specify the separator -->

<!-- ```{r separate_df2} -->
<!-- df = tibble(x = c("I really", "like. _writing R. But not", "R code programs")) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- df %>% separate(x, into = c("first", "second", "third"), extra = "merge") -->
<!-- ``` -->






## 'Find' functions: `stringr`

`str_detect`, `str_subset`, `str_replace`, and `str_replace_all` search for matches to argument pattern within each element of a character vector: they differ in the format of and amount of detail in the results. 

* `str_detect` - returns `TRUE` if `pattern` is found
* `str_subset` - returns only the strings which pattern were detected
    * convenient wrapper around `x[str_detect(x, pattern)]`
* `str_extract` - returns only strings which pattern were detected, but ONLY the pattern
* `str_replace` - replaces `pattern` with `replacement` the first time
* `str_replace_all` - replaces `pattern` with `replacement` as many times matched

## Download Salary FY2014 Data

From https://data.baltimorecity.gov/City-Government/Baltimore-City-Employee-Salaries-FY2015/nsfe-bg53, from  https://data.baltimorecity.gov/api/views/nsfe-bg53/rows.csv

Read the CSV into R `Sal`:

```{r readSal, echo = TRUE, eval = TRUE}
Sal = jhur::read_salaries() # or
Sal = read_csv("https://jhudatascience.org/intro_to_r/data/Baltimore_City_Employee_Salaries_FY2015.csv")
Sal = rename(Sal, Name = name)
Sal
```


## 'Find' functions: Finding Logicals

These are the indices where the pattern match occurs:

```{r RawlMatch_log}
head(str_detect(Sal$Name, "Rawlings"))
```

<!-- ## 'Find' functions: Finding Indices -->

<!-- These are the indices where the pattern match occurs: -->

<!-- ```{r RawlMatch} -->
<!-- which(str_detect(Sal$Name, "Rawlings")) -->
<!-- ``` -->

<!-- ## Showing difference in `str_extract` -->

<!-- `str_extract` extracts just the matched string -->

<!-- ```{r ggrep2} -->
<!-- ss = str_extract(Sal$Name, "Rawling") -->
<!-- head(ss) -->
<!-- ss[ !is.na(ss)] -->
<!-- ``` -->

<!-- str_subset(Sal$Name, "Rawlings") -->


## 'Find' functions: finding values: `stringr` 

```{r}
Sal %>% filter(str_detect(Name, "Rawlings"))
```



<!-- ## Showing differnce in `str_extract` and  `str_extract_all` -->

<!-- `str_extract_all` extracts all the matched strings -->
<!-- - `\\d` searches for DIGITS/numbers -->

<!-- ```{r } -->
<!-- head(str_extract(Sal$AgencyID, "\\d")) -->
<!-- head(str_extract_all(Sal$AgencyID, "\\d"), 2) -->
<!-- ``` -->

## Showing differnce in `str_replace` and  `str_replace_all`

`str_replace` replaces only the first instance.

```{r}
head(Sal$Name, 2)
head(str_replace(Sal$Name, "a", "j"),2)
```

`str_replace` replaces all instances.
```{r}
head(str_replace_all(Sal$Name, "a", "j"), 2)
```


<!-- ## Replace -->

<!-- Let's say we wanted to sort the data set by Annual Salary: -->

<!-- ```{r classSal} -->
<!-- class(Sal$AnnualSalary) -->
<!-- ``` -->

<!-- ```{r destringSal} -->
<!-- head(Sal$AnnualSalary, 4) -->
<!-- head(as.numeric(Sal$AnnualSalary), 4) -->
<!-- ``` -->

<!-- R didn't like the `$` so it thought turned them all to `NA`. -->

<!-- ## Replacing and substituting  -->

<!-- Now we can replace the `$` with nothing (used `fixed("$")` because `$` means ending): -->

<!-- ```{r orderSal} -->
<!-- Sal = Sal %>% mutate( -->
<!--   AnnualSalary = str_replace(AnnualSalary, fixed("$"), ""), -->
<!--   AnnualSalary = as.numeric(AnnualSalary) -->
<!--   ) %>%  -->
<!--   arrange(desc(AnnualSalary)) -->
<!-- ``` -->



## Pasting strings with `paste` and `paste0`

Paste can be very useful for joining vectors together:

```{r Paste}
paste("Visit", 1:5, sep = "_")
paste("Visit", 1:5, sep = "_", collapse = "_")
paste("To", "is going be the ", "we go to the store!", sep = "day ")
# and paste0 can be even simpler see ?paste0 
paste0("Visit",1:5) # no space!
```
## Lab Part 2

[lab part 2](http://jhudatascience.org/intro_to_r/Data_Cleaning/lab/Data_Cleaning_Lab.Rmd)

[Website](http://jhudatascience.org/intro_to_r/index.html)



# Extra Slides

## Using Regular Expressions
* Look for any name that starts with:
    * Payne at the beginning, 
    * Leonard and then an S
    * Spence then capital C

```{r grepstar_stringr}
head(str_subset( Sal$Name, "^Payne.*"), 3)
```

```{r grepstar2_stringr}
head(str_subset( Sal$Name, "Leonard.?S"))
head(str_subset( Sal$Name, "Spence.*C.*"))
```


<!-- ## Uniting columns based on a separator  -->

<!-- * From `tidyr`, you can unite: -->

<!-- ```{r unite_df2} -->
<!-- df = tibble(id = rep(1:5, 3), visit = rep(1:3, each = 5)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df %>% unite(col = "unique_id", id, visit, sep = "_", remove = FALSE) -->
<!-- ``` -->

<!-- ## Paste Depicting How Collapse Works -->

<!-- ```{r Paste2} -->
<!-- paste(1:5) -->
<!-- paste(1:5, collapse = " ") -->
<!-- ``` -->

<!-- ## Useful String Functions -->

<!-- Useful String functions -->

<!-- * `toupper()`, `tolower()` - uppercase or lowercase your data: -->
<!-- * `str_trim()` (in the `stringr` package) or `trimws` in base  -->
<!--     - will trim whitespace -->
<!-- * `nchar` - get the number of characters in a string -->



<!-- ## Sorting characters -->

<!-- - `sort` - reorders the data - characters work, but not correctly -->
<!-- - `rank` - gives the rank of the data - ties are split -->
<!-- - `order` - gives the indices, if subset, would give the data sorted -->
<!--     - `x[order(x)]` is the same as sorting -->

<!-- ```{r orderrank} -->
<!-- sort(c("1", "2", "10")) #  not sort correctly (order simply ranks the data) -->
<!-- order(c("1", "2", "10")) -->
<!-- x = rnorm(10) -->
<!-- x[1] = x[2] # create a tie -->
<!-- rank(x) -->
<!-- ``` -->






















# Comparison of `stringr` to base R - not covered






# Splitting Strings

## Substringing

`stringr`

<!-- * `str_sub(x, start, end)` - substrings from position start to position end -->
* `str_split(string, pattern)` - splits strings up - returns list!


## Splitting String:

In `stringr`, `str_split` splits a vector on a string into a `list`

```{r}
x <- c("I really", "like writing", "R code programs")
y <- stringr::str_split(x, pattern =  " ") # returns a list
y
```



## `str_extract`

`str_extract` extracts  matched strings
- `\\d` searches for DIGITS/numbers

```{r}
head(Sal$AgencyID)
head(str_extract(Sal$AgencyID, "\\d"))
```



<!-- ## 'Find' functions: base R -->

<!-- `grep`: `grep`, `grepl`, `regexpr` and `gregexpr` search for matches to argument pattern within each element of a character vector: they differ in the format of and amount of detail in the results.  -->

<!-- `grep(pattern, x, fixed=FALSE)`, where: -->

<!-- * pattern = character string containing a regular expression to be matched in the given character vector. -->

<!-- * x = a character vector where matches are sought, or an object which can be coerced by as.character to a character vector. -->

<!-- * If fixed=TRUE, it will do exact matching for the phrase anywhere in the vector (regular find) -->


## 'Find' functions: stringr compared to base R

Base R does not use these functions.  Here is a "translator" of the `stringr` function to base R functions

* `str_detect` - similar to `grepl` (return logical)
* `grep(value = FALSE)` is similar to `which(str_detect())`
* `str_subset` - similar to `grep(value = TRUE)` - return value of matched
* `str_replace` - similar to `sub` - replace one time 
* `str_replace_all` - similar to `gsub` - replace many times

## Important Comparisons

Base R:

* Argument order is `(pattern, x)`
* Uses option `(fixed = TRUE)`

`stringr`

* Argument order is `(string, pattern)` aka `(x, pattern)`
* Uses function `fixed(pattern)`

## 'Find' functions: Finding Indices

These are the indices where the pattern match occurs:

```{r}
grep("Rawlings",Sal$Name)
which(grepl("Rawlings", Sal$Name))
which(str_detect(Sal$Name, "Rawlings"))
```

## 'Find' functions: Finding Logicals

These are the indices where the pattern match occurs:

```{r}
head(grepl("Rawlings",Sal$Name))
head(str_detect(Sal$Name, "Rawlings"))
```


## 'Find' functions: finding values, base R 

```{r}
grep("Rawlings",Sal$Name,value=TRUE)
Sal[grep("Rawlings",Sal$Name),]
```



## Showing differnce in `str_extract`

`str_extract` extracts just the matched string

```{r}
ss = str_extract(Sal$Name, "Rawling")
head(ss)
ss[ !is.na(ss)]
```

## Showing differnce in `str_extract` and  `str_extract_all`

`str_extract_all` extracts all the matched strings

```{r}
head(str_extract(Sal$AgencyID, "\\d"))
head(str_extract_all(Sal$AgencyID, "\\d"), 2)
```



## Using Regular Expressions
* Look for any name that starts with:
    * Payne at the beginning, 
    * Leonard and then an S
    * Spence then capital C

```{r}
head(grep("^Payne.*", x = Sal$Name, value = TRUE), 3)
```

```{r}
head(grep("Leonard.?S", x = Sal$Name, value = TRUE))
head(grep("Spence.*C.*", x = Sal$Name, value = TRUE))
```



## Using Regular Expressions: `stringr`
```{r}
head(str_subset( Sal$Name, "^Payne.*"), 3)
```

```{r}
head(str_subset( Sal$Name, "Leonard.?S"))
head(str_subset( Sal$Name, "Spence.*C.*"))
```

## Replace

Let's say we wanted to sort the data set by Annual Salary:

```{r}
class(Sal$AnnualSalary)
```

```{r}
sort(c("1", "2", "10")) #  not sort correctly (order simply ranks the data)
order(c("1", "2", "10"))
```

## Replace

So we must change the annual pay into a numeric:
```{r}
head(Sal$AnnualSalary, 4)
head(as.numeric(Sal$AnnualSalary), 4)
```

R didn't like the `$` so it thought turned them all to `NA`.

`sub()` and `gsub()` can do the replacing part in base R.

## Replacing and subbing 

Now we can replace the `$` with nothing (used `fixed=TRUE` because `$` means ending):

```{r}
Sal$AnnualSalary <- as.numeric(gsub(pattern = "$", replacement="", 
                              Sal$AnnualSalary, fixed=TRUE))
Sal <- Sal[order(Sal$AnnualSalary, decreasing=TRUE), ] 
Sal[1:5, c("Name", "AnnualSalary", "JobTitle")]
```

## Replacing and subbing: `stringr` 

We can do the same thing (with 2 piping operations!) in dplyr
```{r}
dplyr_sal = Sal
dplyr_sal = dplyr_sal %>% mutate( 
  AnnualSalary = AnnualSalary %>%
    str_replace(
      fixed("$"), 
      "") %>%
    as.numeric) %>%
  arrange(desc(AnnualSalary))
check_Sal = Sal
rownames(check_Sal) = NULL
all.equal(check_Sal, dplyr_sal)
```


## Website

[Website](../index.html)

# Extra slides

## Creating Two-way Tables

A two-way table.  If you pass in 2 vectors, `table` creates a 2-dimensional table.

```{r}
tab <- table(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3),
             c(0, 1, 2, 3, 2, 3, 3, 4, 4, 3),
              useNA = "always")
tab
```

## Creating Two-way Tables

```{r}
tab_df = tibble(x = c(0, 1, 2, 3, 2, 3, 3, 2,2, 3),
             y = c(0, 1, 2, 3, 2, 3, 3, 4, 4, 3))
tab_df %>% count(x, y)
```

## Creating Two-way Tables 

```{r}
tab_df %>% 
  count(x, y) %>% 
  group_by(x) %>% mutate(pct_x = n / sum(n))
```

## Creating Two-way Tables 

```{r, message = FALSE}
library(scales)
tab_df %>% 
  count(x, y) %>% 
  group_by(x) %>% mutate(pct_x = percent(n / sum(n)))
```
