---
title: "Esquisse Data Visualization Lab Key"
output: html_document
---

First, make sure you install and load the `esquisse` package using `install.packages` and `library`:

```{r, eval = FALSE}
install.packages("esquisse")
install.packages("ggplot2")
```

```{r, comment = FALSE}
library(esquisse)
library(ggplot2)
library(jhur)
```

1.  Try creating a plot using the `Orange` data that automatically comes with R using the `esquisse` package.

-   Drag and drop the `age` variable to be plotted on the x-axis.
-   Drag and drop the `circumference` variable to be plotted on the y-axis.
-   What happens when you drag `tree` to the facet region of the esquisse GUI?
-   What happens if you move this to color?
-   Copy and paste the code from both options into the chunk below:

```{r}
# esquisser(Orange)
ggplot(Orange) +
  aes(x = age, y = circumference) +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  theme_minimal() +
  facet_wrap(vars(Tree))

ggplot(Orange) +
  aes(x = age, y = circumference, colour = Tree) +
  geom_point(shape = "circle", size = 1.5) +
  scale_color_hue(direction = 1) +
  theme_minimal()
```

2.  Click where it says "point" (may say "auto" depending on how you did the last question) on the far left side and change the plot to a different type of plot. Copy and paste the code into the chunk below. Close Esquisse and run the chunk below to generate a ggplot.

```{r}
ggplot(Orange) +
  aes(x = age, y = circumference, colour = Tree) +
  geom_line(size = 0.5) +
  scale_color_hue(direction = 1) +
  theme_minimal()
```

**Bonus / optional:**

Launch Esquisse on any selection of the following datasets we have worked with before and explore!

```{r}
yts <- read_yts()
tb <- read_tb()
bike <- read_bike()
circ <- read_circulator()
vacc <- read_csv("http://jhudatascience.org/intro_to_r/data/USA_covid19_vaccinations.csv")

# esquisser(vacc)
```
