---
title: "Esquisse Data Visualization Lab"
output: html_document
---


First, make sure you install and load the `esquisse` and `ggplot2` packages using `install.packages` and `library`:

```{r, eval = FALSE}
# install.packages("esquisse")
# install.packages("ggplot2")
```

```{r, comment = FALSE}
library(esquisse)
library(ggplot2)
```


Now try creating a plot using the `Orange` data that automatically comes with R using the `esquisser` function of the `esquisse` package. Drag and drop the `age` variable to be plotted on the x-axis. Drag and drop the `circumfrance` variable to be plotted on the y-axis. What happens when you drag `tree` to the facet region of the esquisse GUI? What happens if you move this to color? Copy and paste the code from both options into the chunk below:

```{r}
# esquisser(Orange) delete this line of code when rendering but keep the code that you copy and paste here from running esquisser to create the two plots.
```

Can you think of another way to plot this data?

Bonus: Click where it says "point" (may say "auto" depending on how you did the last question) on the far left side and change the plot to a different type of plot? Copy and paste the code into the chunk below.

```{r}

```
