---
title: "Syllabus"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
```

Epi/Biostats  Winter Institute - Introduction to R for Public Health Researchers
140.604.73

Class Website: http://jhudatascience.org/intro_to_r/   
CoursePlus: https://courseplus.jhu.edu/core/index.cfm/go/syl:syl.public.view/coid/16733/

**Zoom link will be emailed to students.**

Day/Time: Jan 10 - 21: 8:30AM-11:50AM on Zoom

Instructors: Carrie Wright (cwrigh60@jhu.edu), Ava Hoffman (ava.hoffman@jhu.edu), and Candace Savonen(csavone1@jhu.edu) 

TAs:  Grant Schumock (gschumo1@jhmi.edu) and Qier Meng(qmeng11@jhmi.edu)


**Communication will mainly occur through Slack and we will email you about how to connect to slack.**


Overview: This course will provide "hands-on" training for learning how to analyze data in the R statistical software package. We will cover data input/output, data management and manipulation, and how to make useful and informative graphics
 
Course Format: Each class will consist of 2 or 3 hour-long modules: each module features a lecture and an R programming lab, where student apply the skills taught in the modules to real data. 

By the end of the course, students should be comfortable:

- Reading data into R
- Recoding and manipulating data
- Using R add-on packages
- Making exploratory plots
- Performing basic statistical tests
- Understanding basic programming syntax
- Creating reproducible R documents

# Tentative Schedule:

## Day Overview

| Time (EST)        | Content   |
|-------------------|-----------|
| 8:30am - 9:30am   | Session 1 |
| 9:30am - 9:40am   | Break     |
| 9:40am - 10:40am  | Session 2 |
| 10:40am - 10:50am | Break     |
| 10:50am - 11:50am | Session 3 |


## Day 1

- Introduction 
- RStudio 
- Reproducible Research 


## Day 2
- Basic R: Variables/Objects in R
- Data Input/Output 


## Day 3

- Subsetting Data
- Homework 2

## Day 4

- Summarization 
- Data Classes

## Day 5

 
- Data Cleaning



## Day 6
- Data Manipulation
- Homework 3

## Day 7

-  Data Visualization

## Day 8

- Statistics
- Work on Final Project


## Day 9

- Functions
- Good code practices 
- Work on Final Project

```{r, child=c(paste(here::here(), "/Syllabus/_child_schedule.Rmd", sep = ""))}
```

```{r, results='asis', echo = FALSE}
# Read in schedule from excel - easier to read than md :)
# The mutate_all creates an absolute url to linked documents since there are word and pdf rendered versions of the syllabus
pander::pandoc.table(
  readxl::read_excel("schedule.xlsx")  %>% 
    mutate_all(funs(str_replace_all(., "\\./", "https://jhudatascience.org/intro_to_r/"))),
  missing = "",
  split.tables = Inf,
  style = "rmarkdown"
)
```

```{r, child=c(paste(here::here(), "/Syllabus/_child_assignments.Rmd", sep = ""))}
```
