---
title: "Syllabus"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
```

Epi/Biostats  Winter Institute - Introduction to R for Public Health Researchers
140.604.73

Class Website: http://jhudatascience.org/intro_to_R_class/   
CoursePlus: https://courseplus.jhu.edu/core/index.cfm/go/syl:syl.public.view/coid/16733/

**Zoom link will be emailed to students.**

Day/Time: Jan 10 - 21: 8:30AM-11:50AM on Zoom

Instructors: Carrie Wright (cwrigh60@jhu.edu), Ava Hoffman (ava.hoffman@jhu.edu), and Candace Savonen (csavone1@jhu.edu) 

TAs:  Grant Schumock (gschumo1@jhmi.edu) and Qier Meng(qmeng11@jhmi.edu)


**Communication will mainly occur through Slack and we will email you about how to connect to slack.**


Overview: This course will provide "hands-on" training for learning how to analyze data in the R statistical software package. We will cover data input/output, data management and manipulation, and how to make useful and informative graphics
 
Course Format: Each class will consist of 2 or 3 hour-long modules: each module features a lecture and an R programming lab, where student apply the skills taught in the modules to real data. 

By the end of the course, students should be comfortable:

- Reading data into R
- Recoding and manipulating data
- Using R add-on packages
- Making exploratory plots
- Performing basic statistical tests
- Understanding basic programming syntax
- Creating reproducible R documents

# Tentative Schedule:

## Day Overview

| Time (EST)        | Content   |
|-------------------|-----------|
| 8:30am - 9:30am   | Session 1 |
| 9:30am - 9:40am   | Break     |
| 9:40am - 10:40am  | Session 2 |
| 10:40am - 10:50am | Break     |
| 10:50am - 11:50am | Session 3 |

```{r, child="_child_schedule.Rmd"}
```

```{r, results='asis', echo = FALSE, message=FALSE, warning = FALSE}
# The str_replace_all creates an absolute url to linked documents since there are word and pdf rendered versions of the syllabus
source("../scripts/utils.R")
pander::pandoc.table(
  read_markdown("_schedule_table.Rmd")  %>%
    mutate(across(
      .fns = ~ str_replace_all(., "\\./", "https://jhudatascience.org/intro_to_R_class/")
    )),
  missing = "",
  split.tables = Inf,
  style = "rmarkdown"
)
```

```{r, child="_child_assignments.Rmd"}
```

# Module Details:

## [Day 0 (before the first session)](https://jhudatascience.org/intro_to_R_class/Syllabus/module_details/day0.html)
 - Homework 1

## [Day 1](https://jhudatascience.org/intro_to_R_class/Syllabus/module_details/day1.html)

- Introduction 
- RStudio 
- Reproducible Research 


## [Day 2](https://jhudatascience.org/intro_to_R_class/Syllabus/module_details/day2.html)

- Basic R: Variables/Objects in R
- Data Input/Output 


## [Day 3](https://jhudatascience.org/intro_to_R_class/Syllabus/module_details/day3.html)

- Subsetting Data
- Discuss Homework 2 - work on Homework 2 in class if there is time


## [Day 4](https://jhudatascience.org/intro_to_R_class/Syllabus/module_details/day4.html)

- Summarization 
- Data Classes


## [Day 5](https://jhudatascience.org/intro_to_R_class/Syllabus/module_details/day5.html)


- Data Cleaning
- Discuss Homework 3 - work on Homework 3 in class if there is time


## [Day 6](https://jhudatascience.org/intro_to_R_class/Syllabus/module_details/day6.html)

- Data Manipulation


## [Day 7](https://jhudatascience.org/intro_to_R_class/Syllabus/module_details/day7.html)

- Data Visualization
- Factors

## [Day 8](https://jhudatascience.org/intro_to_R_class/Syllabus/module_details/day8.html)

- Statistics
- Discuss and work on Final Project


## [Day 9](https://jhudatascience.org/intro_to_R_class/Syllabus/module_details/day9.html)

- Functions
- Work on Final Project

