---
title: "Data Classes"
author: "Introduction to R for Public Health Researchers"
output:
  ioslides_presentation:
    css: ../styles.css
    widescreen: yes
  beamer_presentation: default
---

```{r, echo = FALSE, message=FALSE}
# library(dplyr)
suppressPackageStartupMessages(library(dplyr))
library(readr)
```


## Data Classes:

* One dimensional classes ('vectors'):
    * Character: strings or individual characters, quoted
    * Numeric: any real number(s)
    * Integer: any integer(s)/whole numbers
    * Factor: categorical/qualitative variables
    * Logical: variables composed of TRUE or FALSE
    * Date/POSIXct: represents calendar dates and times

## Character and numeric

We have already covered `character` and `numeric` classes.

```{r numChar}
class(c("Andrew", "Jaffe"))
class(c(1, 4, 7))
```

## Integer

`Integer` is a special subset of `numeric` that contains only whole numbers

A sequence of numbers is an example of the integer class

```{r seq}
x = seq(from = 1, to = 5) # seq() is a function
x
class(x)
```

## Integer

The colon `:` is a shortcut for making sequences of numbers

It makes consecutive integer sequence from `[num1]` to `[num2]` by 1


```{r seqShort}
1:5
```

## Logical

`logical` is a class that only has two possible elements: `TRUE` and `FALSE`

```{r logical1}
x = c(TRUE, FALSE, TRUE, TRUE, FALSE)
class(x)
is.numeric(c("Andrew", "Jaffe"))
is.character(c("Andrew", "Jaffe"))
```

## Logical

Note that `logical` elements are NOT in quotes. 
```{r logical2}
z = c("TRUE", "FALSE", "TRUE", "FALSE")
class(z)
as.logical(z)
```

Bonus: `sum()` and `mean()` work on `logical` vectors - they return the total and proportion of `TRUE` elements, respectively.

```{r logical_z}
sum(as.logical(z))
```

## General Class Information

There are two useful functions associated with practically all R classes, which relate to logically checking the underlying class (`is.CLASS_()`) and coercing between classes (`as.CLASS_()`).

```{r logical_coercion}
is.numeric(c("Andrew", "Jaffe"))
is.character(c("Andrew", "Jaffe"))
```

## General Class Information

There are two useful functions associated with practically all R classes, which relate to logically checking the underlying class (`is.CLASS_()`) and coercing between classes (`as.CLASS_()`).

```{r logical_coercion2}
as.character(c(1, 4, 7))
as.numeric(c("Andrew", "Jaffe"))
```


## Factors

A `factor` is a special `character` vector where the elements have pre-defined groups or 'levels'. You can think of these as qualitative or categorical variables:

```{r factor1}
x = factor(c("boy", "girl", "girl", "boy", "girl"))
x 
class(x)
```

Note that levels are, by default, in alphanumerical order.

## Factors

Factors are used to represent categorical data, and can also be used for ordinal data (ie categories have an intrinsic ordering)

Note that R reads in character strings as factors by default in functions like `read.csv()` (but not `read_csv`)

'The function factor is used to encode a vector as a factor (the terms 'category' and 'enumerated type' are also used for factors). If argument ordered is TRUE, the factor levels are assumed to be ordered.'

```
factor(x = character(), levels, labels = levels,
       exclude = NA, ordered = is.ordered(x))
```

## Factors

Suppose we have a vector of case-control status

```{r factor2}
cc = factor(c("case","case","case",
        "control","control","control"))
cc
```
We can reset the levels using the `levels` function, but this is **bad** and can cause problems.  You should do this using the `levels` argument in the `factor()`
```{r}
levels(cc) = c("control","case")
cc
```

## Factors

Note that the levels are alphabetically ordered by default. We can also specify the levels within the factor call

```{r factor_cc_again}
casecontrol = c("case","case","case","control",
          "control","control")
factor(casecontrol, levels = c("control","case") )
factor(casecontrol, levels = c("control","case"), 
       ordered=TRUE)
```

## Factors

Factors can be converted to `numeric` or `character` very easily

```{r factor3}
x = factor(casecontrol,
        levels = c("control","case") )
as.character(x)
as.numeric(x)
```

## Factors

However, you need to be careful modifying the labels of existing factors, as its quite easy to alter the meaning of the underlying data. 

```{r factorCheck}
xCopy = x
levels(xCopy) = c("case", "control") # wrong way
xCopy        
as.character(xCopy) # labels switched
as.numeric(xCopy)
```

## Creating categorical variables

The `rep()` ["repeat"] function is useful for creating new variables 

```{r rep1}
bg = rep(c("boy","girl"),each=50)
head(bg)
bg2 = rep(c("boy","girl"),times=50)
head(bg2)
length(bg) == length(bg2)
```

## Creating categorical variables

One frequently-used tool is creating categorical variables out of continuous variables, like generating quantiles of a specific continuously measured variable. 

A general function for creating new variables based on existing variables is the `ifelse()` function, which "returns a value with the same shape as test which is filled with elements selected from either yes or no depending on whether the element of test is `TRUE` or `FALSE`."

```
ifelse(test, yes, no)

# test: an object which can be coerced
    to logical mode.
# yes: return values for true elements of test.
# no: return values for false elements of test.
```

## Charm City Circulator data

Please download the Charm City Circulator data:

http://johnmuschelli.com/intro_to_r/data/Charm_City_Circulator_Ridership.csv

```{r, message = FALSE}
# paste/paste0 combines strings/character
circ = read_csv(
  paste0("http://johnmuschelli.com/intro_to_r/data",
         "/Charm_City_Circulator_Ridership.csv"))
```

## Creating categorical variables

For example, we can create a new variable that records whether daily ridership on the Circulator was above 10,000.

```{r ifelse1}
hi_rider = ifelse(circ$daily > 10000, "high", "low")
hi_rider = factor(hi_rider, levels = c("low","high"))
head(hi_rider)
table(hi_rider)
```

## Creating categorical variables

You can also nest `ifelse()` within itself to create 3 levels of a variable.

```{r ifelse2}
riderLevels = ifelse(circ$daily < 10000, "low", 
                  ifelse(circ$daily > 20000,
                  "high", "med"))
riderLevels = factor(riderLevels, 
              levels = c("low","med","high"))
head(riderLevels)
table(riderLevels)
```

## Creating categorical variables

However, it's much easier to use `cut()` to create categorical variables from continuous variables. 

'cut divides the range of x into intervals and codes the values in x according to which interval they fall. The leftmost interval corresponds to level one, the next leftmost to level two and so on.'

```
cut(x, breaks, labels = NULL, include.lowest = FALSE, 
  right = TRUE, dig.lab = 3,
  ordered_result = FALSE, ...)
```

## Creating categorical variables

`x`: a numeric vector which is to be converted to a factor by cutting.

`breaks`: either a numeric vector of two or more unique cut points or a single number (greater than or equal to 2) giving the number of intervals into which x is to be cut.

`labels`: labels for the levels of the resulting category. By default, labels are constructed using "(a,b]" interval notation. If `labels = FALSE`, simple integer codes are returned instead of a factor.


## Creating categorical variables


```{r show_cut}
riderLevels2 = cut(
  circ$daily, 
  breaks = c(min(circ$daily, na.rm = TRUE),
             10000, 
             20000,
             max(circ$daily, na.rm = TRUE)),
  labels = c("low", "med", "high"), # one less than breaks
  include.lowest = TRUE)
head(riderLevels2)
table(riderLevels2, riderLevels)
```

## Cut

Now that we know more about factors, `cut()` will make more sense:

```{r cut1}
x = 1:100
cx = cut(x, breaks = c(0,10,25,50,100))
head(cx)  
table(cx)
```

## Cut

We can also leave off the labels

```{r cut2}
cx = cut(x, breaks = c(0,10,25,50,100), labels = FALSE)
head(cx)  
table(cx)
```

## Cut

Note that you have to specify the endpoints of the data, otherwise some of the categories will not be created

```{r cut3}
cx = cut(x, breaks = c(10,25,50), labels = FALSE)
head(cx)  
table(cx)
table(cx, useNA = "ifany")
```

## Date

You can convert date-like strings in the `Date` class (http://www.statmethods.net/input/dates.html for more info)

```{r date}
head(sort(circ$date))
# creating a date for sorting
circ$newDate <- as.Date(circ$date, "%m/%d/%Y")
head(circ$newDate)
range(circ$newDate)
```

## Date 

However, the `lubridate` package is much easier for generating explicit dates:

```{r, message=FALSE}
library(lubridate) # great for dates!
circ = mutate(circ, newDate2 = mdy(date))
head(circ$newDate2)
range(circ$newDate2) # gives you the range of the data
```

## POSIXct

The `POSIXct` class is like a more general date format (with hours, minutes, seconds).

```{r}
theTime = Sys.time()
theTime
class(theTime)
theTime + as.period(20, unit = "minutes") # the future
```

## Date 

However, the `lubridate` package is much easier for generating explicit dates:

```{r, message=FALSE}
circ = circ %>% 
  group_by(day) %>% 
  mutate(first_date = first(newDate2),
         diff_from_first = difftime( # time1 - time2
           time1 = newDate2, time2 = first_date)) 
head(circ$diff_from_first, 10)
units(circ$diff_from_first) = "days"
head(circ$diff_from_first, 10)
```


## Data Classes:

* Two dimensional classes:
    * `data.frame`: traditional 'Excel' spreadsheets
        * Each column can have a different class, from above
    * Matrix: two-dimensional data, composed of rows and columns. Unlike data frames, the entire matrix is composed of one R class, e.g. all numeric or all characters.
    
## Matrices

```{r matrix}
n = 1:9 
n
mat = matrix(n, nrow = 3)
mat
```

## Matrix (and Data frame) Functions

These are in addition to the previous useful vector functions:

* `nrow()` displays the number of rows of a matrix or data frame
* `ncol()` displays the number of columns 
* `dim()` displays a vector of length 2: # rows, # columns
* `colnames()` displays the column names (if any) and `rownames()` displays the row names (if any)

## Data Selection

Matrices have two "slots" you can use to select data, which represent rows and columns, that are separated by a comma, so the syntax is `matrix[row,column]`. Note you cannot use `dplyr` functions on matrices.

```{r subset3}
mat[1, 1] # individual entry: row 1, column 1
mat[1, ] # first row
mat[, 1] # first columns
```

## Data Selection

Note that the class of the returned object is no longer a matrix

```{r subset4}
class(mat[1, ])
class(mat[, 1])
```

## Data Frames

To review, the `data.frame`/`tbl_df` are the other two dimensional variable classes. 

Again, data frames are like matrices, but each column is a vector that can have its own class. So some columns might be `character` and others might be `numeric`, while others maybe a `factor`.

## Lists

* One other data type that is the most generic are `lists`.
* Can be created using list()
* Can hold vectors, strings, matrices, models, list of other list, lists upon lists!
* Can reference data using $ (if the elements are named), or using [], or [[]]

```{r makeList, comment="", prompt=TRUE}
mylist <- list(letters=c("A", "b", "c"), 
        numbers=1:3, matrix(1:25, ncol=5))
```

## List Structure
```{r Lists, comment="", prompt=TRUE}
head(mylist)
```

## List referencing
```{r Listsref1, comment="", prompt=TRUE}
mylist[1] # returns a list
mylist["letters"] # returns a list
```

## List referencing
  
```{r Listsrefvec, comment="", prompt=TRUE}  
mylist[[1]] # returns the vector 'letters'
mylist$letters # returns vector
mylist[["letters"]] # returns the vector 'letters'
```

## List referencing

You can also select multiple lists with the single brackets. 

```{r Listsref2, comment="", prompt=TRUE}
mylist[1:2] # returns a list
```

## List referencing

You can also select down several levels of a list at once

```{r Listsref3, comment="", prompt=TRUE}
mylist$letters[1]
mylist[[2]][1]
mylist[[3]][1:2,1:2]
```


## Website

[Website](../index.html)
