---
title: "Data Class Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Part 1

Load all the libraries we will use in this lab. 

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(jhur)
```

1. Create some data to work with by running the following code chunk.
```{r}
set.seed(1234)

int_vect <- rep(seq(from = 1, to = 10), times = 3)
rand_vect <- sample(x = 1:30, size = 30, replace = TRUE)
TF_vect <- rep(c(TRUE, TRUE, FALSE), times = 10)
TF_vect2 <- rep(c("TRUE", "TRUE", "FALSE"), times = 10)
```

2. Determine the class of each of these new objects.
```{r}

```

3. Create a tibble combining these vectors together called `vect_data` using the following code.
```{r}
vect_data <- tibble(int_vect, rand_vect, TF_vect, TF_vect2)
```

4. Take a look at 5 random rows using the `slice_sample()` function. Try this a few times to see how the results change.
```{r}

```

5. Take a look at 5 random rows using the `slice_sample()` function again but this time with `set.seed(1234)` as the first line of the chunk. Try this a few times to see the results. (Don't forget to not copy the backticks.)
```{r}

```

6. Check to see if the `TF_vect` is logical. Check to see if `TF_vect2` is logical. Why are the results what they are?
```{r}


```


Bonus. Use `mutate()` function to create a new variable in `vect_data` named `type_fact` that is of class `factor` made from the `int_vect` variable. Take a look at the data and observe how the class is different for the new variable compared to `int_vect`.

```{r}



```



# Part 2

1. Read in the Charm City Circulator data using `read_circulator()` function from `jhur` package using the code supplied in the chunk. 

Use the `str()` function to take a look at the data and learn about the column types.

```{r}
circ <- read_circulator()

```

2. Use the `mutate()` function to create a new column named `date_formatted` that is of `Date` class. The new variable is created from `date` column. Hint: use `mdy()` function.

```{r}


```

3. Move the `date_formatted` variable to be before `date` using the `relocate` function. Take a look at the data using `glimpse()`. Note the difference between `date` and `date_formatted` columns.  

```{r}



```

4. Create a list data object called `classes_data` that combines the `vect_data` from the first part of the lab and `circ`. Use `glimpse()` to look at the data.

```{r}



```

Bonus: Use `range()` function on `date_formatted` variable to display the range of dates in the data set. How does this compare to that of `date`? Why? (Hint: use the pull function first to pull the values.)

```{r}


```
 

Bonus: Use the `group_by` function on `day` variable with the `summarize()` function, to display the number of `orangeBoardings` observations with each day (hint: `use sum()`). Which day had the most boardings? Is this true for the other routes (`purpleBoardings`, `greenBoardings`, `bannerBoardings`)?

```{r}



```
 


