---
title: "Data Class Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Part 1

Load all the libraries we will use in this lab. 

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(jhur)
```


1\. Create some data to work with by running the following code chunk.

```{r}
set.seed(1234)

int_vect <- rep(seq(from = 1, to = 10), times = 3)
rand_vect <- sample(x = 1:30, size = 30, replace = TRUE)
TF_vect <- rep(c(TRUE, TRUE, FALSE), times = 10)
TF_vect2 <- rep(c("TRUE", "TRUE", "FALSE"), times = 10)
```


2\. Determine the class of each of these new objects.

```{r}
class(int_vect) # [1] "integer"
class(rand_vect) # [1] "integer"
class(TF_vect) # [1] "logical"
class(TF_vect2) # [1] "character"
```


3\. Create a tibble combining these vectors together called `vect_data` using the following code.

```{r}
vect_data <- tibble(int_vect, rand_vect, TF_vect, TF_vect2)
```


4\. Take a look at 5 random rows using the `slice_sample()` function. Try this a few times to see how the results change.

```{r}
slice_sample(vect_data, n = 5)
slice_sample(vect_data, n = 5)
slice_sample(vect_data, n = 5)
```


5\. Take a look at 5 random rows using the `slice_sample()` function again but this time with `set.seed(1234)` as the first line of the chunk. Try this a few times to see the results. (Don't forget to not copy the backticks.)

```{r}
set.seed(1234)
slice_sample(vect_data, n = 5)
slice_sample(vect_data, n = 5)
slice_sample(vect_data, n = 5)
```


6\. Check to see if the `TF_vect` is logical. Check to see if `TF_vect2` is logical. Why are the results what they are?

```{r}
is.logical(TF_vect)
is.logical(TF_vect2)
```

> Logical vectors do not have quotes around `TRUE` and `FALSE` values.


**Bonus / Extra practice**: Use `mutate()` function to create a new variable in `vect_data` named `type_fact` that is of class `factor` made from the `int_vect` variable. Take a look at the data and observe how the class is different for the new variable compared to `int_vect`. 

```{r}
vect_data <- mutate(vect_data, type_fact = factor(int_vect))
vect_data
```


# Part 2

7\. Read in the Charm City Circulator data using `read_circulator()` function from `jhur` package using the code supplied in the chunk. 

- Use the `str()` function to take a look at the data and learn about the column types.

```{r}
circ <- read_circulator()
str(circ)
```


8\. Use the `mutate()` function to create a new column named `date_formatted` that is of `Date` class. The new variable is created from `date` column. Hint: use `mdy()` function.

```{r}
circ <- mutate(circ, date_formatted = mdy(date))
```


9\. Move the `date_formatted` variable to be before `date` using the `relocate` function. Take a look at the data using `glimpse()`. Note the difference between `date` and `date_formatted` columns.  


```{r}
circ <- circ %>% relocate(date_formatted, .before = date)
glimpse(circ)
```


10\. Create a list data object called `classes_data` that combines the `vect_data` from the first part of the lab and `circ`. Use `glimpse()` to look at the data.

```{r}
classes_data <- list(vect_data, circ)
glimpse(classes_data)
```


**Bonus / Extra practice**: Use `range()` function on `date_formatted` variable to display the range of dates in the data set. How does this compare to that of `date`? Why? (Hint: use the pull function first to pull the values.)

```{r}
pull(circ, date_formatted) %>% range()
pull(circ, date) %>% range()
```

> The max of `pull(circ, date) %>% range()` is numerical not based on date.


**Bonus / Extra practice**: Use the `group_by` function on `day` variable with the `summarize()` function, to display the number of `orangeBoardings` observations with each day (hint: `use sum()`). Which day had the most boardings? Is this true for the other routes (`purpleBoardings`, `greenBoardings`, `bannerBoardings`)?

```{r}
circ %>%
  group_by(day) %>%
  summarize(n = sum(orangeBoardings, na.rm = TRUE))

circ %>%
  group_by(day) %>%
  summarize(n = sum(purpleBoardings, na.rm = TRUE))

circ %>%
  group_by(day) %>%
  summarize(n = sum(greenBoardings, na.rm = TRUE))

circ %>%
  group_by(day) %>%
  summarize(n = sum(bannerBoardings, na.rm = TRUE))
```
 
