---
title: "Data Class Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Part 1

Load all the libraries we will use in this lab. 

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(jhur)
```

1. Create some data to work with. 

First, create a vector that has class integer called `int_vect` that starts at 1 and goes up to 10 and repeats this sequence 3 times using `rep`  (hint use `seq()`).

```{r}

```

2. Next, create a random vector of 30 values from a set of integers from 1 to 30 called `rand_vect` (hint use `sample()` and set the replace argument to `TRUE`).

Because we are using a random vector let's use the `set.seed()` function to make sure we all have the same result - this code is already in the code chunk for you. Simply create the vector below the `set.seed` line.


```{r}
set.seed(1234)


```

3. Create a vector that repeats `c(TRUE, TRUE, FALSE)` 10 times called `TF_vect`. Also create a vector that repeats `c("TRUE", "TRUE", "FALSE")` 10 times called `TF_vect2`.

```{r}


```

Create a tibble combining these vectors together called `vect_data` using the following code.

```{r, eval = FALSE}
vect_data <- tibble(int_vect, rand_vect, TF_vect, TF_vect2)
```


4. Take a look at 5 random rows using the `slice_sample()` function. Try this a few times to see how the results change.

```{r}

```

5. Take a look at 5 random rows using the `slice_sample()` function again but this time with `set.seed(1234)` as the first line of the chunk. Try this a few times to see the results. (Don't forget to not copy the backticks.)

```{r}

```


6. Check to see if the `TF_vect` is logical. Check to see if `TF_vect2` is logical. Why are the results what they are?

```{r}


```


Bonus. Use `mutate()` function to create a new variable in `vect_data` named `type_fact` that is of class `factor` made from the `int_vect` variable. Take a look at the data and observe how the class is different for the new variable compared to `int_vect`.

```{r}



```



# Part 2

1. Read in the Charm City Circulator data using `read_circulator()` function from `jhur` package using the code supplied in the chunk. 

Use the `str()` function to take a look at the data and learn about the column types.

```{r}
circ <- read_circulator()

```

2. Use the `mutate()` function to create a new column named `date_formatted` that is of `Date` class. The new variable is created from `date` column. Hint: use `mdy()` function.

```{r}


```

3. Move the `date_formatted` variable to be before `date` using the `relocate` function. Take a look at the data using `glimpse()`. Note the difference between `date` and `date_formatted` columns.  

```{r}



```

4. Create a list data object called `classes_data` that combines the `vect_data` from the first part of the lab and `circ`. Use `glimpse()` to look at the data.

```{r}



```

Bonus: Use `range()` function on `date_formatted` variable to display the range of dates in the data set. How does this compare to that of `date`? Why? (Hint: use the pull function first to pull the values.)

```{r}


```
 

Bonus: Use the `group_by` function on `day` variable with the `summarize()` function, to display the number of `orangeBoardings` observations with each day (hint: `use sum()`). Which day had the most boardings? Is this true for the other routes (`purpleBoardings`, `greenBoardings`, `bannerBoardings`)?

```{r}



```
 


