---
title: "Data Class Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Part 1

Load all the libraries we will use in this lab. 

```{r}
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(jhur)
```


1\. Create some data to work with by running the following code chunk.

```{r}
set.seed(1234)

int_vect <- rep(seq(from = 1, to = 10), times = 3)
rand_vect <- sample(x = 1:30, size = 30, replace = TRUE)
TF_vect <- rep(c(TRUE, TRUE, FALSE), times = 10)
TF_vect2 <- rep(c("TRUE", "TRUE", "FALSE"), times = 10)
```


2\. Determine the class of each of these new objects.

```{r}
class(int_vect) # [1] "integer"
class(rand_vect) # [1] "integer"
class(TF_vect) # [1] "logical"
class(TF_vect2) # [1] "character"
```


3\. Are `TF_vect` and `TF_vect2` different classes? Why or why not?

```{r}
# Yes!
# Logical vectors do not have quotes around `TRUE` and `FALSE` values.
```


4\. Create a tibble combining these vectors together called `vect_data` using the following code.

```{r}
vect_data <- tibble(int_vect, rand_vect, TF_vect, TF_vect2)
```


5\. Coerce `rand_vect` to character class using `as.character()`. Save this vector as `rand_char_vect`. How is the output for `rand_vect` and `rand_char_vect` different?

```{r}
rand_char_vect <- as.character(rand_vect)
rand_char_vect # Numbers now have quotation marks
```



6\. Read in the Charm City Circulator data using `read_circulator()` function from `jhur` package using the code supplied in the chunk. Or alternatively using the url link.


- Use the `str()` function to take a look at the data and learn about the column types.

```{r}
circ <- read_circulator()
circ <- read_csv(file = "http://jhudatascience.org/intro_to_r/data/Charm_City_Circulator_Ridership.csv")
str(circ)
```


7\. Use the `mutate()` function to create a new column named `date_formatted` that is of `Date` class. The new variable is created from `date` column. Hint: use `mdy()` function. Reassign to `circ`.

```
# General format
NEWDATA <- OLD_DATA %>% mutate(NEW_COLUMN = OLD_COLUMN)
```

```{r}
circ <- mutate(circ, date_formatted = mdy(date))
```



**Bonus / Extra practice**: Move the `date_formatted` variable to be before `date` using the `relocate` function. Take a look at the data using `glimpse()`. Note the difference between `date` and `date_formatted` columns.

```{r}
circ <- circ %>% relocate(date_formatted, .before = date)
glimpse(circ)
```


**Bonus / Extra practice**: Use `range()` function on `date_formatted` variable to display the range of dates in the data set. How does this compare to that of `date`? Why? (Hint: use the pull function first to pull the values.)

```{r}
pull(circ, date_formatted) %>% range()
pull(circ, date) %>% range()
```

> The max of `pull(circ, date) %>% range()` is numerical not based on date.



 
