---
title: "Data Cleaning"
output:
  ioslides_presentation:
    css: ../styles.css
    widescreen: yes
---
 

```{r, echo = FALSE, message = FALSE}
library(knitr)
opts_chunk$set(comment = "")
library(readr)
suppressPackageStartupMessages(library(dplyr))
library(tidyverse)
```


## Data Cleaning{.emphasized}

In general, data cleaning is a process of investigating your data for inaccuracies, or recoding it in a way that makes it more manageable.

::: {style="color: red;"}
`r emo::ji("warning")` MOST IMPORTANT RULE - LOOK `r emo::ji("look")` AT YOUR DATA! `r emo::ji("warning")`
:::


# Dealing with Missing Data

## Missing data types
One of the most important aspects of data cleaning is missing values.  

Types of "missing" data:

* `NA` - general missing data
* `NaN` - stands for "**N**ot **a** **N**umber", happens when you do 0/0.
* `Inf` and `-Inf` - Infinity, happens when you take a positive number (or negative number) by 0.

## Finding Missing data

Each missing data type has a function that returns `TRUE` if the data is missing:

* `NA` - `is.na`
* `NaN` - `is.nan`
* `Inf` and `-Inf` - `is.infinite`

## Useful checking functions{.small}

- `is.na` - is `TRUE` if the data is `FALSE` otherwise
- `!` - negation (NOT) 
    - if `is.na(x)` is `TRUE`, then `!is.na(x)` is `FALSE`

- `any` will be `TRUE` if ANY are true
    - `any(is.na(x))` - do we have any `NA`'s in `x`?

```{r isna}
A = c(1, 2, 4, NA)
B = c(1, 2, 3, 4)
any(is.na(A)) # are there any NAs - YES/TRUE
any(is.na(B)) # are there any NAs- NO/FALSE
```


<!-- ## Complete.cases -->

<!-- `complete.cases` - returns `TRUE` if EVERY value of a row is NOT `NA` -->
<!--     - very stringent condition  -->
<!--     - `FALSE` missing one value (even if not important) -->


<!-- ```{r} -->
<!-- complete.cases(mtcars) -->
<!-- ``` -->



## naniar
Sometimes you need to look at lots of data though... the [`naniar` package](https://cran.r-project.org/web/packages/naniar/vignettes/getting-started-w-naniar.html) is a good option.

The `pct_complete()` function shows the percentage that is complete for a given data object.

```{r, error=FALSE}
#install.packages("naniar")
library(naniar)
x = c(0, NA, 2, 3, 4, -0.5, 0.2)
naniar::pct_complete(x)
```

## Air quality data

The airquality dataset comes with R about air quality in New York in 1973.

```{r, message=FALSE}
?airquality # use this to find out more about the data
airqual <-tibble(airquality)
airqual
```

## naniar: `pct_complete()`{.small}

```{R}
pct_complete(airquality)
```

## Naniar plots

The `gg_miss_var()` function creates a nice plot about the number of missing values for each variable.

```{r, fig.height=4, warning=FALSE, fig.align='center'}
naniar::gg_miss_var(airqual)
```


## Naniar plots

We can use the `facet` argument to make more plots about a specific variable.

```{r, fig.height=4, warning=FALSE, fig.align='center'}
naniar::gg_miss_var(airqual, facet = Month)
```

## Missing Data Issues

Recall that mathematical operations with `NA` often result in `NA`s.
```{r}
sum(c(1,2,3,NA))
mean(c(2,4,NA))
median(c(1,2,3,NA))
```

```{r, echo = FALSE}
red_blue <- tibble(color = rep(c("red", "blue", NA), 3))
red_blue <- red_blue %>% count(color)
red_blue <- rename(red_blue, col_count = n)
```




## Missing Data Issues

This is also true for logicals.  This is a good thing. The NA data could be `> 2` or not, we don't know,
so `R` says there is no `TRUE` or `FALSE`, so that is missing.
```{r}
x = c(0, NA, 2, 3, 4, -0.5, 0.2)
x > 2
```




<!-- ## Missing Data with Logicals -->

<!-- What to do?  What if we want if `x > 2` and `x` isn't `NA`? -->
<!-- Don't do `x != NA`, do `x > 2` and `x` is NOT `NA`: -->

<!-- ```{r} -->
<!-- x > 2 & x != NA -->
<!-- x > 2 & !is.na(x) -->
<!-- ``` -->

<!-- ## Missing Data with Logicals -->

<!-- What about seeing if a value is equal to multiple values?  You can do `(x == 1 | x == 2) & !is.na(x)`, but that is not efficient.   -->

<!-- ```{r} -->
<!-- (x == 0 | x == 2) # has NA -->
<!-- (x == 0 | x == 2) & !is.na(x) # No NA -->
<!-- ``` -->

<!-- what to do? -->

## filter() and missing data

Be careful with missing data using subsetting: 

`filter()` removes missing values by default.
To keep them need to add `is.na()`:

```{r}
x # looks like the 1st and 3rd element should be TRUE
x %in% c(0, 2) # uh oh - not good!
x %in% c(0, 2) | is.na(x) # do this

```


## filter() and missing data{.codesmall}
```{r, echo = FALSE}
df <-tibble(Dog = c(0, NA, 2, 3, 1, 1), 
            Cat= c(NA, 8, 6, NA, 2, NA))
```

```{r}
df

df %>% filter(Dog < 3)
```

## to remove rows with NAs for one variable use `drop_na()`

Avoid using filter for NA values. Instead use `drop_na()`

```{R}
df %>% drop_na(Dog)
```

!NA` does not work as you might expect because you can't tell if something is not actually `NA` - R doesn't ever assume to know what the value of `NA` is 

```{r}
NA == NA
NA != NA
```



## tidyr::drop_na{.codesmall}

This function will drop rows with **any** missing data in **any** column when used on a df.

```{r}
df
drop_na(df)
```


## Think about `NA`

Sometimes removing `NA` values leads to distorted math - be careful! 
Think about what your `NA` means for your data (are you sure ?).

Is an `NA` for values so low they could not be reported?
Or is it this and also if there was a different issue? 

## Think about `NA`

If it is something more like a zero then you might want it included in your data like a zero.

Example:
- survey reports NA if student has never tried cigarettes
- survey reports 0 if student has tried cigarettes but did not smoke that week


You might want to keep the NA values so that you know the original sample size.

 
## Word of caution{.codesmall}

Calculating percentages will give you a different result depending on your choice to include NA values.

```{r}
red_blue
red_blue %>% mutate(percent = 
                      col_count/sum(pull(red_blue, col_count)))
```

## Word of caution{.codesmall}
```{r}
red_blue %>% mutate(percent = 
                      col_count/sum(pull(drop_na(red_blue), col_count)))

# Should you be dividing by 9 or 6? It depends on your data
# Pay attention to your data and your NAs!
```

<!-- ## Missing Data with Operations -->

<!-- Similarly with logicals, operations/arithmetic with `NA` will result in `NA`s: -->

<!-- ```{r} -->
<!-- x = c(0, NA, 2, 3, 4, -0.5, 0.2) -->
<!-- x + 2 -->
<!-- x * 2 -->
<!-- ``` -->


## Check values{.codesmall}

Check the values for your variables, are they what you expect?

`count()` is a great option because it gives tells you:

1) The unique values
2) the amount of these values

Check if rare values make sense

```{r, message=FALSE}
bike <-jhur::read_bike()

bike %>% count(subType)

```



<!-- ## Useful checking functions -->

<!--   - `table(x, useNA = "ifany")` - will have row NA if there are NA values -->
<!--   - `table(x, useNA = "always")` - will have row NA even if there are no NA values -->

<!-- Here we will use `table` to make tabulations of the data.   Look at `?table` to see options for missing data. -->

<!-- ```{r table} -->
<!-- z = c("A", "B", "A", "B") -->
<!-- table(z, useNA = "ifany") -->
<!-- table(z, useNA = "always") -->
<!-- ``` -->


<!-- ## Creating One-way Tables  -->

<!-- `useNA = "ifany"` will not have NA in table heading if no `NA`: -->
<!-- ```{r onetab_ifany} -->
<!-- table(c(0, 1, 2, 3, 2, 3, 3, 2, 2, 3),  -->
<!--         useNA = "ifany") -->
<!-- tibble(x = c(0, 1, 2, 3, 2, 3, 3, 2, 2, 3)) %>%  count(x) -->
<!-- ``` -->

<!-- ## Creating One-way Tables  -->

<!-- You can set `useNA = "always"` to have it always have a column for `NA` -->
<!-- ```{r onetab} -->
<!-- table(c(0, 1, 2, 3, 2, 3, 3, 2, 2, 3),  -->
<!--         useNA = "always") -->
<!-- ``` -->

<!-- ## Tables with Factors  -->

<!-- If you use a `factor`, all levels will be given even if no exist!  -->
<!--   - (May be wanted or not): -->

<!-- ```{r onetab_fact, error = TRUE} -->
<!-- fac = factor(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3), -->
<!--              levels = 1:4) -->
<!-- tab = table(fac) -->
<!-- tab -->
<!-- tab[ tab > 0 ] -->
<!-- ``` -->


<!-- ## Factors cont -->
<!-- ```{r} -->
<!-- tibble(x = fac) %>% count(x) -->
<!-- ``` -->


<!-- ## Finding Row or Column Totals -->

<!-- `margin.table` finds the marginal sums of the table.  `margin` is 1 for rows, 2 for columns in general in `R`.  Here is the column sums of the table: -->

<!-- ```{r margin} -->
<!-- tab -->
<!-- margin.table(tab, 2) -->
<!-- ``` -->

<!-- ## Proportion Tables  -->

<!-- `prop.table` finds the marginal proportions of the table. Think of it dividing the table by it's respective marginal totals.  If `margin` not set, divides by overall total. -->

<!-- ```{r table2} -->
<!-- prop.table(tab) -->
<!-- prop.table(tab,1) * 100 -->
<!-- ``` -->

## Lab Part 1

[lab part 1](http://jhudatascience.org/intro_to_r/Data_Cleaning/lab/Data_Cleaning_Lab.Rmd)

[Website](http://jhudatascience.org/intro_to_r/index.html)





<!-- ```{r readSal_csv, echo= FALSE, eval = TRUE} -->
<!-- Sal = read_csv("https://jhudatascience.org/intro_to_r/data/Baltimore_City_Employee_Salaries_FY2015.csv", col_types = cols( -->
<!--   name = col_character(), -->
<!--   JobTitle = col_character(), -->
<!--   AgencyID = col_character(), -->
<!--   Agency = col_character(), -->
<!--   HireDate = col_character(), -->
<!--   AnnualSalary = col_character(), -->
<!--   GrossPay = col_character() -->
<!-- )) -->
<!-- Sal = rename(Sal, Name = name) -->
<!-- ``` -->

<!-- ## Checking for logical conditions  -->

<!-- * `any()` - checks if there are any `TRUE`s -->
<!-- * `all()` - checks if ALL are true -->

<!-- ```{r isna} -->
<!-- head(Sal,2) -->
<!-- any(is.na(Sal$Name)) # are there any NAs? -->
<!-- ``` -->

# Recoding Variables

## Example of Recoding

```{r, echo=FALSE}
data_diet <- tibble(Diet = rep(c("A", "B", "B"),
                               times = 4), 
                    Gender = c("Male", 
                               "m",
                               "Other",
                               "F", 
                               "Female",
                               "M", 
                               "f",
                               "O", 
                               "Man",
                               "f",
                               "F",
                               "O"), 
                    Weight_start = sample(100:250, size = 12),
                    Weight_change = sample(-10:20, size = 12))

```

Say we have some data about samples in a diet study:

```{r}
data_diet
```

## Oh dear...

This needs lots of recoding.

```{r}
data_diet %>%
  count(Gender, Diet)
```

## `dplyr` can help!

Using Excel to find all of the different ways `gender` has been coded, would be a matter of filtering and changing all by hand or using if statements. This can be hectic!

In `dplyr` you can use the `recode` function (need `mutate` here too!):


```{r, eval = FALSE}
# General Format - this is not code!
{data_input} %>%
  mutate({variable_to_fix} = {Variable_fixing, {old_value} = {new_value},
                                       {another_old_value} = {new_value})

```



```{r, eval = FALSE}

data_diet %>% 
  mutate(Gender = recode(Gender, M = "Male", 
                                 m = "Male", 
                               Man = "Male",
                                 O = "Other",
                                 f = "Female",
                                 F = "Female")) %>%
  count(Gender, Diet)
```

## Or you can use `case_when()`. 


The `case_when()` function of `dplyr` can help us to do this as well.

```{r, eval = FALSE}
# General Format - this is not code!
{data_input} %>%
  mutate({variable_to_fix} = case_when{Variable_fixing}condition 
                                              ~ {value_for_cond}))

```

Note that automatically values not reassigned explicitly by case_when will be `NA`.


## Use of `case_when()`
```{r}
data_diet %>% 
  mutate(Gender = case_when(Gender =="M" ~ "Male"))
```


## More complicated case_when()

```{r}
data_diet %>% 
  mutate(Gender = case_when(
    Gender %in% c("M", "male", "Man", "m", "Male") ~ "Male",
    Gender %in% c("F", "Female", "f", "female")~ "Female",
    Gender %in% c("O", "Other") ~ "Other"))

```

## Another reason for `case_when()`

`case_when` can do very sophisticated comparisons

```{r}

data_diet <-data_diet %>% 
      mutate(Effect = case_when(Weight_change > 0 ~ "Increase",
                                Weight_change == 0 ~ "Same",
                                Weight_change < 0 ~ "Decrease"))

head(data_diet)

```


```{r, echo= FALSE}
data_diet %>% 
  count(Diet, Effect)
```

```{r, echo = FALSE}
diet_effect <-data_diet %>% 
  count(Diet, Effect)

data_diet %>% 
  count(Diet, Effect)%>%
  ggplot(aes(x = Effect,y = n, fill = Diet)) + 
  geom_col(position = position_dodge()) +
  labs(y = "Individuals", 
       title = "Effect of diet A & B on participants")


```

## What if our data looked like this?

```{r, echo = FALSE}
diet_comb<-diet_effect %>% unite("change" , Diet:Effect, remove = TRUE)
```

```{r}
diet_comb
```

<!-- ## Example of Cleaning: more complicated -->

<!-- Sometimes though, it's not so simple.  That's where functions that find patterns come in very useful. -->

<!-- ```{r gender, echo=FALSE} -->
<!-- set.seed(4) # random sample below - make sure same every time -->
<!-- gender <- sample(c("Male", "mAle", "MaLe", "M", "MALE", "Ma", "FeMAle", "F", "Woman", "Man", "Fm", "FEMALE"), 1000, replace = TRUE) -->
<!-- ``` -->

<!-- ```{r gentab} -->
<!-- table(gender) -->
<!-- ``` -->


<!-- ## Example of Cleaning: more complicated -->

<!-- ```{r gender2, echo=FALSE} -->
<!-- gender = gender %>%  -->
<!--   tolower %>%  -->
<!--   recode(m = "Male", f = "Female", ma = "Male", -->
<!--          woman = "Female", man = "Male") -->
<!-- ``` -->

<!-- ```{r gentab2} -->
<!-- table(gender) -->
<!-- ``` -->

## Separating columns based on a separator 

* From `tidyr`, you can split a data set into multiple columns:

```{r}
diet_comb %>% 
  separate(change, into = c("Diet", "Change"))
```

## Separating columns based on a separator 
You can specify the separator with `sep`.

```{r, echo = FALSE}
diet_comb<-diet_effect %>% unite("change" , Diet:Effect, remove = TRUE, sep = "_diet ")
``` 

```{r}
diet_comb %>% 
  separate(change, into = c("Diet", "Change"), sep = " ")
```

## Uniting columns based on a separator 

* From `tidyr`, you can unite:

```{r unite_df}
df = tibble(id = rep(1:5, 3), visit = rep(1:3, each = 5))
head(df, 4)
```


```{r}
df_united <- df %>% unite(col = "unique_id", id, visit, sep = "_")
head(df_united, 4)
```



# Strings functions
## Splitting/Find/Replace and Regular Expressions

* R can do much more than find exact matches for a whole string!


## The `stringr` package

The `stringr` package:

* Modifying or finding **part** or all of a character string
* We will not cover `grep` or `gsub` - base R functions
    - are used on forums for answers
* Almost all functions start with `str_*`


##  `stringr`

`str_detect`, and `str_replace` search for matches to argument pattern within each element of a character vector (not data frame or tibble!).

* `str_detect` - returns `TRUE` if `pattern` is found
<!-- * `str_subset` - returns only the strings which pattern were detected -->
<!--     * convenient wrapper around `x[str_detect(x, pattern)]` -->
<!-- * `str_extract` - returns only strings which pattern were detected, but ONLY the pattern -->
* `str_replace` - replaces `pattern` with `replacement` 

## Download Salary FY2014 Data

From https://data.baltimorecity.gov/City-Government/Baltimore-City-Employee-Salaries-FY2015/nsfe-bg53, from  https://data.baltimorecity.gov/api/views/nsfe-bg53/rows.csv

Read the CSV into R `Sal`:

```{r readSal, echo = TRUE, eval = TRUE, message=FALSE, warning=FALSE}
Sal = jhur::read_salaries() # or
```

```{r}
head(Sal)
```


<!-- ## 'Find' functions: Finding Logicals -->

<!-- These are the indices where the pattern match occurs: -->

<!-- ```{r RawlMatch_log} -->
<!-- head(str_detect(Sal$name, "Rawlings")) -->
<!-- ``` -->

<!-- ## 'Find' functions: Finding Indices -->

<!-- These are the indices where the pattern match occurs: -->

<!-- ```{r RawlMatch} -->
<!-- which(str_detect(Sal$Name, "Rawlings")) -->
<!-- ``` -->

<!-- ## Showing difference in `str_extract` -->

<!-- `str_extract` extracts just the matched string -->

<!-- ```{r ggrep2} -->
<!-- ss = str_extract(Sal$Name, "Rawling") -->
<!-- head(ss) -->
<!-- ss[ !is.na(ss)] -->
<!-- ``` -->

<!-- str_subset(Sal$Name, "Rawlings") -->


## 'Find'`str_detect()` function: finding values: `stringr` 

```{r}
Sal %>% filter(str_detect(name, "Rawlings"))
```



## Showing differnce in `str_replace` and  `str_replace_all`

`str_replace` replaces only the first instance.

```{r, warning = FALSE}
head(pull(Sal, JobTitle))
head(str_replace(pull(Sal, JobTitle), "II", "2"))
```

`str_replace` replaces all instances.
```{r}
head(str_replace_all(pull(Sal, name), "a", "j"), 2)
```


## Lab Part 2

[lab part 2](http://jhudatascience.org/intro_to_r/Data_Cleaning/lab/Data_Cleaning_Lab.Rmd)

[Website](http://jhudatascience.org/intro_to_r/index.html)

# Extra Slides

##  String Splitting

* `str_split(string, pattern)` - splits strings up - returns list!

```{r str_split_orig}
library(stringr)
x <- c("I really like writing R code")
df = tibble(x = c("I really", "like writing", "R code programs"))
y <- unlist(str_split(x, " "))
y
length(y)
```

## A bit on Regular Expressions

* http://www.regular-expressions.info/reference.html
* They can use to match a large number of strings in one statement
* `.` matches any single character
* `*` means repeat as many (even if 0) more times the last character
* `?` makes the last thing optional
* `^` matches start of vector `^a` - starts with "a"
* `$` matches end of vector `b$` - ends with "b"

## Let's look at modifiers for `stringr`

`?modifiers`

* `fixed` - match everything exactly
<!-- * `regex` - default - uses **reg**ular **exp**ressions -->
* `ignore_case` is an option to not have to use `tolower`

## Using a fixed expression

One example case is when you want to split on a period "`.`".  In regular expressions `.` means **ANY** character, so we need to specify that we want R to interpret "." as simply a period.

```{r}
str_split("I.like.strings", ".")
str_split("I.like.strings", fixed("."))
str_split("I.like.strings", "\\.")
```

<!-- ## Let's extract from `y` -->

<!-- ```{r stsplit2} -->
<!-- y[[2]] -->
<!-- # sapply(y, dplyr::first) # on the fly -->
<!-- # sapply(y, nth, 2) # on the fly -->
<!-- # sapply(y, last) # on the fly -->
<!-- ``` -->




<!-- ## Separating columns based on a separator  -->

<!-- * From `tidyr`, you can split a data set into multiple columns: -->

<!-- ```{r separate_df2_for_easy} -->
<!-- df = tibble(x = c("I really", "like writing", "R code programs")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df %>% separate(x, into = c("first", "second")) -->
<!-- ``` -->


<!-- ## Separating columns based on a separator  -->

<!-- * `extra = "merge"` will not drop data.  Also, you can specify the separator -->

<!-- ```{r separate_df2} -->
<!-- df = tibble(x = c("I really", "like. _writing R. But not", "R code programs")) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- df %>% separate(x, into = c("first", "second", "third"), extra = "merge") -->
<!-- ``` -->




<!-- ## Showing differnce in `str_extract` and  `str_extract_all` -->

<!-- `str_extract_all` extracts all the matched strings -->
<!-- - `\\d` searches for DIGITS/numbers -->

<!-- ```{r } -->
<!-- head(str_extract(Sal$AgencyID, "\\d")) -->
<!-- head(str_extract_all(Sal$AgencyID, "\\d"), 2) -->
<!-- ``` -->


<!-- ## Replace -->

<!-- Let's say we wanted to sort the data set by Annual Salary: -->

<!-- ```{r classSal} -->
<!-- class(Sal$AnnualSalary) -->
<!-- ``` -->

<!-- ```{r destringSal} -->
<!-- head(Sal$AnnualSalary, 4) -->
<!-- head(as.numeric(Sal$AnnualSalary), 4) -->
<!-- ``` -->

<!-- R didn't like the `$` so it thought turned them all to `NA`. -->

<!-- ## Replacing and substituting  -->

<!-- Now we can replace the `$` with nothing (used `fixed("$")` because `$` means ending): -->

<!-- ```{r orderSal} -->
<!-- Sal = Sal %>% mutate( -->
<!--   AnnualSalary = str_replace(AnnualSalary, fixed("$"), ""), -->
<!--   AnnualSalary = as.numeric(AnnualSalary) -->
<!--   ) %>%  -->
<!--   arrange(desc(AnnualSalary)) -->
<!-- ``` -->



## Pasting strings with `paste` and `paste0`

Paste can be very useful for joining vectors together:

```{r Paste}
paste("Visit", 1:5, sep = "_")
paste("Visit", 1:5, sep = "_", collapse = "_")
# and paste0 can be even simpler see ?paste0 
paste0("Visit",1:5) # no space!
```

<!-- ```{r } -->
<!-- paste("Visit", 1:5, sep = "_") -->
<!-- paste("Visit", 1:5, sep = "_", collapse = "_") -->
<!-- paste("To", "is going be the ", "we go to the store!", sep = "day ") -->
<!-- # and paste0 can be even simpler see ?paste0  -->
<!-- paste0("Visit",1:5) # no space! -->
<!-- ``` -->




!-- # Before Cleaning - Subsetting with Brackets -->

<!-- ## Select specific elements using an index -->

<!-- Often you only want to look at subsets of a data set at any given time.  Elements of an R object are selected using the brackets (`[` and `]`). -->

<!-- For example, `x` is a vector of numbers and we can select the second element of `x` using the brackets and an index (2): -->

<!-- ```{r} -->
<!-- x = c(1, 4, 2, 8, 10) -->
<!-- x[2] -->
<!-- ``` -->

<!-- <!-- `dplyr`: --> -->

<!-- <!-- ```{r} --> -->
<!-- <!-- nth(x, n = 2) --> -->
<!-- <!-- ``` --> -->

<!-- ## Select specific elements using an index -->

<!-- We can select the fifth or second AND fifth elements below: -->

<!-- ```{r} -->
<!-- x = c(1, 2, 4, 8, 10) -->
<!-- x[5] -->
<!-- x[c(2,5)] -->
<!-- ``` -->


<!-- ```{r, error=TRUE} -->
<!-- nth(x, n = c(2, 5)) # nth only returns one number -->
<!-- ``` -->


<!-- ## Subsetting by deletion of entries -->

<!-- You can put a minus (`-`) before integers inside brackets to remove these indices from the data. -->

<!-- ```{r negativeIndex} -->
<!-- x[-2] # all but the second -->
<!-- ``` -->

<!-- Note that you have to be careful with this syntax when dropping more than 1 element: -->

<!-- ```{r negativeIndex2} -->
<!-- x[-c(1,2,3)] # drop first 3 -->
<!-- # x[-1:3] # shorthand. R sees as -1 to 3 -->
<!-- x[-(1:3)] # needs parentheses -->
<!-- ``` -->

<!-- ## Select specific elements using logical operators -->

<!-- What about selecting rows based on the values of two variables? We use logical statements.  Here we select only elements of `x` greater than 2: -->

<!-- ```{r} -->
<!-- x -->
<!-- x > 2 -->
<!-- x[ x > 2 ] -->
<!-- ``` -->

<!-- ## Select specific elements using logical operators -->
<!-- You can have multiple logical conditions using the following: -->

<!-- * `&` : AND -->
<!-- * `|` : OR -->

<!-- ```{r} -->
<!-- x[ x > 2 & x < 5 ] -->
<!-- x[ x > 5 | x == 2 ] -->
<!-- ``` -->


<!-- ## which function -->

<!-- The `which` functions takes in logical vectors and returns the index for the elements where the logical value is `TRUE`.   -->

<!-- ```{r} -->
<!-- which(x > 5 | x == 2) # returns index -->
<!-- x[ which(x > 5 | x == 2) ] -->
<!-- x[ x > 5 | x == 2 ] -->
<!-- ``` -->

## Using Regular Expressions
* Look for any name that starts with:
    * Payne at the beginning, 
    * Leonard and then an S
    * Spence then capital C

```{r grepstar_stringr}
head(str_subset( Sal$name, "^Payne.*"), 3)
```

```{r grepstar2_stringr}
head(str_subset( Sal$name, "Leonard.?S"))
head(str_subset( Sal$name, "Spence.*C.*"))
```


<!-- ## Uniting columns based on a separator  -->

<!-- * From `tidyr`, you can unite: -->

<!-- ```{r unite_df2} -->
<!-- df = tibble(id = rep(1:5, 3), visit = rep(1:3, each = 5)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df %>% unite(col = "unique_id", id, visit, sep = "_", remove = FALSE) -->
<!-- ``` -->

<!-- ## Paste Depicting How Collapse Works -->

<!-- ```{r Paste2} -->
<!-- paste(1:5) -->
<!-- paste(1:5, collapse = " ") -->
<!-- ``` -->

<!-- ## Useful String Functions -->

<!-- Useful String functions -->

<!-- * `toupper()`, `tolower()` - uppercase or lowercase your data: -->
<!-- * `str_trim()` (in the `stringr` package) or `trimws` in base  -->
<!--     - will trim whitespace -->
<!-- * `nchar` - get the number of characters in a string -->



<!-- ## Sorting characters -->

<!-- - `sort` - reorders the data - characters work, but not correctly -->
<!-- - `rank` - gives the rank of the data - ties are split -->
<!-- - `order` - gives the indices, if subset, would give the data sorted -->
<!--     - `x[order(x)]` is the same as sorting -->

<!-- ```{r orderrank} -->
<!-- sort(c("1", "2", "10")) #  not sort correctly (order simply ranks the data) -->
<!-- order(c("1", "2", "10")) -->
<!-- x = rnorm(10) -->
<!-- x[1] = x[2] # create a tie -->
<!-- rank(x) -->
<!-- ``` -->






















# Comparison of `stringr` to base R - not covered






# Splitting Strings

## Substringing

`stringr`

<!-- * `str_sub(x, start, end)` - substrings from position start to position end -->
* `str_split(string, pattern)` - splits strings up - returns list!


## Splitting String:

In `stringr`, `str_split` splits a vector on a string into a `list`

```{r}
x <- c("I really", "like writing", "R code programs")
y <- stringr::str_split(x, pattern =  " ") # returns a list
y
```



## `str_extract`

`str_extract` extracts  matched strings
- `\\d` searches for DIGITS/numbers

```{r}
head(Sal$AgencyID)
head(str_extract(Sal$AgencyID, "\\d"))
```



<!-- ## 'Find' functions: base R -->

<!-- `grep`: `grep`, `grepl`, `regexpr` and `gregexpr` search for matches to argument pattern within each element of a character vector: they differ in the format of and amount of detail in the results.  -->

<!-- `grep(pattern, x, fixed=FALSE)`, where: -->

<!-- * pattern = character string containing a regular expression to be matched in the given character vector. -->

<!-- * x = a character vector where matches are sought, or an object which can be coerced by as.character to a character vector. -->

<!-- * If fixed=TRUE, it will do exact matching for the phrase anywhere in the vector (regular find) -->


## 'Find' functions: stringr compared to base R

Base R does not use these functions.  Here is a "translator" of the `stringr` function to base R functions

* `str_detect` - similar to `grepl` (return logical)
* `grep(value = FALSE)` is similar to `which(str_detect())`
* `str_subset` - similar to `grep(value = TRUE)` - return value of matched
* `str_replace` - similar to `sub` - replace one time 
* `str_replace_all` - similar to `gsub` - replace many times

## Important Comparisons

Base R:

* Argument order is `(pattern, x)`
* Uses option `(fixed = TRUE)`

`stringr`

* Argument order is `(string, pattern)` aka `(x, pattern)`
* Uses function `fixed(pattern)`

## 'Find' functions: Finding Indices

These are the indices where the pattern match occurs:

```{r}
grep("Rawlings",Sal$Name)
which(grepl("Rawlings", Sal$Name))
which(str_detect(Sal$Name, "Rawlings"))
```

## 'Find' functions: Finding Logicals

These are the indices where the pattern match occurs:

```{r}
head(grepl("Rawlings",Sal$Name))
head(str_detect(Sal$Name, "Rawlings"))
```


## 'Find' functions: finding values, base R 

```{r}
grep("Rawlings",Sal$Name,value=TRUE)
Sal[grep("Rawlings",Sal$Name),]
```



## Showing differnce in `str_extract`

`str_extract` extracts just the matched string

```{r}
ss = str_extract(Sal$Name, "Rawling")
head(ss)
ss[ !is.na(ss)]
```

## Showing differnce in `str_extract` and  `str_extract_all`

`str_extract_all` extracts all the matched strings

```{r}
head(str_extract(Sal$AgencyID, "\\d"))
head(str_extract_all(Sal$AgencyID, "\\d"), 2)
```



## Using Regular Expressions
* Look for any name that starts with:
    * Payne at the beginning, 
    * Leonard and then an S
    * Spence then capital C

```{r}
head(grep("^Payne.*", x = Sal$name, value = TRUE), 3)
```

```{r}
head(grep("Leonard.?S", x = Sal$name, value = TRUE))
head(grep("Spence.*C.*", x = Sal$name, value = TRUE))
```



## Using Regular Expressions: `stringr`
```{r}
head(str_subset( Sal$name, "^Payne.*"), 3)
```

```{r}
head(str_subset( Sal$name, "Leonard.?S"))
head(str_subset( Sal$name, "Spence.*C.*"))
```

## Replace

Let's say we wanted to sort the data set by Annual Salary:

```{r}
class(Sal$AnnualSalary)
```

```{r}
sort(c("1", "2", "10")) #  not sort correctly (order simply ranks the data)
order(c("1", "2", "10"))
```

## Replace

So we must change the annual pay into a numeric:
```{r}
head(Sal$AnnualSalary, 4)
head(as.numeric(Sal$AnnualSalary), 4)
```

R didn't like the `$` so it thought turned them all to `NA`.

`sub()` and `gsub()` can do the replacing part in base R.

## Replacing and subbing 

Now we can replace the `$` with nothing (used `fixed=TRUE` because `$` means ending):

```{r}
Sal$AnnualSalary <- as.numeric(gsub(pattern = "$", replacement="", 
                              Sal$AnnualSalary, fixed=TRUE))
Sal <- Sal[order(Sal$AnnualSalary, decreasing=TRUE), ] 
Sal[1:5, c("name", "AnnualSalary", "JobTitle")]
```

## Replacing and subbing: `stringr` 

We can do the same thing (with 2 piping operations!) in dplyr
```{r}
dplyr_sal = Sal
dplyr_sal = dplyr_sal %>% mutate( 
  AnnualSalary = AnnualSalary %>%
    str_replace(
      fixed("$"), 
      "") %>%
    as.numeric) %>%
  arrange(desc(AnnualSalary))
check_Sal = Sal
rownames(check_Sal) = NULL
all.equal(check_Sal, dplyr_sal)
```


## Website

[Website](../index.html)

# Extra slides

## Creating Two-way Tables

A two-way table.  If you pass in 2 vectors, `table` creates a 2-dimensional table.

```{r}
tab <- table(c(0, 1, 2, 3, 2, 3, 3, 2,2, 3),
             c(0, 1, 2, 3, 2, 3, 3, 4, 4, 3),
              useNA = "always")
tab
```

## Creating Two-way Tables

```{r}
tab_df = tibble(x = c(0, 1, 2, 3, 2, 3, 3, 2,2, 3),
             y = c(0, 1, 2, 3, 2, 3, 3, 4, 4, 3))
tab_df %>% count(x, y)
```

## Creating Two-way Tables 

```{r}
tab_df %>% 
  count(x, y) %>% 
  group_by(x) %>% mutate(pct_x = n / sum(n))
```

## Creating Two-way Tables 

```{r, message = FALSE}
library(scales)
tab_df %>% 
  count(x, y) %>% 
  group_by(x) %>% mutate(pct_x = percent(n / sum(n)))
```
