---
title: "Functions"
output:
  ioslides_presentation:
    css: ../../docs/styles.css
    widescreen: yes
---

```{r, echo = FALSE, message = FALSE}
library(dplyr)
library(jhur)
library(knitr)
library(stringr)
library(tidyr)
library(emo)
opts_chunk$set(comment = "")
```

## Writing your own functions

So far we've seen many functions, like `c()`, `class()`, `filter()`, `dim()` ...

**Why create your own functions?**

- Cut down on repetitive code (easier to fix things!)
- Organize code into manageable chunks
- Avoid running code unintentionally
- Use names that make sense to you

## Writing your own functions

Here we will write a function that multiplies some number (x) by 2:

```{r comment=""}
times_2 <- function(x) x * 2
```

When you run the line of code above, you make it ready to use (no output yet!). Let's test it!

```{r comment=""}
times_2(x = 10)
```

## Writing your own functions: `{ }`

Adding the curly brackets - `{}` - allows you to use functions spanning multiple lines:

```{r comment=""}
times_2 <- function(x) {
  x * 2
}
times_2(x = 10)
```

## Writing your own functions: `return`

If we want something specific for the function's output, we use `return()`:

```{r comment=""}
times_2 <- function(x) {
  output <- x * 2
  return(output)
}
times_2(x = 10)
```

## Writing your own functions

The general syntax for a function is: 

```
functionName <- function(inputs) {
 <function body>
return(value)
}
```

## Writing your own functions: multiple inputs

Functions can take multiple inputs:

```{r comment=""}
times_2_plus_y <- function(x, y) x * 2 + y
times_2_plus_y(x = 10, y = 3)
```

## Writing your own functions: defaults

Functions can have "default" arguments. This lets us use the function without using an argument later:

```{r comment=""}
times_2_plus_y <- function(x = 10, y = 3) x * 2 + y
times_2_plus_y()
```

## Writing another simple function

Let's write a function, `sqdif`, that:

1. takes two numbers `x` and `y` with default values of 2 and 3.
2. takes the difference
3. squares this difference
4. then returns the final value 

## Writing another simple function

```{r comment=""}
sqdif <- function(x = 2, y = 3) (x - y)^2

sqdif()
sqdif(x = 10, y = 5)
sqdif(10, 5)
```

## Writing your own functions: characters

Functions can have any kind of input. Here is a function with characters:

```{r comment=""}
loud <- function(word) {
  output <- rep(toupper(word), 5)
  return(output)
}
loud(word = "hooray!")
```

## Functions for tibbles

We can use `filter(row_number()==n)` to extract a row of a tibble:

```{r message=FALSE}
cars <- read_kaggle()

get_row <- function(dat, row) dat %>% filter(row_number() == row)
```

```{r echo=FALSE}
# So extra columns don't clutter the slide
cars <- cars %>% select(1:10)
```

```{r}
get_row(dat = cars, row = 10)
```

```{r echo=FALSE}
# Restore dataset
cars <- read_kaggle()
```

## Functions for tibbles

`select(n)` will choose column `n`:

```{r message=FALSE}
get_index <- function(dat, row, col) {
  dat %>%
    filter(row_number() == row) %>%
    select(col)
}

get_index(dat = cars, row = 10, col = 8)
```

## Functions for tibbles

Including default values for arguments:

```{r message=FALSE}
get_top <- function(dat, row = 1, col = 1) {
  dat %>%
    filter(row_number() == row) %>%
    select(col)
}

get_top(dat = cars)
```

## Using your custom functions: `sapply()`

Now that you've made a function... You can "apply" functions easily with `sapply()`!

These functions take the form:

```   
sapply(<a vector or list>, some_function)
```

## Using your custom functions: `sapply()`

`r emo::ji("rotating_light")` There are no parentheses on the functions! `r emo::ji("rotating_light")`

```{r comment=""}
sapply(cars, class)
```

## Using your custom functions: `sapply()`

```{r}
sapply(pull(cars, VehOdo), times_2_plus_y)
```

## Using your custom functions "on the fly" to iterate

```{r comment=""}
sapply(pull(cars, VehOdo), function(x) x / 1000)
```

## Applying functions with `across` from `dplyr`

`across()` makes it easy to apply the same transformation to multiple columns. Usually used with `summarize()`.

```
across( .cols = <columns>, .fns = function, ... )
```

- List columns first : `.cols = `
- List function next: `.fns = `
- Then list any arguments for the function

## Applying functions with `across` from `dplyr`.{.codesmall}

Combining with `summarize()`:

```{r warning=FALSE}
cars_dbl <- cars %>% select(Make, Model, where(is.double))

cars_dbl %>%
  group_by(Make) %>%
  summarize(across(.cols = everything(), .fns = mean))
```

## Applying functions with `across` from `dplyr`.{.codesmall}

Combining with `summarize()`:

```{r warning=FALSE}
# Adding arguments to the end!
#
cars_dbl %>%
  group_by(Make) %>%
  summarize(across(.cols = everything(), .fns = mean, na.rm = TRUE))
```

## Applying functions with `across` from `dplyr`.{.codesmall}

Using different `tidyselect()` options:

```{r warning=FALSE}
cars_dbl %>%
  group_by(Make) %>%
  summarize(across(.cols = starts_with("Veh"), .fns = mean))
```

## Applying functions with `across` from `dplyr`.{.codesmall}

Combining with `mutate()`:

```{r}
cars_dbl %>%
  mutate(across(.cols = starts_with("Veh"), .fns = round, digits = -3))
```

## Applying functions with `across` from `dplyr`.{.codesmall}

Combining with `mutate()`:

```{r}
cars_dbl %>%
  mutate(across(
    .cols = everything(),
    .fns = str_replace_all,
    pattern = "A",
    replacement = "a"
  ))
```

## Applying functions with `across` from `dplyr`.{.codesmall}

Combining with `mutate()`:

```{r warning=FALSE, message=FALSE}
# Child mortality data
mort <- read_mortality() %>% rename(country = `...1`)

mort %>%
  select(country, starts_with("194")) %>%
  mutate(across(
    .cols = c(`1943`, `1944`, `1945`),
    .fns = replace_na,
    replace = 0
  ))
```

## Summary

- Simple functions take the form:
  - `NEW_FUNCTION <- function(x, y) x + y ..`
  - Can specify defaults like `function(x = 1, y = 2)`
- Apply your functions with `sapply(<a vector or list>, some_function)`
- Use `across()` to apply functions across multiple columns of data

## Website

🏠 [Class Website](https://jhudatascience.org/intro_to_r/)  

💻 [Lab](https://jhudatascience.org/intro_to_r/modules/Functions/lab/Functions_Lab.Rmd)

The end!
