---
title: "Materials + Schedule"
output:
  html_document
---

```{r, child="_child_schedule.Rmd"}

```

### Day Overview

| Time (EST)        | Content   |
|-------------------|-----------|
| 8:30am - 9:30am   | Session 1 |
| 9:30am - 9:40am   | Break     |
| 9:40am - 10:40am  | Session 2 |
| 10:40am - 10:50am | Break     |
| 10:50am - 11:50am | Session 3 |

### Detailed Schedule + Materials

```{r, results='asis', echo = FALSE, message=FALSE, warning = FALSE}
# The str_replace_all creates an absolute url to linked documents since there are word and pdf rendered versions of the syllabus
source("../scripts/utils.R")
pander::pandoc.table(
  read_markdown("_schedule_table.Rmd")  %>%
    mutate(across(
      .fns = ~ str_replace_all(., "\\./", "https://jhudatascience.org/intro_to_r/")
    )),
  missing = "",
  split.tables = Inf,
  style = "rmarkdown"
)
```

### Module Details

## [Day 0 (before the first session)](https://jhudatascience.org/intro_to_r/Syllabus/module_details/day0.html)
 - Homework 1

## [Day 1](https://jhudatascience.org/intro_to_r/Syllabus/module_details/day1.html)

- Introduction 
- RStudio
- Reproducible Research

## [Day 2](https://jhudatascience.org/intro_to_r/Syllabus/module_details/day2.html)

- Basic R: Variables/Objects in R
- Data Input/Output


## [Day 3](https://jhudatascience.org/intro_to_r/Syllabus/module_details/day3.html)

- Subsetting Data
- Discuss Homework 2 - work on Homework 2 in class if there is time


## [Day 4](https://jhudatascience.org/intro_to_r/Syllabus/module_details/day4.html)

- Summarization
- Data Classes


## [Day 5](https://jhudatascience.org/intro_to_r/Syllabus/module_details/day5.html)


- Data Cleaning
- Discuss Homework 3 - work on Homework 3 in class if there is time


## [Day 6](https://jhudatascience.org/intro_to_r/Syllabus/module_details/day6.html)

- Data Manipulation
- Data Visualization with Esquisse


## [Day 7](https://jhudatascience.org/intro_to_r/Syllabus/module_details/day7.html)

- Data Visualization Continued
- Factors

## [Day 8](https://jhudatascience.org/intro_to_r/Syllabus/module_details/day8.html)

- Statistics
- Discuss and work on Final Project


## [Day 9](https://jhudatascience.org/intro_to_r/Syllabus/module_details/day9.html)

- Functions
- Work on Final Project
