---
title: "Factors Lab"
output: html_document
editor_options: 
  chunk_output_type: console
---

0. Load all the libraries we will use in this lab. 

```{r message=FALSE}
library(dplyr)
library(ggplot2)
```

0. Load the Youth Tobacco Survey data (using `jhur::read_yts()`). `select` "Sample_Size",  "Education", and "LocationAbbr". Name this data "yts". 

```{r}
yts <- jhur::read_yts() %>% select(Sample_Size, Education, LocationAbbr)

```

1. Create a boxplot showing the difference in "Sample_Size" between Middle School and High School "Education". **Hint**: Use `aes(x = Education, y = Sample_Size)` and `geom_boxplot()`.

```{r}

```

2. Use `count` to count up the number of observations of data for each "Education" group.

```{r}

```

3. Make "Education" a factor using the `mutate` and `factor` functions. Use the `levels` argument inside `factor` to reorder "Education". Reorder this variable so that "Middle School" comes before "High School". Assign the output the name "yts_fct".

```{r}

```

4. Repeat #1 and #2 using the "yts_fct" data. You should see different ordering in the plot and `count` table.

```{r}

```

Bonus/optional practice. Convert "LocationAbbr" (state) in "yts_fct" into a factor using the `mutate` and `factor` functions. Do not add a `levels =` argument.

```{r}

```

Bonus/optional practice. We want to create a new column that contains the group-level median sample size. 

- Using the "yts_fct" data, `group_by` "LocationAbbr". 
- Then, use `mutate` to create a new column "med_sample_size" that is the median "Sample_Size". 
- **Hint**: Since you have already done `group_by`, a median "Sample_Size" will automatically be created for each unique level in "LocationAbbr". Use the `median` function with `na.rm = TRUE`.

```{r}

```

Bonus/optional practice. We want to plot the "LocationAbbr" (state) by the "med_sample_size" column we created above. Using the `forcats` package, create a plot that:

- Has "LocationAbbr" on the x-axis
- Uses the `mapping` argument and the `fct_reorder` function to order the x-axis by "med_sample_size"
- Has "Sample_Size" on the y-axis
- Is a boxplot (`geom_boxplot`)
- Has the x axis label of `State`
(Don't worry if you get a warning about not being able to plot `NA` values.)

Save your plot using `ggsave()` with a width of 10 and height of 3.

Which state has the largest median sample size?

```{r}

```
