---
title: "Statistics Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Part 1

1. Load the libraries needed in this lab. Read in the following child mortality data using `read_mortality()` function from `jhur` package. Assign it to `mort` variable. Change its first column name `X1` into `country`

https://jhudatascience.org/intro_to_R_class/data/mortality.csv
```{r}
library(tidyverse)
library(broom)
library(jhur)
library(psych)

mort = read_mortality()

mort = mort %>% 
  rename(country = X1)
```

2. Compute the correlation between the `1980`, `1990`, `2000`, and `2010` mortality variables.
(No need to save this in an object. Just display the result to the screen.) 

Hint: you can use `select()` function to first subset the data frame to keep the four columns only. Hint 2: to use column name in `select()` that starts with a number, surround it with back quotes. 

```{r}
mort_sub = 
  mort %>% 
  select(`1980`, `1990`, `2000`, `2010`)

cor(mort_sub)
cor(mort_sub, use = "complete.obs")
```

```{r}
# alternatively
mort_sub = mort[, c("1980", "1990", "2000", "2010")]
cor(mort_sub, use = "complete.obs")
```

3. BONUS. Compute the correlation between the `Myanmar`, `China`, and `United States` mortality data over the years. 

Note: column names are in 1st column of `mort` data frame.

```{r}
# make data frame that has data from the three countries only
mort_sub_contry = mort %>% 
  filter(country %in% c( "China", "Myanmar", "United States")) 

# remove country column
mortality_only <- 
  mort_sub_contry %>%
  select(!country)

# transpose and rename 
mortality_only_transposed <- as.data.frame(t(mortality_only))
colnames(mortality_only_transposed) <- c( "China", "Myanmar", "United States")

# compute correlation matrix
country_cor <- cor(mortality_only_transposed, use = "complete.obs")
country_cor
```

# Part 2

1. Read in the Kaggle cars auction dataset from "http://jhudatascience.org/intro_to_R_class/data/kaggleCarAuction.csv" with the argument `col_types = cols(VehBCost = col_double())`:

```{r}
cars = read_csv("http://jhudatascience.org/intro_to_R_class/data/kaggleCarAuction.csv",
                col_types = cols(VehBCost = col_double()))

```

2. Fit a linear regression model with vehicle cost (`VehBCost`) as the outcome and 
vehicle age (`VehicleAge`) and whether it's an online sale (`IsOnlineSale`) as predictors.  

Save the model fit in an object called `lmfit_cars` and display the summary table.

```{r}
lmfit_cars = lm(VehBCost ~ VehicleAge + IsOnlineSale, data = cars)
summary(lmfit_cars)
```

3. Fit a linear regression model with vehicle cost (`VehBCost`) as the outcome predictors:  
vehicle age (`VehicleAge`), whether it's an online sale (`IsOnlineSale`), and interaction between `VehicleAge` and `IsOnlineSale`.

To specify interaction, use `VehicleAge * IsOnlineSale` in the formula. 

Save the model fit in an object called `lmfit_cars_2` and display the summary table.

```{r}
lmfit_cars_2 = lm(VehBCost ~ VehicleAge * IsOnlineSale, data = cars)
summary(lmfit_cars_2)
```


# Part 3

1. Fit a logistic regression model where the outcome is "bad buy" status and predictors are the cost (`VehBCost`) and vehicle age (`VehicleAge`).

Save the model fit in an object called `logfit_cars` and display the summary table.

Use `summary()` or `tidy(logfit_cars, conf.int = TRUE, exponentiate = TRUE)`
or `tidy(logfit_cars, conf.int = TRUE, exponentiate = FALSE)` for log odds ratios.  

```{r}
logfit_cars = glm(IsBadBuy ~ VehBCost + VehicleAge, data = cars, family = binomial())
summary(logfit_cars)

tidy(logfit_cars, conf.int = TRUE, exponentiate = TRUE)
tidy(logfit_cars, conf.int = TRUE, exponentiate = FALSE)
```

2. Randomly sample `10,000` observations (rows) with replacement from the `cars` dataset and store this new dataset in an object called `cars_subsample`.  Hint: `sample_n` function

```{r}
set.seed(1)
cars_subsample = cars %>% 
  sample_n(size = 10000)
```

3. Fit the same logistic regression model as in problem 1. above  with this sample and display the summary. Call this model `logfit_cars_sub`. 

How do the results compare?  

```{r}
logfit_cars_sub = glm(IsBadBuy ~ VehBCost + VehicleAge, data = cars_subsample, 
                      family = binomial())

summary(logfit_cars_sub)
tidy(logfit_cars_sub, conf.int = TRUE, exponentiate = TRUE)
tidy(logfit_cars, conf.int = TRUE, exponentiate = FALSE)
```
