---
title: "Statistics Lab Key"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1

1. Load the libraries needed in this lab. Read in the following child mortality data using `read_mortality()` function from `jhur` package. Assign it to the "mort" variable. Change its first column name from `...1` into `country`. You can also find the data here: https://jhudatascience.org/intro_to_R_class/data/mortality.csv

```{r}
# library(tidyverse)
# library(broom)
library(jhur)
# library(psych)

mort <- read_mortality()

mort <- mort %>% 
  rename(country = `...1`)
```

2. Compute the correlation between the `2006` and `2007` mortality variables. (No need to save this in an object. Just display the result to the screen.) Use `pull()` function to first extract these columns. To use a column name in `pull()` that starts with a number, surround it with backticks (\`). Then, use the `cor` function.

```{r}
x <- pull(mort, `2006`)
y <- pull(mort, `2007`)

cor(x, y)
```

3. Compute the correlation between the `1980`, `1990`, `2000`, and `2010` mortality variables. (No need to save this in an object. Just display the result to the screen.) Use `select()` function to first subset the data frame to keep the four columns only. To use a column name in `select()` that starts with a number, surround it with backticks (\`). 

```{r}
mort_sub <- 
  mort %>% 
  select(`1980`, `1990`, `2000`, `2010`)

cor(mort_sub)
cor(mort_sub, use = "complete.obs")
```

# Part 2

1. Read in the Kaggle cars auction dataset from "http://jhudatascience.org/intro_to_R_class/data/kaggleCarAuction.csv" with the argument `col_types = cols(VehBCost = col_double())`:

```{r}
cars = read_csv("http://jhudatascience.org/intro_to_R_class/data/kaggleCarAuction.csv",
                col_types = cols(VehBCost = col_double()))

```

2. Fit a linear regression model with vehicle cost (`VehBCost`) as the outcome and 
vehicle age (`VehicleAge`) and whether it's an online sale (`IsOnlineSale`) as predictors.  

Save the model fit in an object called `lmfit_cars` and display the summary table.

```{r}
lmfit_cars = lm(VehBCost ~ VehicleAge + IsOnlineSale, data = cars)
summary(lmfit_cars)
```

3. Fit a linear regression model with vehicle cost (`VehBCost`) as the outcome predictors:  
vehicle age (`VehicleAge`), whether it's an online sale (`IsOnlineSale`), and interaction between `VehicleAge` and `IsOnlineSale`.

To specify interaction, use `VehicleAge * IsOnlineSale` in the formula. 

Save the model fit in an object called `lmfit_cars_2` and display the summary table.

```{r}
lmfit_cars_2 = lm(VehBCost ~ VehicleAge * IsOnlineSale, data = cars)
summary(lmfit_cars_2)
```


# Part 3

1. Fit a logistic regression model where the outcome is "bad buy" status and predictors are the cost (`VehBCost`) and vehicle age (`VehicleAge`).

Save the model fit in an object called `logfit_cars` and display the summary table.

Use `summary()` or `tidy(logfit_cars, conf.int = TRUE, exponentiate = TRUE)`
or `tidy(logfit_cars, conf.int = TRUE, exponentiate = FALSE)` for log odds ratios.  

```{r}
logfit_cars = glm(IsBadBuy ~ VehBCost + VehicleAge, data = cars, family = binomial())
summary(logfit_cars)

tidy(logfit_cars, conf.int = TRUE, exponentiate = TRUE)
tidy(logfit_cars, conf.int = TRUE, exponentiate = FALSE)
```

2. Randomly sample `10,000` observations (rows) with replacement from the `cars` dataset and store this new dataset in an object called `cars_subsample`.  Hint: `sample_n` function

```{r}
set.seed(1)
cars_subsample = cars %>% 
  sample_n(size = 10000)
```

3. Fit the same logistic regression model as in problem 1. above  with this sample and display the summary. Call this model `logfit_cars_sub`. 

How do the results compare?  

```{r}
logfit_cars_sub = glm(IsBadBuy ~ VehBCost + VehicleAge, data = cars_subsample, 
                      family = binomial())

summary(logfit_cars_sub)
tidy(logfit_cars_sub, conf.int = TRUE, exponentiate = TRUE)
tidy(logfit_cars, conf.int = TRUE, exponentiate = FALSE)
```
